### 2020-11-30 abstract.tex
paper structure

\begin{abstract}
    \todo[inline]{TODO}
\end{abstract}

### 2020-11-30 cleave.tex
paper structure

\section{CLEAVE - ControL bEnmArking serVice on the Edge}\label{sec:cleave}

### 2020-11-30 experiments.tex
paper structure

\section{Experimental validation}\label{sec:experiments}

### 2020-12-29 cleave.tex
progress

@@
-\section{CLEAVE - ControL bEnmArking serVice on the Edge}\label{sec:cleave}
+\pagebreak
+\section{The CLEAVE Framework}\label{sec:cleave}
+
+CLEAVE (\emph{ControL bEnchmArking serVice on the Edge}) is our framework for quick, robust, and repeatable benchmarking of networked control systems.
+CLEAVE tackles this challenge through a number of key design decisions that we will discuss in the following.
+
+\begin{description}[wide]
+    \item[Emulation approach.] 
+    Existing benchmarking platforms for NCS either take a fully simulated approach --- running plant, network, and controller inside a completely simulated environment --- or one requiring an actual physical plant that interacts with a real network and controller.
+    CLEAVE walks the line between these two approaches, employing a strategy based on having a real-time emulated representation of a Plant interact with the real network and controller.
+    \item[Plant- and Controller-agnosticism.] 
+    CLEAVE makes as few assumptions as possible about the inner workings of both the Plant and the Controller. 
+    The only requirements for a system to be able to be emulated in CLEAVE are
+    \begin{enumerate*}[itemjoin*={{ and }}]
+        \item that its behavior can be described in a discrete-time fashion;
+        \item that its state and the actions performed on it can be described by a finite number of scalar and/or vector variables.
+    \end{enumerate*}
+
+    This agnosticism is achieved through careful abstraction of the the system into a well-defined API based around the following components:
+    \begin{itemize}
+        \item A \texttt{State} object, which implements the discrete-time evolution of the physical system being controlled.
+        This object also holds a number of named properties that can be measured or actuated upon.
+        \item An arbitrary number of \texttt{Sensor} objects which measure named properties of the \texttt{State}, and potentially transform them (for instance, to add noise).
+        \item A \texttt{Controller} object which receives values of named properties processed by the \texttt{Sensor} objects and returns new values for the named properties of the \texttt{State} that can be actuated upon.
+        \item Finally, an arbitrary number of \texttt{Actuator} objects which receive the values generated by the \texttt{Controller} and process them before passing them on the \texttt{State}.
+    \end{itemize}
+    Users implement their desired behavior by extending the abstract base classes defining these components provided by the framework.
+
+\end{description}
+
+\todo[inline]{Fill in core concepts}
+\todo[inline]{Link to repo}
+
+\subsection{General architecture}
+
+The general architecture of CLEAVE centers around two components:
+
+\begin{enumerate}
+    \item the Plant, corresponding to the emulated physical system running on some arbitrary client device;
+    \item the Controller Service running on some arbitrary cloudlet, which receives samples of the state of the Plant and produces control commands;
+\end{enumerate}
+
+These components are connected through the real network
+
+In the following, we will briefly discuss the design and implementation of these components.
+
+\subsubsection{Plant}
+
+As
+
+\subsubsection{Controller Service}
+\subsection{Network}

### 2020-12-30 cleave.tex
more

@@
 
-    This agnosticism is achieved through careful abstraction of the the system into a well-defined API based around the following components:
+    \item[API based on well-defined components.]
+    The previously mentioned agnosticism is achieved through careful abstraction of the the system into an API based around the following components:
     \begin{itemize}
@@
 
+    \item[Plain Python configuration.]
+    As mentioned above, users implement their desired behavior for the core components by extending classes provided by the framework.
+    This is done in configuration files written in full-featured Python simply defining a number of required top-level variables and which are then passed to the main script provided with CLEAVE.
+    This allows for easy configuration of arbitrarily complex emulations.
+
 \end{description}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3cm]{example-image-a}
+    \caption{Execution flow of a time step in CLEAVE.}
+\end{figure}
+
+\begin{listing}
+    \begin{minted}[linenos=true,
+        breaklines, 
+        breakafter=d,
+        fontsize=\small]{Python}
+from cleave.api.plant import SimpleConstantActuator, SimpleSensor
+from cleave.impl.inverted_pendulum import InvPendulumState
+
+host = 'localhost'
+port = 8080
+
+controller_class = 'InvPendulumController'
+
+tick_rate = 100
+emu_duration = '30m'
+
+state = InvPendulumState(fail_angle_rad=-1)
+
+sensors = [
+    SimpleSensor('position', 100),
+    SimpleSensor('speed', 100),
+    SimpleSensor('angle', 100),
+    SimpleSensor('ang_vel', 100),
+]
+
+actuators = [
+    SimpleConstantActuator(initial_value=0, prop_name='force')
+]
+
+output_dir = './plant_metrics'
+    \end{minted}
+    \caption{Example configuration file.}
+    \label{lst:config}
+\end{listing}
 

### 2021-01-06 cleave.tex
start describing experiments

@@
 
-host = 'localhost'
-port = 8080
+host = '0.0.0.0'  # address of dispatch server
+port = 8080       # providing Controllers
 
+# Controller parameters
 controller_class = 'InvPendulumController'
+controller_params = {'ref': 0.0}
 
+# Plant emulation parameters 
 tick_rate = 100
 emu_duration = '30m'
-
-state = InvPendulumState(fail_angle_rad=-1)
+state = InvPendulumState(fail_angle_rad=0.34)
+output_dir = './plant_metrics'
 
@@
 ]
-
-output_dir = './plant_metrics'
     \end{minted}
-    \caption{Example configuration file.}
+    \caption{
+        Example configuration file.
+        Users are simply required to define a number of top-level variables, but are otherwise free to include arbitrary code.
+        This allows for the easy extension of emulations.
+    }
     \label{lst:config}
 \end{listing}
-
-\todo[inline]{Fill in core concepts}
-\todo[inline]{Link to repo}
-
-\subsection{General architecture}
-
-The general architecture of CLEAVE centers around two components:
-
-\begin{enumerate}
-    \item the Plant, corresponding to the emulated physical system running on some arbitrary client device;
-    \item the Controller Service running on some arbitrary cloudlet, which receives samples of the state of the Plant and produces control commands;
-\end{enumerate}
-
-These components are connected through the real network
-
-In the following, we will briefly discuss the design and implementation of these components.
-
-\subsubsection{Plant}
-
-As
-
-\subsubsection{Controller Service}
-\subsection{Network}

### 2021-01-06 experiments.tex
start describing experiments

@@
 \section{Experimental validation}\label{sec:experiments}
+
+In this section, we validate the utility of the CLEAVE framework through scalability measurements of a networked control system running on a WiFi link.
+With this, we aim to answer questions about the ability of CLEAVE to provide relevant and accurate metrics on the performance of such setups as \emph{system load} increases.
+In this context, we will focus on the load as it manifests on the network link in terms of the reliability of the link and the associated latencies. 
+We adopt this constrained view of load due to the characteristics of NCS, where controllers are usually relatively lightweight in terms of computing resources, leaving the network link as the principal bottleneck in the system.
+
+Our experimental setup is depicted in \cref{fig:expsetup}.
+A number between 1 and 12 of CLEAVE Plants running on dedicated light-weight general purpose computing devices (in this case, Raspberry Pi 4's) connect wirelessly to a WiFi Access Point (AP).
+This AP in turn connects via Ethernet to a host on which the CLEAVE Dispatcher and Controllers are executed.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3cm]{example-image-a}
+    \caption{Experiment setup}
+    \label{fig:expsetup}
+\end{figure}

### 2021-01-07 cleave.tex
rewriting

@@
 \pagebreak
+\section{Approach}
+
+A system for the benchmarking of Networked Control Systems needs to be able to 
+\todo[inline]{Explain what it needs}
+Existing platforms take one of two approaches.
+The first is a fully simulated approach, in which the plant, network and controller reside in a completely simulated environment.
+This allows for unparalleled flexibility, as parts of the system can be swapped out or modified without greater difficulty.
+However, this comes at the cost of realism, as these models are always approximations of reality; and time, as these simulations are often very computationally expensive.
+
+The alternative is an approach using a testbed comprised of a \emph{physical} plant that interacts with a \emph{real} network and controller.
+This of course allows for great realism, but provides very little flexibility in terms of the components of the system, limiting the approach to whatever plant, network and controller are available.
+
+CLEAVE tries to mitigate the shortcomings of these approaches by walking the line between them.
+We employ a third strategy based on having a real-time emulated representation of a control system interacting with a real network.
+This allows for great flexibility when it comes to the components of the control system, as these are as easily switched out as in the fully simulated approach, while maintaining the realism of arguably the most complex and limiting component of an NCS, namely the network itself.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{images/CLEAVE_NCS_structure.png}
+    \caption{
+        Structure of an emulated NCS on top of CLEAVE.
+        The State object implements the discrete-time evolution of the Plant.
+        For this purpose, it holds two sets of special variables; Actuated Variables and Sensed Variables.
+        Actuated Variables are updated from the values obtained from the Actuator objects at the beginning of each time step.
+        These are used to perform a time step update in the State, and update the values of the Sensed Variables.
+        The values of the Sensed Variables are then, at the end of each time step, sampled and passed on to the Sensor objects, which in turn process them (for instance, to add noise) and send the processed values to the Controller.
+        The Controller uses these values to calculate the new actuation commands and sends these to the Actuators.
+        Finally, the Actuators process the actuation commands and hold the new values for the Actuated Variables until they are read at the next time step.
+    }
+\end{figure*}
+
+Concretely, in the following we describe the procedure for benchmarking an NCS on the CLEAVE framework:
+
+\begin{enumerate}
+    \item A discrete-time emulation of the target Plant is implemented.
+    CLEAVE provides a simple API for this task in the form of an \emph{abstract base class} \mintinline{Python}{class State(abc.ABC)} users need to extend.
+    This class provides an abstract method \mintinline{Python}{State.advance(delta_t: float)} which is called by CLEAVE on each emulation time step with the number of seconds since the previous iteration.
+    Additionally, the class provides functionality for the assigning special instance variables that can be sampled from or actuated upon during the emulation.
+
+    \item A Controller for the Plant is developed.
+    Once again, CLEAVE provides an API for this in the form of the \mintinline{Python}{class Controller(abc.ABC)} abstract base class.
+    Users need to extend this class and implement the \mintinline[breaklines]{Python}{Controller.process(sensor_values: Mapping) -> Mapping} method, which will be called with a mapping from property names to sampled values and should return a mapping from property names to actuated values.
+
+    \item A configuration file for the benchmarking run is written.
+    Deployments in CLEAVE are configured through fully-featured Python scripts which simply need to define a number of required (and some optional) top-level variables.
+    These top-level variables include, for instance, a \mintinline[breaklines]{Python}{state} variable pointing to an instance of the target Plant emulation implementation and the name of the Controller class to use for the run, among other emulation parameters.
+
+\end{enumerate}
+
 \section{The CLEAVE Framework}\label{sec:cleave}

### 2021-01-07 cleave.tex
first draft of approach

@@
 
-A system for the benchmarking of Networked Control Systems needs to be able to 
-\todo[inline]{Explain what it needs}
-Existing platforms take one of two approaches.
+% A system for the benchmarking of Networked Control Systems needs to be able to 
+% \todo[inline]{Explain what it needs}
+Existing platforms for NCS benchmarking take one of two approaches.
 The first is a fully simulated approach, in which the plant, network and controller reside in a completely simulated environment.
@@
 This allows for great flexibility when it comes to the components of the control system, as these are as easily switched out as in the fully simulated approach, while maintaining the realism of arguably the most complex and limiting component of an NCS, namely the network itself.
+
+In the following we will explain the approach taken in CLEAVE by walking the reader through the steps necessary to deploy an arbitrary NCS emulation on top of the framework.
+An overview of the structure of an NCS emulation on CLEAVE can be seen in \cref{fig:cleave:ncs:struct}.
 
@@
         Finally, the Actuators process the actuation commands and hold the new values for the Actuated Variables until they are read at the next time step.
-    }
+    }\label{fig:cleave:ncs:struct}
 \end{figure*}
 
-Concretely, in the following we describe the procedure for benchmarking an NCS on the CLEAVE framework:
+\subsection{Defining a Plant}
 
-\begin{enumerate}
-    \item A discrete-time emulation of the target Plant is implemented.
-    CLEAVE provides a simple API for this task in the form of an \emph{abstract base class} \mintinline{Python}{class State(abc.ABC)} users need to extend.
-    This class provides an abstract method \mintinline{Python}{State.advance(delta_t: float)} which is called by CLEAVE on each emulation time step with the number of seconds since the previous iteration.
-    Additionally, the class provides functionality for the assigning special instance variables that can be sampled from or actuated upon during the emulation.
+The first step in developing an NCS emulation on top of CLEAVE is the definition of a Plant which represents the physical system to be controlled.
 
-    \item A Controller for the Plant is developed.
-    Once again, CLEAVE provides an API for this in the form of the \mintinline{Python}{class Controller(abc.ABC)} abstract base class.
-    Users need to extend this class and implement the \mintinline[breaklines]{Python}{Controller.process(sensor_values: Mapping) -> Mapping} method, which will be called with a mapping from property names to sampled values and should return a mapping from property names to actuated values.
-
-    \item A configuration file for the benchmarking run is written.
-    Deployments in CLEAVE are configured through fully-featured Python scripts which simply need to define a number of required (and some optional) top-level variables.
-    These top-level variables include, for instance, a \mintinline[breaklines]{Python}{state} variable pointing to an instance of the target Plant emulation implementation and the name of the Controller class to use for the run, among other emulation parameters.
-
-\end{enumerate}
-
-\section{The CLEAVE Framework}\label{sec:cleave}
-
-CLEAVE (\emph{ControL bEnchmArking serVice on the Edge}) is our framework for quick, robust, and repeatable benchmarking of networked control systems.
-CLEAVE tackles this challenge through a number of key design decisions that we will discuss in the following.
-
-\begin{description}[wide]
-    \item[Emulation approach.] 
-    Existing benchmarking platforms for NCS either take a fully simulated approach --- running plant, network, and controller inside a completely simulated environment --- or one requiring an actual physical plant that interacts with a real network and controller.
-    CLEAVE walks the line between these two approaches, employing a strategy based on having a real-time emulated representation of a Plant interact with the real network and controller.
-    \item[Plant- and Controller-agnosticism.] 
-    CLEAVE makes as few assumptions as possible about the inner workings of both the Plant and the Controller. 
-    The only requirements for a system to be able to be emulated in CLEAVE are
-    \begin{enumerate*}[itemjoin*={{ and }}]
-        \item that its behavior can be described in a discrete-time fashion;
-        \item that its state and the actions performed on it can be described by a finite number of scalar and/or vector variables.
+In abstract terms, Plants in CLEAVE can be understood to be composed of
+\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
+    \item\label{item:plant:state} an emulation of a physical system which is updated in a discrete-time fashion
+    \item\label{item:plant:actuators} procedures to modify or act upon the physical emulation at each time-step
+    \item\label{item:plant:sensors} procedures to sample relevant values from the physical emulation at each time-step
+    \item\label{item:plant:network} procedures to interact with the network and send and receive data from the Controller.
 \end{enumerate*}
 
-    \item[API based on well-defined components.]
-    The previously mentioned agnosticism is achieved through careful abstraction of the the system into an API based around the following components:
-    \begin{itemize}
-        \item A \texttt{State} object, which implements the discrete-time evolution of the physical system being controlled.
-        This object also holds a number of named properties that can be measured or actuated upon.
-        \item An arbitrary number of \texttt{Sensor} objects which measure named properties of the \texttt{State}, and potentially transform them (for instance, to add noise).
-        \item A \texttt{Controller} object which receives values of named properties processed by the \texttt{Sensor} objects and returns new values for the named properties of the \texttt{State} that can be actuated upon.
-        \item Finally, an arbitrary number of \texttt{Actuator} objects which receive the values generated by the \texttt{Controller} and process them before passing them on the \texttt{State}.
-    \end{itemize}
-    Users implement their desired behavior by extending the abstract base classes defining these components provided by the framework.
+While \cref{item:plant:network} is handled transparently by CLEAVE, by hiding it through layers of abstraction, the rest of the items need to be provided in more explicit ways by the user.
 
-    \item[Plain Python configuration.]
-    As mentioned above, users implement their desired behavior for the core components by extending classes provided by the framework.
-    This is done in configuration files written in full-featured Python simply defining a number of required top-level variables and which are then passed to the main script provided with CLEAVE.
-    This allows for easy configuration of arbitrarily complex emulations.
+For \cref{item:plant:state}, the user needs to extend an abstract base class \texttt{State} provided by the framework.
+This class provides the base functionality and defines a simple API for the implementation of discrete-time emulations by
+\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
+    \item automatically tracking variables that can be modified and/or sampled by the framework
+    \item defining a single abstract method \texttt{State.advance()} which is invoked by the framework on each time step of the emulation.
+\end{enumerate*}
 
-\end{description}
+\Cref{item:plant:actuators,item:plant:sensors} are in the same way provided by the user through the extension of the \texttt{Actuator} and \texttt{Sensor} abstract base classes, respectively.
+\texttt{Actuator} objects process the outputs of the Controller and update the values of the \texttt{State} accordingly.
+Conversely \texttt{Sensor} objects sample relevant values from the \texttt{State} object, process them, and finally pass them on to the Controller through the network.
+
+\subsection{Defining a Controller}
+
+After the Plant has been defined, users need to implement a Controller for it.
+This is once again done by extending an abstract base class, \texttt{Controller}, provided by CLEAVE.
+This class defines a single abstract method, \texttt{Controller.process()} which is called by the framework whenever samples are provided by the Plant.
+When invoked, it is passed a mapping of property names to sampled values.
+The framework also expects implementations of this method to return yet another mapping, this time of actuated property names to target values.
+Users are otherwise free to implement any kind of logic or process within extending classes.
+
+\subsection{Configuring and running the emulation}
+
+Once the core components for the emulation have been defined and implemented, the emulation needs to be configured to register them with the framework.
+This is done through fully-featured Python scripts which simply define a number of required (and some optional) top-level variables.
+These Python configuration files are then passed as arguments to the \texttt{cleave} bash command provided by the framework, which sets up and executes the emulation.
+The use of Python scripts for configuration allows for a high degree of flexibility and complexity in the definition of emulation setups.
+
+At this point, yet another core component of CLEAVE needs to be mentioned: the Dispatcher server.
+Although Controllers in CLEAVE can be started manually, this quickly becomes cumbersome when performing experimentation with more than a few Plants.
+The CLEAVE Dispatcher server solves this by providing a way for remote deployment of Controllers through a RESTful HTTP service that the Plant connects to (see \cref{fig:cleave:dispatch}).
+
+Example configuration files for a Plant and the Dispatcher are presented in \cref{lst:config:plant,lst:config:dispatcher}.
 
@@
     \centering
-    \includegraphics[width=3cm]{example-image-a}
-    \caption{Execution flow of a time step in CLEAVE.}
+    \includegraphics[width=.85\columnwidth]{images/CLEAVE_dispatcher.png}
+    \caption{
+        Diagram of the interaction between CLEAVE Plant, Dispatcher and Controller.
+        The Dispatcher is instantiated and listens on a specific port on the Controller Host. 
+        As the Plant initializes its emulation, it requests a Controller from the Dispatcher; the Dispatcher spawns the corresponding Controller and returns the host and port on which it is listening.
+        The Plant proceeds to execute its emulation configuration.
+        Finally, on shutdown, the Plant once again communicates with the Dispatcher to request the shutdown of the associated Controller.
+    }\label{fig:cleave:dispatch}
 \end{figure}
@@
     \caption{
-        Example configuration file.
-        Users are simply required to define a number of top-level variables, but are otherwise free to include arbitrary code.
-        This allows for the easy extension of emulations.
+        Example configuration file for a CLEAVE Plant.
     }
-    \label{lst:config}
+    \label{lst:config:plant}
 \end{listing}
+
+\begin{listing}
+    \begin{minted}[linenos=true,
+        breaklines, 
+        breakafter=d,
+        fontsize=\small]{Python}
+from cleave.impl.inverted_pendulum import InvPendulumController
+
+host = 'localhost'
+port = 8080
+
+controllers = {
+    'InvPendulumController': InvPendulumController,
+}
+
+output_dir = './controllers'
+    \end{minted}
+    \caption{
+        Example configuration file for the CLEAVE Dispatcher.
+        The previously defined Controller is registered with the framework and will then be available from the Dispatcher server for use with a Plant.
+    }
+    \label{lst:config:dispatcher}
+\end{listing}

### 2021-01-09 cleave.tex
progress on experiment section

@@
 
+# Sensors to attach to the Plant
 sensors = [
@@
 actuators = [
-    SimpleConstantActuator(initial_value=0, prop_name='force')
+    SimpleConstantActuator(
+        initial_value=0,
+        prop_name='force')
 ]
@@
     \caption{
-        Example configuration file for a CLEAVE Plant.
+        Example configuration file for a CLEAVE Plant, defining the parameters for a single emulation.
+        The ``controller parameters'' define the class of Controller to request from the Dispatcher.
+        ``Plant emulation parameters'' define key emulation parameters, such as the emulation discrete-time update rate (the \emph{tick rate}).
+        The Sensors and Actuators are then provided in lists, each Sensor and Actuator attached to a specific property identified by a string.
+        Additionally, Sensors have a sampling rate in Hz which needs to be provided as an integer.
     }

### 2021-01-09 experiments.tex
first draft of approach
progress on experiment section

@@
 
+\todo[inline]{WIP}
+
 In this section, we validate the utility of the CLEAVE framework through scalability measurements of a networked control system running on a WiFi link.
-With this, we aim to answer questions about the ability of CLEAVE to provide relevant and accurate metrics on the performance of such setups as \emph{system load} increases.
+We aim to answer questions about the ability of CLEAVE to provide relevant and accurate metrics on the performance of such setups as \emph{system load} increases.
 In this context, we will focus on the load as it manifests on the network link in terms of the reliability of the link and the associated latencies. 
@@
 
-Our experimental setup is depicted in \cref{fig:expsetup}.
-A number between 1 and 12 of CLEAVE Plants running on dedicated light-weight general purpose computing devices (in this case, Raspberry Pi 4's) connect wirelessly to a WiFi Access Point (AP).
+Our experimental setup is depicted in \cref{fig:cleave:expsetup}.
+A number between 1 and 12 of CLEAVE Plants running on dedicated light-weight general purpose computing devices (in this case, Raspberry Pi 4's) connect wirelessly to an 802.11n WiFi Access Point (AP).
 This AP in turn connects via Ethernet to a host on which the CLEAVE Dispatcher and Controllers are executed.
@@
     \centering
-    \includegraphics[width=3cm]{example-image-a}
-    \caption{Experiment setup}
-    \label{fig:expsetup}
+    \includegraphics[width=.95\columnwidth]{images/CLEAVE_experiment_setup}
+    \caption{Experiment setup}\label{fig:cleave:expsetup}
 \end{figure}
+
+In terms of the control system deployed on top of CLEAVE for the experiments, we chose a two-dimensional inverted pendulum system.
+This system was chosen due to its relative simplicity and prevalence in the field of automatic control as one of the fundamental examples basically every control engineer has encountered at some point.
+
+The physical system, represented in \cref{fig:invpend}, was implemented using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}.
+For the controller, a proportional-differential strategy was employed, implemented using the framework Controller API and the NumPy numeric computation library~\autocite{harris2020array}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.95\columnwidth]{images/inverted_pendulum.png}
+    \caption{
+        The two-dimensional inverted pendulum system.
+        The cart moves on the X-axis, and the pendulum on top of it swings freely.
+        The objective of the system is to balance the pendulum vertically through the application of horizontal forces on the cart.
+    }\label{fig:invpend}
+\end{figure}

### 2021-12-29 cleave.tex
writing

@@
 This is done through fully-featured Python scripts which simply define a number of required (and some optional) top-level variables.
+An example of such a configuration file can be seen in \cref{lst:config:plant}.
 These Python configuration files are then passed as arguments to the \texttt{cleave} bash command provided by the framework, which sets up and executes the emulation.
 The use of Python scripts for configuration allows for a high degree of flexibility and complexity in the definition of emulation setups.
-
-At this point, yet another core component of CLEAVE needs to be mentioned: the Dispatcher server.
-Although Controllers in CLEAVE can be started manually, this quickly becomes cumbersome when performing experimentation with more than a few Plants.
-The CLEAVE Dispatcher server solves this by providing a way for remote deployment of Controllers through a RESTful HTTP service that the Plant connects to (see \cref{fig:cleave:dispatch}).
-
-Example configuration files for a Plant and the Dispatcher are presented in \cref{lst:config:plant,lst:config:dispatcher}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=.85\columnwidth]{images/CLEAVE_dispatcher.png}
-    \caption{
-        Diagram of the interaction between CLEAVE Plant, Dispatcher and Controller.
-        The Dispatcher is instantiated and listens on a specific port on the Controller Host. 
-        As the Plant initializes its emulation, it requests a Controller from the Dispatcher; the Dispatcher spawns the corresponding Controller and returns the host and port on which it is listening.
-        The Plant proceeds to execute its emulation configuration.
-        Finally, on shutdown, the Plant once again communicates with the Dispatcher to request the shutdown of the associated Controller.
-    }\label{fig:cleave:dispatch}
-\end{figure}
 
@@
 from cleave.impl.inverted_pendulum import InvPendulumState
-
-host = '0.0.0.0'  # address of dispatch server
-port = 8080       # providing Controllers
-
-# Controller parameters
-controller_class = 'InvPendulumController'
-controller_params = {'ref': 0.0}
 
@@
         Example configuration file for a CLEAVE Plant, defining the parameters for a single emulation.
-        The ``controller parameters'' define the class of Controller to request from the Dispatcher.
         ``Plant emulation parameters'' define key emulation parameters, such as the emulation discrete-time update rate (the \emph{tick rate}).
@@
 \end{listing}
-
-\begin{listing}
-    \begin{minted}[linenos=true,
-        breaklines, 
-        breakafter=d,
-        fontsize=\small]{Python}
-from cleave.impl.inverted_pendulum import InvPendulumController
-
-host = 'localhost'
-port = 8080
-
-controllers = {
-    'InvPendulumController': InvPendulumController,
-}
-
-output_dir = './controllers'
-    \end{minted}
-    \caption{
-        Example configuration file for the CLEAVE Dispatcher.
-        The previously defined Controller is registered with the framework and will then be available from the Dispatcher server for use with a Plant.
-    }
-    \label{lst:config:dispatcher}
-\end{listing}

### 2021-12-29 experiments.tex
writing

--- 2020-11-30 experiments.tex
@@
 \section{Experimental validation}\label{sec:experiments}
+
+\todo[inline]{WIP}
+
+In this section, we aim to validate the utility of the CLEAVE framework through a series of experiments.
+
+\todo[inline]{talk about goal of experiments}
+
+Our experimental setup is depicted in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
+We employ a Raspberry Pi 4B single-board computer together with a general-purpose \verb|x86_64|-based workstation.
+These devices are connected through an 802.11n WiFi link.
+For ease of deployment and reparametrization, we deploy our CLEAVE control loop as a pair of Docker~\cite{merkel2014docker} containers\todo{add link to image?}; one for the plant, running on the Rasberry Pi device, and one for the Controller, running on the workstation.
+These instances are then connected through an overlay network configured through Docker.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.95\columnwidth]{images/CLEAVE_experiment_setup}
+    \caption{Experimental setup.
+Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Overlay Network spanning a Raspberry Pi 4B Plant host connected to an x86 Controller host over a 802.11n WiFi link.}\label{fig:cleave:expsetup}
+\end{figure}
+
+We chose a two-dimensional inverted pendulum system for the control system deployed inside CLEAVE for the experiments, 
+This system was chosen due to its relative simplicity and prevalence in the field of automatic control as one of the fundamental examples of linear control.
+The physical system, represented in \cref{fig:invpend}, was implemented using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}.
+For the controller, a proportional-differential strategy was employed, implemented using the framework Controller API and the NumPy numeric computation library~\autocite{harris2020array}.
+
+This setup is then used to run a series of experiments with varying parametrization of the controlled system.
+Specifically, we modify:
+\begin{itemize}
+    \item the sampling rate of the Plant state, and set it to \SIlist[list-units=single,list-final-separator={, or }]{5;10;20;40;60;120}{\hertz};
+    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-units=single,list-final-separator={, or }]{0;25;50;100;200}{\milli\second} after the processing of each sample.
+\end{itemize}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.95\columnwidth]{images/inverted_pendulum.png}
+    \caption{
+        The two-dimensional inverted pendulum system.
+        The cart moves on the X-axis, and the pendulum on top of it swings freely.
+        The objective of the system is to balance the pendulum vertically through the application of horizontal forces on the cart.
+    }\label{fig:invpend}
+\end{figure}

### 2021-12-30 experiments.tex
writing

@@
 
-Our experimental setup is depicted in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
+We depict our experimental setup in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
 We employ a Raspberry Pi 4B single-board computer together with a general-purpose \verb|x86_64|-based workstation.
-These devices are connected through an 802.11n WiFi link.
+These devices are connected to each other by an 802.11n WiFi link.
 For ease of deployment and reparametrization, we deploy our CLEAVE control loop as a pair of Docker~\cite{merkel2014docker} containers\todo{add link to image?}; one for the plant, running on the Rasberry Pi device, and one for the Controller, running on the workstation.
-These instances are then connected through an overlay network configured through Docker.
+These instances are then connected to an overlay network configured through Docker which sits on top of and abstracts away the real network configuration, and allows for transparent communication between the Plant and Controller.
 
@@
 
-We chose a two-dimensional inverted pendulum system for the control system deployed inside CLEAVE for the experiments, 
+We use a two-dimensional inverted pendulum system for the control system deployed inside CLEAVE for the experiments, 
 This system was chosen due to its relative simplicity and prevalence in the field of automatic control as one of the fundamental examples of linear control.
-The physical system, represented in \cref{fig:invpend}, was implemented using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}.
-For the controller, a proportional-differential strategy was employed, implemented using the framework Controller API and the NumPy numeric computation library~\autocite{harris2020array}.
+The physical system, represented in \cref{fig:invpend}, is implemented using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}.
+For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the NumPy numeric computation library~\autocite{harris2020array}.
 
@@
 \begin{itemize}
-    \item the sampling rate of the Plant state, and set it to \SIlist[list-units=single,list-final-separator={, or }]{5;10;20;40;60;120}{\hertz};
+    \item the sampling rate of the Plant state, setting it to \SIlist[list-units=single,list-final-separator={, or }]{5;10;20;40;60;120}{\hertz};
     \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-units=single,list-final-separator={, or }]{0;25;50;100;200}{\milli\second} after the processing of each sample.
 \end{itemize}
+Each one of the \num{30} possible combinations of these parameters is run on the experimental testbed depicted in \cref{fig:cleave:expsetup} for \num{10} repetitions lasting \SI{5}{\minute} each, yielding \SI{25}{\hour} worth of data.
+During each repetition, we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
 

### 2021-12-30 intro.tex
writing

\section{Introduction}\label{sec:intro}

\todo[inline]{change this}

With the advent of the Internet of Things and Industry 4.0, the number and applications of cyber physical systems is ever increasing [1].
Cyber-Physical Systems (CPS) are formed by an integration of computational and physical processes, namely, it comprises of actuators and sensors connected by virtue of a computational process [2]. A Networked Control System (NCS) is composed of a CPS wherein multiple networked actuators and sensors form a part of the same automatic control system [3].
Zhang et al. in [4] proposed a control communication co-design approach wherein the stabilization of an NCS in which telemetry between the sensors and actuators of the plant with a remote controller utilizing a shared communication medium is discussed.
Zoppi et al. in [5] proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS.
Their methodology utilizes joint knowledge of control, computation, and communication.
In their work various architectural elements and the corresponding delays associated with the NCS are modelled.
Multiple experimental parameters and certain observable Key Performance Indicators (KPIs) are defined and utilized in the implementation.
This work however utilizes a physical LEGO [6] based plant for the implementation, preventing instantaneous changes in the plant characteristics and their nature, in terms of flexibility.
We overcome this issue by proposing a completely virtual plant allowing for unparalleled flexibility in changing the plant model and characteristic features of the experiments.
This work experiments over Ethernet and Wireless Local Area Network (WLAN) based networks.
We perform experiments with similar network configurations and additionally benchmark the number of plants which can be simultaneously sustained by the controller at the same time over varying physical sampling frequencies and tick rates.

\todo[inline]{change this}

### 2021-12-30 cleave.tex
adding acronyms

@@
 % \todo[inline]{Explain what it needs}
-Existing platforms for NCS benchmarking take one of two approaches.
+Existing platforms for \acp{NCS} benchmarking take one of two approaches.
 The first is a fully simulated approach, in which the plant, network and controller reside in a completely simulated environment.
@@
 The alternative is an approach using a testbed comprised of a \emph{physical} plant that interacts with a \emph{real} network and controller.
-This of course allows for great realism, but provides very little flexibility in terms of the components of the system, limiting the approach to whatever plant, network and controller are available.
+This of course allows for great realism, but provides very little flexibility in terms of the components of the system, limiting the approach to whatever plant, network, and controller are available.
 
-CLEAVE tries to mitigate the shortcomings of these approaches by walking the line between them.
+Our approach, CLEAVE, tries to mitigate the shortcomings of these approaches by walking the line between them.
 We employ a third strategy based on having a real-time emulated representation of a control system interacting with a real network.
-This allows for great flexibility when it comes to the components of the control system, as these are as easily switched out as in the fully simulated approach, while maintaining the realism of arguably the most complex and limiting component of an NCS, namely the network itself.
+This allows for great flexibility when it comes to the components of the control system, as these are as easily switched out as in the fully simulated approach, while maintaining the realism of arguably the most complex and limiting component of an \ac{NCS}, namely the network itself.
 
-In the following we will explain the approach taken in CLEAVE by walking the reader through the steps necessary to deploy an arbitrary NCS emulation on top of the framework.
-An overview of the structure of an NCS emulation on CLEAVE can be seen in \cref{fig:cleave:ncs:struct}.
+In the following we will explain the approach taken in CLEAVE by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation on top of the framework.
+An overview of the structure of such an emulation on CLEAVE can be seen in \cref{fig:cleave:ncs:struct}.
 
@@
     \caption{
-        Structure of an emulated NCS on top of CLEAVE.
+        Structure of an emulated \ac{NCS} on top of CLEAVE.
         The State object implements the discrete-time evolution of the Plant.
@@
 
-The first step in developing an NCS emulation on top of CLEAVE is the definition of a Plant which represents the physical system to be controlled.
+The first step in developing an \ac{NCS} emulation on top of CLEAVE is the definition of a Plant which represents the physical system to be controlled.
 
@@
 After the Plant has been defined, users need to implement a Controller for it.
-This is once again done by extending an abstract base class, \texttt{Controller}, provided by CLEAVE.
+This is once again done by extending an abstract base class, \texttt{Controller}, provided by CLEAVE.\@
 This class defines a single abstract method, \texttt{Controller.process()} which is called by the framework whenever samples are provided by the Plant.

### 2021-12-30 intro.tex
adding acronyms
drafting intro

@@
 
-\todo[inline]{change this}
+The number and applications of \acp{CPS} --- that is, systems in which a real, physical mechanism is controlled by a computer --- have exploded in later years.
+This rise in popularity is in great part driven by the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge Computing.
+\todo[inline]{references}
 
+In particular, \acp{NCS}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system, have been of great interest to the scientific community in the last decade.
+Due to their potential advantages for industrial settings, a large body of work exists dedicated to the modelling and performance characterization of these systems.
+\acp{NCS} enhance the flexibility of and improve existing control systems by allowing for the distribution of control functions over and across networks.
+This allows for --- for instance --- simultaneous centralized coordination, control, and monitoring of potentially hundreds of devices.
+\todo[inline]{references}
+
+In spite of this, however, we identify a gap in the literature pertaining to the reproducibility and comparison of experimental studies on \ac{NCS}.
+Experimental research in the literature has been conducted on a huge variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
+This stems from the inherent inter-domain nature of \acp{NCS}, which intertwines knowledge from the fields of communications, computing, and control theory.
+This leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
+Additionally, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
+
+\todo[inline]{
 With the advent of the Internet of Things and Industry 4.0, the number and applications of cyber physical systems is ever increasing [1].
@@
 We perform experiments with similar network configurations and additionally benchmark the number of plants which can be simultaneously sustained by the controller at the same time over varying physical sampling frequencies and tick rates.
-
-\todo[inline]{change this}
+}

### 2022-01-01 intro.tex
working

--- 2021-12-30 intro.tex
@@
 
-\todo[inline]{change this}
+\todo[inline]{references}
 
-With the advent of the Internet of Things and Industry 4.0, the number and applications of cyber physical systems is ever increasing [1].
-Cyber-Physical Systems (CPS) are formed by an integration of computational and physical processes, namely, it comprises of actuators and sensors connected by virtue of a computational process [2].
+The number and applications of \acp{CPS}~\cite{Rajkumar2010CPS} --- that is, systems in which a real, physical mechanism is controlled by a computer --- have exploded in later years.
+This rise in popularity is in great part driven by the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge Computing.
+
+In particular, \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system, have been of great interest to the scientific community in the last decade.
+Due to their potential advantages for industrial settings~\cite{Lu2015WSAN}, a large body of work exists dedicated to the modelling and performance characterization of these systems~\cite{Hespanha2007Survey,Zhang2013Survey,Zhang2016Survey}.
+\Acp{NCS} enhance the flexibility of and improve existing control systems by allowing for the distribution of control functions over and across networks.
+This allows for --- for instance --- simultaneous centralized coordination, control, and monitoring of potentially hundreds of devices.
+
+Most of the large literature concerning \acp{NCS}, however, follows a theoretical approach, and only a small fraction of it deals with experimental studies.
+The inherent inter-domain nature of \acp{NCS}, which intertwines knowledge from the fields of communications, computing, and control theory, coupled with the complexity of experimental platforms make these kinds of studies hard to perform.
+Moreover, we identify a gap in knowledge pertaining to the reproducibility and comparison of experimental studies on these systems.
+Existing experimental research includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
+In turn, this leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
+Additionally, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
+
+The most straightforward approach to experimental research in \ac{NCS} use setups in which the complete system uses real hardware.
+\cite{Drew2005NCSWLAN,Baumann2018LowPower} are examples of studies employing this approach.
+\textcite{Drew2005NCSWLAN} present \iac{NCS} model which considers both random packet delay and loss on both the sensor and actuator sides of the feedback loop; \textcite{Baumann2018LowPower}, on the other hand, develop an evaluation methodology for wireless \ac{NCS}.
+Both works validate their results on physical testbeds, consisting of inverted penduli plants controlled wirelessly over WiFi from a central computation point.
+\todo[inline]{More examples?}
+
+Other
+
+\textcite{Zoppi2020NCSBench} made a first attempt at tackling this issue through their \emph{NCSbench} platform.
+\todo[inline]{Describe NCSbench}
+
+% \todo[inline]{
+% With the advent of the Internet of Things and Industry 4.0, the number and applications of cyber physical systems is ever increasing [1].
+% Cyber-Physical Systems (CPS) are formed by an integration of computational and physical processes, namely, it comprises of actuators and sensors connected by virtue of a computational process [2].
 A Networked Control System (NCS) is composed of a CPS wherein multiple networked actuators and sensors form a part of the same automatic control system [3].
-Zhang et al.
+% Zhang et al.
 in [4] proposed a control communication co-design approach wherein the stabilization of an NCS in which telemetry between the sensors and actuators of the plant with a remote controller utilizing a shared communication medium is discussed.
-Zoppi et al. in [5] proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS.
-Their methodology utilizes joint knowledge of control, computation, and communication.
-In their work various architectural elements and the corresponding delays associated with the NCS are modelled.
-Multiple experimental parameters and certain observable Key Performance Indicators (KPIs) are defined and utilized in the implementation.
-This work however utilizes a physical LEGO [6] based plant for the implementation, preventing instantaneous changes in the plant characteristics and their nature, in terms of flexibility.
-We overcome this issue by proposing a completely virtual plant allowing for unparalleled flexibility in changing the plant model and characteristic features of the experiments.
-This work experiments over Ethernet and Wireless Local Area Network (WLAN) based networks.
-We perform experiments with similar network configurations and additionally benchmark the number of plants which can be simultaneously sustained by the controller at the same time over varying physical sampling frequencies and tick rates.
-
-\todo[inline]{change this}
+% Zoppi et al. in [5] proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS.
+% Their methodology utilizes joint knowledge of control, computation, and communication.
+% In their work various architectural elements and the corresponding delays associated with the NCS are modelled.
+% Multiple experimental parameters and certain observable Key Performance Indicators (KPIs) are defined and utilized in the implementation.
+% This work however utilizes a physical LEGO [6] based plant for the implementation, preventing instantaneous changes in the plant characteristics and their nature, in terms of flexibility.
+% We overcome this issue by proposing a completely virtual plant allowing for unparalleled flexibility in changing the plant model and characteristic features of the experiments.
+% This work experiments over Ethernet and Wireless Local Area Network (WLAN) based networks.
+% We perform experiments with similar network configurations and additionally benchmark the number of plants which can be simultaneously sustained by the controller at the same time over varying physical sampling frequencies and tick rates.
+% }

### 2022-01-02 intro.tex
writing more stuff

@@
 In turn, this leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
-Additionally, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
+Furthermore, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
 
 The most straightforward approach to experimental research in \ac{NCS} use setups in which the complete system uses real hardware.
-\cite{Drew2005NCSWLAN,Baumann2018LowPower} are examples of studies employing this approach.
 \textcite{Drew2005NCSWLAN} present \iac{NCS} model which considers both random packet delay and loss on both the sensor and actuator sides of the feedback loop; \textcite{Baumann2018LowPower}, on the other hand, develop an evaluation methodology for wireless \ac{NCS}.
 Both works validate their results on physical testbeds, consisting of inverted penduli plants controlled wirelessly over WiFi from a central computation point.
-\todo[inline]{More examples?}
+Similarly, \textcite{Li2014Wireless} implement a wireless microcontroller-based system for vibration control and validate it on a physical prototype consisting of a cantilever beam controlled over a IEEE 802.15.4 \ac{WPAN} system.
 
-Other
+Conversely, some studies choose to instead use completely \emph{simulated} \ac{NCS} setups~\cite{Andersson2005Simulation,Eyisi2012NCSWT}.
+Such an approach is employed by \textcite{Du2009Smith}; the authors develop a novel \emph{Smith} predictor to compensate for varying latencies in wireless \acp{NCS} and validate it using the \emph{TrueTime}~\cite{Henriksson2002TrueTime} simulator.
+In \textcite{Chen2015synccontrol} the authors validate and evaluate a number of synchronous control strategies for three-motor setups using completely simulated \acp{NCS} environments.
+\citeauthor{Wu2012NPC} introduce the concept of \emph{network predictive control} in\ \cite{Wu2012NPC} and use it to balance a simulated inverted pendulum over a simulated wireless network.
+
+Finally, a number of experimental studies instead employ hardware- and network-in-the-loop approaches.
+Hardware-in-the-loop refers to \ac{NCS} setups in which a real control system interacts with an emulated network.
+Such setups have been used by studies
 

### 2022-01-03 intro.tex
writing more shite

@@
 \textcite{Drew2005NCSWLAN} present \iac{NCS} model which considers both random packet delay and loss on both the sensor and actuator sides of the feedback loop; \textcite{Baumann2018LowPower}, on the other hand, develop an evaluation methodology for wireless \ac{NCS}.
-Both works validate their results on physical testbeds, consisting of inverted penduli plants controlled wirelessly over WiFi from a central computation point.
-Similarly, \textcite{Li2014Wireless} implement a wireless microcontroller-based system for vibration control and validate it on a physical prototype consisting of a cantilever beam controlled over a IEEE 802.15.4 \ac{WPAN} system.
+Both works validate their results on physical testbeds, consisting of inverted penduli plants controlled wirelessly over IEEE 802.11 WiFi from a central computation point.
+Similarly, \textcite{Li2014Wireless} implement a wireless microcontroller-based system for vibration control and validate it on a physical prototype consisting of a cantilever beam controlled over an IEEE 802.15.4 \ac{WPAN} system.
+\textcite{Cuenca2019UAV} design and implement periodic event-triggered sampling and dual-rate control techniques for wireless control, validating their contributions on a four-rotor autonomous drone platform controlled over WiFi.
 
@@
 \citeauthor{Wu2012NPC} introduce the concept of \emph{network predictive control} in\ \cite{Wu2012NPC} and use it to balance a simulated inverted pendulum over a simulated wireless network.
+In \cite{Ma2019DynamicSched}, \citeauthor{Ma2019DynamicSched} propose an optimal dynamic scheduling strategy that optimizes performance of multi-loop control systems and validate it by simulating a four-loop control system over an IEEE 802.15.4 \ac{WPAN}.
 
-Finally, a number of experimental studies instead employ hardware- and network-in-the-loop approaches.
-Hardware-in-the-loop refers to \ac{NCS} setups in which a real control system interacts with an emulated network.
-Such setups have been used by studies
+Finally, a number of experimental studies instead employ \emph{hardware-} and \emph{network-in-the-loop} approaches.
+Hardware-in-the-loop refers to \ac{NCS} setups in which a real network interacts with a simulated or emulated control system.
+This kind of setup is particularly prevalent in the field of \emph{smart grid} control, and has been used by studies such as \textcite{Wang2020VoltageControl}, in which the authors validate a novel three-level coordinated control method for photovoltaic inverters.
+
+On the other hand, network-in-the-loop refers to experimental setups in which the \emph{network} is instead simulated or emulated.
+An example of this approach can be found in \textcite{Natale2004InvPendEthernet}, in which the authors interface an real inverted pendulum system with an emulated IEEE 802.3 Ethernet network to study the effects of the \ac{CSMA/CD} medium access control scheme on plant stability.
 

### 2022-01-03 intro.tex
blah

@@
 \section{Introduction}\label{sec:intro}
-
-\todo[inline]{references}
 
@@
 The inherent inter-domain nature of \acp{NCS}, which intertwines knowledge from the fields of communications, computing, and control theory, coupled with the complexity of experimental platforms make these kinds of studies hard to perform.
-Moreover, we identify a gap in knowledge pertaining to the reproducibility and comparison of experimental studies on these systems.
-Existing experimental research includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
-In turn, this leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
-Furthermore, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
 
@@
 
-\textcite{Zoppi2020NCSBench} made a first attempt at tackling this issue through their \emph{NCSbench} platform.
-\todo[inline]{Describe NCSbench}
+As evidence by the discussion above, experimental research in \acp{NCS} includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
+This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
+Furthermore, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
+We identify thus a gap in knowledge pertaining to the reproducibility and comparison of experimental studies on these systems.
 
-% \todo[inline]{
-% With the advent of the Internet of Things and Industry 4.0, the number and applications of cyber physical systems is ever increasing [1].
-% Cyber-Physical Systems (CPS) are formed by an integration of computational and physical processes, namely, it comprises of actuators and sensors connected by virtue of a computational process [2].
-A Networked Control System (NCS) is composed of a CPS wherein multiple networked actuators and sensors form a part of the same automatic control system [3].
-% Zhang et al.
-in [4] proposed a control communication co-design approach wherein the stabilization of an NCS in which telemetry between the sensors and actuators of the plant with a remote controller utilizing a shared communication medium is discussed.
-% Zoppi et al. in [5] proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS.
-% Their methodology utilizes joint knowledge of control, computation, and communication.
-% In their work various architectural elements and the corresponding delays associated with the NCS are modelled.
-% Multiple experimental parameters and certain observable Key Performance Indicators (KPIs) are defined and utilized in the implementation.
-% This work however utilizes a physical LEGO [6] based plant for the implementation, preventing instantaneous changes in the plant characteristics and their nature, in terms of flexibility.
-% We overcome this issue by proposing a completely virtual plant allowing for unparalleled flexibility in changing the plant model and characteristic features of the experiments.
-% This work experiments over Ethernet and Wireless Local Area Network (WLAN) based networks.
-% We perform experiments with similar network configurations and additionally benchmark the number of plants which can be simultaneously sustained by the controller at the same time over varying physical sampling frequencies and tick rates.
-% }
+\textcite{Zoppi2020NCSBench} made a first attempt at tackling this challenge with their \emph{NCSbench} platform.
+\emph{NCSbench} is an open-source \ac{NCS} benchmarking platform designed with reproducibility in mind, built using the \citetitle{LEGOMindstormsEV3}~\cite{LEGOMindstormsEV3} platform.
+It allows for easy and highly reproducible experimentation and benchmarking of \acp{NCS}, due to its low-cost, accessibility, and ease of reconfiguration.
+
+Although the work of \textcite{Zoppi2020NCSBench} is a relevant and important step towards the reproducibility of \ac{NCS} benchmarks, we believe it still falls short on some aspects.
+In particular, we argue the reliance of \emph{NCSbench} on specific hardware --- accessible as it may be --- hampers the potential of the platform, as it fundamentally limits both the physical systems that can be benchmarked and the scalability of the performed benchmarks.
+
+In this paper, we present thus the first fully-software-based \ac{NCS} benchmarking framework with scalability and repeatability as its main focus.
+This framework is built on top of the \emph{Python 3.8}~\cite{Python3.8} language, making it highly extensible and able to harness the tremendous amount of already existing user-provided libraries and packages~\cite{pypi}.
+Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}~\cite{merkel2014docker}, making it suitable for automated deployment and benchmarking on industry-standard deployment using container orchestration technologies.
+We name this framework \ac{CLEAVE}.
+
+The rest of this paper is structured as follows.

### 2022-01-03 intro.tex
intro

@@
 The rest of this paper is structured as follows.
+\Cref{sec:approach} presents the design principles of the framework.
+Next, in \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.
+Finally, we conclude this paper in \cref{sec:conclusion} with a brief discussion on the strengths and weaknesses of the platform, as well as on future directions to take.

### 2022-01-03 cleave.tex
blah
intro
cleave typo
editing

@@
-\pagebreak
-\section{Approach}
+\section{Approach}\label{sec:approach}
 
-% A system for the benchmarking of Networked Control Systems needs to be able to 
-% \todo[inline]{Explain what it needs}
-Existing platforms for \acp{NCS} benchmarking take one of two approaches.
-The first is a fully simulated approach, in which the plant, network and controller reside in a completely simulated environment.
-This allows for unparalleled flexibility, as parts of the system can be swapped out or modified without greater difficulty.
-However, this comes at the cost of realism, as these models are always approximations of reality; and time, as these simulations are often very computationally expensive.
+In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}.
+Our design follows an approach similar to the above-discussed \emph{hardware-in-the-loop} setups, in that we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
+This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of \acp{NCS}, namely the network itself.
 
-The alternative is an approach using a testbed comprised of a \emph{physical} plant that interacts with a \emph{real} network and controller.
-This of course allows for great realism, but provides very little flexibility in terms of the components of the system, limiting the approach to whatever plant, network, and controller are available.
-
-Our approach, CLEAVE, tries to mitigate the shortcomings of these approaches by walking the line between them.
-We employ a third strategy based on having a real-time emulated representation of a control system interacting with a real network.
-This allows for great flexibility when it comes to the components of the control system, as these are as easily switched out as in the fully simulated approach, while maintaining the realism of arguably the most complex and limiting component of an \ac{NCS}, namely the network itself.
-
-In the following we will explain the approach taken in CLEAVE by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation on top of the framework.
-An overview of the structure of such an emulation on CLEAVE can be seen in \cref{fig:cleave:ncs:struct}.
+In the following we will explain the approach taken in \ac{CLEAVE} by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation on top of the framework.
+An overview of the structure of such an emulation on \ac{CLEAVE} can be seen in \cref{fig:cleave:ncs:struct}.
 
@@
 
-The first step in developing an \ac{NCS} emulation on top of CLEAVE is the definition of a Plant which represents the physical system to be controlled.
+The first step in developing an \ac{NCS} emulation on top of \ac{CLEAVE} is the definition of a Plant which represents the physical system to be controlled.
 
-In abstract terms, Plants in CLEAVE can be understood to be composed of
+In abstract terms, Plants in \ac{CLEAVE} can be understood to be composed of
 \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
@@
     \caption{
-        Example configuration file for a CLEAVE Plant, defining the parameters for a single emulation.
+        Example configuration file for a \ac{CLEAVE} Plant, defining the parameters for a single emulation.
         ``Plant emulation parameters'' define key emulation parameters, such as the emulation discrete-time update rate (the \emph{tick rate}).

### 2022-01-04 intro.tex
editing
changes

@@
 
+\todo[inline, caption={}]{
+    \begin{itemize}
+        \item You miss a bit the balance between general introduction (really only the first paragraph), discussion of related work (almost one page) and the contribution of the paper (again one paragraph).
+Try to find a better balance, and the best way to do that is to shorten perhaps the discussion of related work.
+        \item The aspect of edge computing is a bit missing in the intro.
+That is perhaps not a big problem, but you could introduce it more strongly towards the contribution part.
+One useful angle could be that we see more and more concepts of the ``edge'' --- near edge, far edge, telco edge etc.
+While edge computing is seen as important for closed loop applications in general, understanding the usefullness of such different edge concepts will become more important --- which is where Cleave can help obviously.
+        \item In the intro (first paragraph) you can be a bit more specific as 5G together with edge computing is introduced to become the backbone of CPS in the future.
+As of today, not many CPS really run on cellular networks.
+That does not come out from your intro, and might require you to change the narrative of the first paragraph
+    \end{itemize}
+}
+
 The number and applications of \acp{CPS}~\cite{Rajkumar2010CPS} --- that is, systems in which a real, physical mechanism is controlled by a computer --- have exploded in later years.
-This rise in popularity is in great part driven by the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge Computing.
+This rise in popularity is in great part driven by the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing.
 
@@
 
-The most straightforward approach to experimental research in \ac{NCS} use setups in which the complete system uses real hardware.
+The most straightforward approach to experimental research in \ac{NCS} uses setups in which the complete system is built on top of real hardware.
 \textcite{Drew2005NCSWLAN} present \iac{NCS} model which considers both random packet delay and loss on both the sensor and actuator sides of the feedback loop; \textcite{Baumann2018LowPower}, on the other hand, develop an evaluation methodology for wireless \ac{NCS}.
@@
 
-\textcite{Zoppi2020NCSBench} made a first attempt at tackling this challenge with their \emph{NCSbench} platform.
+\textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge with their \emph{NCSbench} platform.
 \emph{NCSbench} is an open-source \ac{NCS} benchmarking platform designed with reproducibility in mind, built using the \citetitle{LEGOMindstormsEV3}~\cite{LEGOMindstormsEV3} platform.
@@
 This framework is built on top of the \emph{Python 3.8}~\cite{Python3.8} language, making it highly extensible and able to harness the tremendous amount of already existing user-provided libraries and packages~\cite{pypi}.
-Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}~\cite{merkel2014docker}, making it suitable for automated deployment and benchmarking on industry-standard deployment using container orchestration technologies.
+Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}~\cite{merkel2014docker}, making it suitable for automated deployment and benchmarking on industry-standard setups using container orchestration technologies.
 We name this framework \ac{CLEAVE}.

### 2022-01-04 intro.tex
fixing intro

@@
 The number and applications of \acp{CPS}~\cite{Rajkumar2010CPS} --- that is, systems in which a real, physical mechanism is controlled by a computer --- have exploded in later years.
-This rise in popularity is in great part driven by the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing.
+However, this rapid increase in adoption of these systems has been mostly limited to industrial contexts.
+Although \acp{CPS} present huge opportunities for all facets of society, they have yet to reach our daily lives in any relevant scale due to their stringent operational requirements.
+This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing~\cite{Satya2009Case}, comsumer-grade \acp{CPS} will be made possible.
+These technologies meet two key requirements of \acp{CPS}: real-time capabilities (through extremely low end-to-end latencies), and context- and locality-awareness, and will most likely become the backbone of \ac{CPS} in the future.
 
-In particular, \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system, have been of great interest to the scientific community in the last decade.
-Due to their potential advantages for industrial settings~\cite{Lu2015WSAN}, a large body of work exists dedicated to the modelling and performance characterization of these systems~\cite{Hespanha2007Survey,Zhang2013Survey,Zhang2016Survey}.
+A subcategory of \acp{CPS} which stands to particularly benefit from the adoption of the above technologies is \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system.
+Depending on the physical system being controlled, \acp{NCS} can have extremely stringent timing and reliability requirements for the communication between plant and controller; requirements that conventional cloud computing and cellular networks are unable to meet~\cite{Liu2017Review,Wan2020Efficient}.
+
+% These systems have been of great interest to the scientific community in the last decade.
+Due to their potential advantages for industrial and commercial settings~\cite{Lu2015WSAN}, a large body of work exists dedicated to the modelling and performance characterization of \aclp{NCS}~\cite{Hespanha2007Survey,Zhang2013Survey,Zhang2016Survey}.
 \Acp{NCS} enhance the flexibility of and improve existing control systems by allowing for the distribution of control functions over and across networks.
@@
 
-As evidence by the discussion above, experimental research in \acp{NCS} includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
+As evidenced by the discussion above, experimental research in \acp{NCS} includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
 This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
@@
 
-In this paper, we present thus the first fully-software-based \ac{NCS} benchmarking framework with scalability and repeatability as its main focus.
-This framework is built on top of the \emph{Python 3.8}~\cite{Python3.8} language, making it highly extensible and able to harness the tremendous amount of already existing user-provided libraries and packages~\cite{pypi}.
-Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}~\cite{merkel2014docker}, making it suitable for automated deployment and benchmarking on industry-standard setups using container orchestration technologies.
-We name this framework \ac{CLEAVE}.
+In this paper, we present thus the first fully-software-based \ac{NCS} benchmarking framework with scalability and repeatability as its main focus, built with the Edge in mind.
+In the years since the advent of the Edge computing paradigm, more and more variations of it have appeared in literature as the paradigm starts to be adopted by industry.
+``Near'', ``far'', ``core'', and ``telco'' edge, among others, are all terms which describe variations of the original concept and which are becoming ubiquitous in new research.
+While the main idea of Edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the usefulness, strenghts, and weaknesses of such different edge concepts is becoming more and more important.
+
+Our framework, which we name \ac{CLEAVE}, thus aims to simplify the repeatable and scalable benchmarking of such systems. 
+It is built on top of the \emph{Python 3.8}~\cite{Python3.8} language, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries and packages~\cite{pypi}.
+Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}~\cite{merkel2014docker}, making it suitable for automated deployment and benchmarking on industry-standard cloud and (more importantly) edge setups using container orchestration technologies.
 

### 2022-01-04 intro.tex
more edits

@@
     \begin{itemize}
-        \item You miss a bit the balance between general introduction (really only the first paragraph), discussion of related work (almost one page) and the contribution of the paper (again one paragraph).
-Try to find a better balance, and the best way to do that is to shorten perhaps the discussion of related work.
-        \item The aspect of edge computing is a bit missing in the intro.
-That is perhaps not a big problem, but you could introduce it more strongly towards the contribution part.
-One useful angle could be that we see more and more concepts of the ``edge'' --- near edge, far edge, telco edge etc.
-While edge computing is seen as important for closed loop applications in general, understanding the usefullness of such different edge concepts will become more important --- which is where Cleave can help obviously.
-        \item In the intro (first paragraph) you can be a bit more specific as 5G together with edge computing is introduced to become the backbone of CPS in the future.
-As of today, not many CPS really run on cellular networks.
-That does not come out from your intro, and might require you to change the narrative of the first paragraph
+        \item Still missing balance, and too long. Will fix towards the end.
     \end{itemize}
@@
 Although \acp{CPS} present huge opportunities for all facets of society, they have yet to reach our daily lives in any relevant scale due to their stringent operational requirements.
-This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing~\cite{Satya2009Case}, comsumer-grade \acp{CPS} will be made possible.
+This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing~\cite{Satya2009Case,Satya2017Emergence}, comsumer-grade \acp{CPS} will be made possible.
 These technologies meet two key requirements of \acp{CPS}: real-time capabilities (through extremely low end-to-end latencies), and context- and locality-awareness, and will most likely become the backbone of \ac{CPS} in the future.
@@
 
-The most straightforward approach to experimental research in \ac{NCS} uses setups in which the complete system is built on top of real hardware.
+The most straightforward approach to experimental research in \acp{NCS} uses setups in which the complete system is built on top of real hardware.
 \textcite{Drew2005NCSWLAN} present \iac{NCS} model which considers both random packet delay and loss on both the sensor and actuator sides of the feedback loop; \textcite{Baumann2018LowPower}, on the other hand, develop an evaluation methodology for wireless \ac{NCS}.
@@
 On the other hand, network-in-the-loop refers to experimental setups in which the \emph{network} is instead simulated or emulated.
-An example of this approach can be found in \textcite{Natale2004InvPendEthernet}, in which the authors interface an real inverted pendulum system with an emulated IEEE 802.3 Ethernet network to study the effects of the \ac{CSMA/CD} medium access control scheme on plant stability.
+An example of this approach can be found in \textcite{Natale2004InvPendEthernet}, in which the authors interface a real inverted pendulum system with an emulated IEEE 802.3 Ethernet network to study the effects of the \ac{CSMA/CD} medium access control scheme on plant stability.
 
-As evidenced by the discussion above, experimental research in \acp{NCS} includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
+As evidenced above, experimental research in \acp{NCS} includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
 This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
@@
 
-In this paper, we present thus the first fully-software-based \ac{NCS} benchmarking framework with scalability and repeatability as its main focus, built with the Edge in mind.
+In this paper, we present the first fully-software-based \ac{NCS} benchmarking framework with scalability and repeatability as its main focus, built with the Edge in mind.
 In the years since the advent of the Edge computing paradigm, more and more variations of it have appeared in literature as the paradigm starts to be adopted by industry.
@@
 
-Our framework, which we name \ac{CLEAVE}, thus aims to simplify the repeatable and scalable benchmarking of such systems. 
+Our framework, which we name \ac{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems. 
 It is built on top of the \emph{Python 3.8}~\cite{Python3.8} language, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries and packages~\cite{pypi}.

### 2022-01-04 experiments.tex
more writing

@@
 In this section, we aim to validate the utility of the CLEAVE framework through a series of experiments.
+In particular, we wish to show the utility of CLEAVE as a tool for repeatable and scalable benchmarking of \acp{NCS} in Edge environments.
 
@@
 
-We depict our experimental setup in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
-We employ a Raspberry Pi 4B single-board computer together with a general-purpose \verb|x86_64|-based workstation.
-These devices are connected to each other by an 802.11n WiFi link.
-For ease of deployment and reparametrization, we deploy our CLEAVE control loop as a pair of Docker~\cite{merkel2014docker} containers\todo{add link to image?}; one for the plant, running on the Rasberry Pi device, and one for the Controller, running on the workstation.
-These instances are then connected to an overlay network configured through Docker which sits on top of and abstracts away the real network configuration, and allows for transparent communication between the Plant and Controller.
+This section is structured in two parts.
+\Cref{ssec:expsetup} details the experimental setup and describes the experiments performed.
+\Cref{ssec:results} then presents and discusses the numerical results.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=.95\columnwidth]{images/CLEAVE_experiment_setup}
-    \caption{Experimental setup.
-Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Overlay Network spanning a Raspberry Pi 4B Plant host connected to an x86 Controller host over a 802.11n WiFi link.}\label{fig:cleave:expsetup}
-\end{figure}
+\subsection{Experimental Setup}\label{ssec:expsetup}
 
-We use a two-dimensional inverted pendulum system for the control system deployed inside CLEAVE for the experiments, 
-This system was chosen due to its relative simplicity and prevalence in the field of automatic control as one of the fundamental examples of linear control.
-The physical system, represented in \cref{fig:invpend}, is implemented using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}.
+The \acl{NCS} employed for these experiments corresponds to an inverted pendulum, chosen for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
+The physical system, represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}, executed at a constant \SI{120}{\hertz}.
 For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the NumPy numeric computation library~\autocite{harris2020array}.
-
-This setup is then used to run a series of experiments with varying parametrization of the controlled system.
-Specifically, we modify:
-\begin{itemize}
-    \item the sampling rate of the Plant state, setting it to \SIlist[list-units=single,list-final-separator={, or }]{5;10;20;40;60;120}{\hertz};
-    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-units=single,list-final-separator={, or }]{0;25;50;100;200}{\milli\second} after the processing of each sample.
-\end{itemize}
-Each one of the \num{30} possible combinations of these parameters is run on the experimental testbed depicted in \cref{fig:cleave:expsetup} for \num{10} repetitions lasting \SI{5}{\minute} each, yielding \SI{25}{\hour} worth of data.
-During each repetition, we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
 
@@
 \end{figure}
+
+We depict our experimental setup in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
+We employ a Raspberry Pi 4B single-board computer together with a general-purpose \verb|x86_64|-based workstation; the \ac{OS} configured on both corresponds to Ubuntu Server 20.04 LTS~\cite{Ubuntu}.
+These devices are connected to each other by an 802.11n WiFi link.
+For ease of orchestration and re-parametrization, as well as to mimic real-world deployments, we package our CLEAVE control loop as a pair of Docker~\cite{merkel2014docker} containers.
+For our main set of experiments, a two-node Docker \emph{Swarm}~\cite{Swarm2021} is set up with the \verb|x86_64| host configured as a Manager and the Raspberry Pi as a Worker node.
+The containers are deployed inside the Swarm, with one of them configured as a Controller, assigned to the \verb|x86_64| host, and the other one as the Inverted Pendulum Plant and assigned to the Raspberry Pi.
+These instances communicate over an overlay network which sits on top of and abstracts away the real network configuration.
+Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which we simply execute both Plant and Controller containers co-located on the \verb|x86_64| host.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.95\columnwidth]{images/CLEAVE_experiment_setup}
+    \caption{Experimental setup.
+Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Swarm Overlay Network spanning a Raspberry Pi 4B Plant host connected to an x86 Controller host over a 802.11n WiFi link.}\label{fig:cleave:expsetup}
+\end{figure}
+
+These setups are then used to run a series of experiments with varying parametrization of the controlled system.
+Specifically, we modify:
+\begin{itemize}
+    \item the sampling rate of the Plant state, setting it to \SIlist[list-units=single,list-final-separator={, or }]{5;10;20;40;60}{\hertz};
+    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-units=single,list-final-separator={, or }]{0;25;50;100}{\milli\second} after the processing of each sample.
+\end{itemize}
+We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with interesting and relevant results are then repeated an additional \num{20} times for better statistical significance in the results.
+During each repetition, we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
+
+\subsection{Results}\label{ssec:results}

### 2022-01-05 experiments.tex
writing

@@
 
-\todo[inline]{WIP}
+\input{tables/hardware.tex}
 
-In this section, we aim to validate the utility of the CLEAVE framework through a series of experiments.
-In particular, we wish to show the utility of CLEAVE as a tool for repeatable and scalable benchmarking of \acp{NCS} in Edge environments.
-
-\todo[inline]{talk about goal of experiments}
+In this section, we demonstrate the practical utility of \ac{CLEAVE} through a series of experiments on an emulated \ac{NCS} running between a client device and a cloudlet.
+We aim to answer question relating to the ability of our framework to provide relevant, accurate, and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
 
@@
 
-We depict our experimental setup in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
-We employ a Raspberry Pi 4B single-board computer together with a general-purpose \verb|x86_64|-based workstation; the \ac{OS} configured on both corresponds to Ubuntu Server 20.04 LTS~\cite{Ubuntu}.
+We depict our experimental edge setup in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
+We employ a Raspberry Pi 4B single-board computer as a client device, together with a general-purpose \verb|x86_64|-based workstation which acts as the cloudlet.
 These devices are connected to each other by an 802.11n WiFi link.
 For ease of orchestration and re-parametrization, as well as to mimic real-world deployments, we package our CLEAVE control loop as a pair of Docker~\cite{merkel2014docker} containers.
-For our main set of experiments, a two-node Docker \emph{Swarm}~\cite{Swarm2021} is set up with the \verb|x86_64| host configured as a Manager and the Raspberry Pi as a Worker node.
-The containers are deployed inside the Swarm, with one of them configured as a Controller, assigned to the \verb|x86_64| host, and the other one as the Inverted Pendulum Plant and assigned to the Raspberry Pi.
-These instances communicate over an overlay network which sits on top of and abstracts away the real network configuration.
-Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which we simply execute both Plant and Controller containers co-located on the \verb|x86_64| host.
+For our main set of experiments, a two-node Docker \emph{Swarm}~\cite{Swarm2021} is set up with the cloudlet host configured as a Manager and the client device as a Worker node.
+The containers are deployed inside the Swarm, with one of them configured as a Controller, assigned to the cloudlet, and the other one as the Inverted Pendulum Plant and assigned to the client.
+These instances communicate using the \ac{UDP} over an overlay network which sits on top of and abstracts away the real network configuration.
+Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which we simply execute both Plant and Controller containers co-located on the cloudlet.
 
@@
 \subsection{Results}\label{ssec:results}
+
+The results presented below provide valuable insights on both system limits and on the chosen \ac{NCS} itself, as well as on the capabilities of \ac{CLEAVE}.
+
+We begin with a simple analysis of the stability of the Plant.
+\Cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
+It provides a very clear overview of the minimum requirements
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.8\columnwidth]{topple_fraction.png}
+    \caption{Fraction of plants that toppled, per experimental setup.}%
+    \label{fig:topple:fraction}
+\end{figure}
+
+When it comes to network reliability, mean packet losses for all setups were below \SI{0.15}{\percent}. 
+\todo[inline]{Discuss which network results to add.}

### 2022-01-06 experiments.tex
writing

@@
 
-We begin with a simple analysis of the stability of the Plant.
-\Cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
-It provides a very clear overview of the minimum requirements
+We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
+For instance, \cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
+These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value will depend on the parametrization of the controller, and in this case corresponds to \SI{20}{\degree}, or \SI{0.35}{\radian}} were marked as ``toppled''.
 
@@
 
+\Ac{CLEAVE} also allows for more fine-grained analysis of stability.
+\Cref{fig:topple:rms} shows the \ac{RMS} of the angles for each setup, averaged over all repetitions of each experiment.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.8\columnwidth]{angle_rms.png}
+    \caption{RMS}%
+    \label{fig:topple:rms}
+\end{figure}
+
 When it comes to network reliability, mean packet losses for all setups were below \SI{0.15}{\percent}. 

### 2022-01-06 experiments.tex
writing

@@
 
+\todo[inline]{Capacity experiment description.}
+
 \subsection{Results}\label{ssec:results}
@@
 
+\subsubsection{Single Plant Scenario}
+
 We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
 For instance, \cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
-These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value will depend on the parametrization of the controller, and in this case corresponds to \SI{20}{\degree}, or \SI{0.35}{\radian}} were marked as ``toppled''.
+These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value will depend on the parametrization of the controller, and in this case corresponds to \SI{20}{\degree}, or \SI{0.35}{\radian}.} were marked as ``toppled''.
 
@@
 \Cref{fig:topple:rms} shows the \ac{RMS} of the angles for each setup, averaged over all repetitions of each experiment.
+This figure shows an intuitive but important result --- less stable setups show higher amplitude in the oscillations of their penduli.
 
@@
     \includegraphics[width=.8\columnwidth]{angle_rms.png}
-    \caption{RMS}%
+    \caption{\Ac{RMS} of the pendulum angles averaged over all runs for each scenario.
+    Error bars indicate \SI{95}{\percent} \acp{CI}.}%
     \label{fig:topple:rms}
@@
 
-When it comes to network reliability, mean packet losses for all setups were below \SI{0.15}{\percent}. 
+Finally, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
+\Cref{fig:network} shows average \acp{RTT}s and packet losses per scenario.
+
+\begin{figure}
+    \centering
+    % \includegraphics[width=.8\columnwidth]{angle_rms.png}
+    \caption[caption]{Mean \acp{RTT} and packet losses for each scenario; \acp{RTT} were calculated by subtracting the synthetic delay from the absolute end-to-end latency\textsuperscript{a}.
+Error bars indicate \SI{95}{\percent} \acp{CI}.}%
+    \small\textsuperscript{a} The systematic offset of \SI{2}{\milli\second} on scenarios with synthetic delay is an artifact of our implementation, and will be fixed in the next release of the framework.
+    \label{fig:network}
+\end{figure}
+
 \todo[inline]{Discuss which network results to add.}

### 2022-01-07 experiments.tex
adding results

@@
 The \acl{NCS} employed for these experiments corresponds to an inverted pendulum, chosen for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
-The physical system, represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}, executed at a constant \SI{120}{\hertz}.
+The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}, executed at a constant \SI{120}{\hertz}.
 For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the NumPy numeric computation library~\autocite{harris2020array}.
+The plant and controller are then packaged into Docker\cite{merkel2014docker}, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
 
@@
 
-We depict our experimental edge setup in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
-We employ a Raspberry Pi 4B single-board computer as a client device, together with a general-purpose \verb|x86_64|-based workstation which acts as the cloudlet.
-These devices are connected to each other by an 802.11n WiFi link.
-For ease of orchestration and re-parametrization, as well as to mimic real-world deployments, we package our CLEAVE control loop as a pair of Docker~\cite{merkel2014docker} containers.
-For our main set of experiments, a two-node Docker \emph{Swarm}~\cite{Swarm2021} is set up with the cloudlet host configured as a Manager and the client device as a Worker node.
-The containers are deployed inside the Swarm, with one of them configured as a Controller, assigned to the cloudlet, and the other one as the Inverted Pendulum Plant and assigned to the client.
-These instances communicate using the \ac{UDP} over an overlay network which sits on top of and abstracts away the real network configuration.
-Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which we simply execute both Plant and Controller containers co-located on the cloudlet.
+We depict our experimental edge setup in \cref{fig:cleave:expsetup}.
+It consists of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose Cloudlet.
+On top of this physical architecture we configure a Docker \emph{Swarm}~\cite{merkel2014docker,Swarm2021} managed centrally from the Cloudlet.
+For each experimental scenario we deploy a number of control loops inside this Swarm, executing plant containers exclusively in the Raspberry Pi clients (up to a single plant per client) and controller containers in the Cloudlet.
+Plants and controllers communicate using the \ac{UDP} over an overlay network which sits on top of and abstracts away the real network configuration.
+Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which plants and controllers are executed co-located on the Cloudlet.
 
@@
     \caption{Experimental setup.
-Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Swarm Overlay Network spanning a Raspberry Pi 4B Plant host connected to an x86 Controller host over a 802.11n WiFi link.}\label{fig:cleave:expsetup}
+Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Swarm Overlay Network spanning \num{10} Raspberry Pi 4B clients connected to a single Cloudlet over an IEEE 802.11n \ac{AP}.}\label{fig:cleave:expsetup}
+    \todo[inline]{Change fig.}
 \end{figure}
 
-These setups are then used to run a series of experiments with varying parametrization of the controlled system.
-Specifically, we modify:
+\subsubsection{Single Loop Baseline Scenarios}
+
+We first run a series of single loop scenarios with varying parametrization of the controlled system.
+\todo[inline]{Purpose of these scenarios?}
+We modify:
 \begin{itemize}
-    \item the sampling rate of the Plant state, setting it to \SIlist[list-units=single,list-final-separator={, or }]{5;10;20;40;60}{\hertz};
-    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-units=single,list-final-separator={, or }]{0;25;50;100}{\milli\second} after the processing of each sample.
+    \item the sampling rate of the Plant state, setting it to \SIlist[list-final-separator={, or }]{5;10;20;40;60}{\hertz};
+    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-final-separator={, or }]{0;25;50;100}{\milli\second} after the processing of each sample.
 \end{itemize}
 We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with interesting and relevant results are then repeated an additional \num{20} times for better statistical significance in the results.
-During each repetition, we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
+Each repetition lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
 
-\todo[inline]{Capacity experiment description.}
+\subsubsection{Realistic Scenario with Resource Contention}
+
+Next, we run a more complex scenario to validate the utility of \ac{CLEAVE} in a more realistic setting where network resources are shared with video stream traffic.
+Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
+
+In this scenario, we deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
+On the remaining \num{4} clients we run the \emph{iperf3}~\cite{iperf3} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of \ac{UDP} uplink traffic.
+This traffic emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
+We execute this scenario with \ac{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
+Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
 
@@
 \todo[inline]{Discuss which network results to add.}
+
+\subsubsection{Realistic Scenario}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[h]{\columnwidth}
+        \centering
+        \includegraphics[width=.8\columnwidth]{video_topple_frac}
+        \caption{}\label{fig:video:toppled}
+    \end{subfigure}\\%
+    \begin{subfigure}[h]{\columnwidth}
+        \centering
+        \includegraphics[width=.8\columnwidth]{video_angle_rms}
+        \caption{}\label{fig:video:rms}
+    \end{subfigure}%
+    \caption{
+        Stability metrics for the realistic scenario.
+        \labelcref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
+        \labelcref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
+        Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
+    }\label{fig:video:stability}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[h]{\columnwidth}
+        \centering
+        \includegraphics[width=.8\columnwidth]{video_drop_frac}
+        \caption{}\label{fig:video:drop}
+    \end{subfigure}\\%
+    \begin{subfigure}[h]{\columnwidth}
+        \centering
+        \includegraphics[width=.8\columnwidth]{video_rtt}
+        \caption{}\label{fig:video:rtt}
+    \end{subfigure}%
+    \caption{
+        Network metrics for the realistic scenario.
+        \labelcref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
+        \labelcref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
+        Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
+    }\label{fig:video:network}
+\end{figure}

### 2022-01-08 conclusion.tex
results

\section{Conclusion}\label{sec:conclusion}
\todo[inline]{TODO}

### 2022-01-08 experiments.tex
results

@@
 The \acl{NCS} employed for these experiments corresponds to an inverted pendulum, chosen for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
-The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}, executed at a constant \SI{120}{\hertz}.
-For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the NumPy numeric computation library~\autocite{harris2020array}.
-The plant and controller are then packaged into Docker\cite{merkel2014docker}, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
+The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk 2d physics library: \url{http://www.pymunk.org/en/latest/}}, executed at a constant \SI{120}{\hertz}.
+For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
+The plant and controller are then packaged into Docker\footnote{Docker Engine: \url{https://www.docker.com/}} containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
 
@@
     \centering
-    \includegraphics[width=.95\columnwidth]{images/inverted_pendulum.png}
+    \includegraphics[width=.7\columnwidth]{images/inverted_pendulum.png}
     \caption{
@@
 It consists of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose Cloudlet.
-On top of this physical architecture we configure a Docker \emph{Swarm}~\cite{merkel2014docker,Swarm2021} managed centrally from the Cloudlet.
+On top of this physical architecture we configure a Docker Swarm\footnote{Swarm mode overview: \url{https://docs.docker.com/engine/swarm/}} managed centrally from the Cloudlet.
 For each experimental scenario we deploy a number of control loops inside this Swarm, executing plant containers exclusively in the Raspberry Pi clients (up to a single plant per client) and controller containers in the Cloudlet.
@@
 Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Swarm Overlay Network spanning \num{10} Raspberry Pi 4B clients connected to a single Cloudlet over an IEEE 802.11n \ac{AP}.}\label{fig:cleave:expsetup}
-    \todo[inline]{Change fig.}
 \end{figure}
@@
 We first run a series of single loop scenarios with varying parametrization of the controlled system.
-\todo[inline]{Purpose of these scenarios?}
-We modify:
+These scenarios are intended as initial baselines to evaluate the accuracy of the \ac{CLEAVE} framework and showcase its flexibility.
+
+For these scenarios, we vary:
 \begin{itemize}
@@
 \end{itemize}
+
 We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with interesting and relevant results are then repeated an additional \num{20} times for better statistical significance in the results.
@@
 
-\subsubsection{Realistic Scenario with Resource Contention}
+Scenarios are executed automatically in batches using a simple Python script which interacts with Docker through the widely adopted \emph{docker-py}\footnote{Docker SDK for Python: \url{https://docker-py.readthedocs.io/en/stable/}} library.
+This is \ac{CLEAVE}'s first advantage over existing frameworks; it is designed with cloud and edge technology and paradigms in mind, making integration with existing systems straightforward.
 
-Next, we run a more complex scenario to validate the utility of \ac{CLEAVE} in a more realistic setting where network resources are shared with video stream traffic.
+\subsubsection{Realistic Scenario with Network Resource Contention}
+
+Next, we run a multi-loop scenario to validate the utility of \ac{CLEAVE} in a more realistic setting where network resources are shared with video stream traffic.
 Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
@@
 In this scenario, we deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
-On the remaining \num{4} clients we run the \emph{iperf3}~\cite{iperf3} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of \ac{UDP} uplink traffic.
-This traffic emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
+On the remaining \num{4} clients we run the \emph{iperf3}\footnote{iperf3: \url{https://iperf.fr/}} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of uplink \ac{UDP} traffic.
+This emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
 We execute this scenario with \ac{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
 Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
+Once again, repetitions of this scenario are executed automatically in batches using a simple Python script and \emph{docker-py}.
 
@@
 We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
-For instance, \cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
+\Cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
 These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value will depend on the parametrization of the controller, and in this case corresponds to \SI{20}{\degree}, or \SI{0.35}{\radian}.} were marked as ``toppled''.
+
+Next, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
+\Cref{fig:single:network} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios, and packet losses for the networked single-loop scenario\footnote{Packet losses in the local scenario were, of course, \num{0}.}.
+
+From these results we can already infer interesting consequences for \ac{NCS} deployments on the edge.
+For instance, it is clear from \cref{fig:topple:fraction} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
+A corollary of this is that, conversely, at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
+Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.
 
@@
     \centering
-    \includegraphics[width=.8\columnwidth]{topple_fraction.png}
+    \includegraphics[width=.8\columnwidth]{single_loop_toppled}
     \caption{Fraction of plants that toppled, per experimental setup.}%
@@
 
-\Ac{CLEAVE} also allows for more fine-grained analysis of stability.
-\Cref{fig:topple:rms} shows the \ac{RMS} of the angles for each setup, averaged over all repetitions of each experiment.
-This figure shows an intuitive but important result --- less stable setups show higher amplitude in the oscillations of their penduli.
-
 \begin{figure}
     \centering
-    \includegraphics[width=.8\columnwidth]{angle_rms.png}
-    \caption{\Ac{RMS} of the pendulum angles averaged over all runs for each scenario.
+    \begin{subfigure}[h]{\columnwidth}
+        \centering
+        \includegraphics[width=.8\columnwidth]{single_loop_rtts}
+        \caption{}\label{fig:single:rtt}
+    \end{subfigure}\\%
+    \begin{subfigure}[h]{\columnwidth}
+        \centering
+        \includegraphics[width=.8\columnwidth]{single_loop_drop_frac}
+        \caption{}\label{fig:single:drop}
+    \end{subfigure}%
+    \caption[caption]{Mean latency due to network and processing (\labelcref{fig:single:rtt}) for both single-loop scenarios, and packet losses (\labelcref{fig:single:drop}) for the single-loop scenario over the WiFi link.
     Error bars indicate \SI{95}{\percent} \acp{CI}.}%
-    \label{fig:topple:rms}
+    % \vspace{2pt}\raggedright{}\small\textsuperscript{a} The systematic offset of approx. \SI{2}{\milli\second} on scenarios with synthetic delay is due to an artifact of our implementation.
+    \label{fig:single:network}
 \end{figure}
 
-Finally, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
-\Cref{fig:network} shows average \acp{RTT}s and packet losses per scenario.
+\subsubsection{Realistic Scenario}
 
-\begin{figure}
-    \centering
-    % \includegraphics[width=.8\columnwidth]{angle_rms.png}
-    \caption[caption]{Mean \acp{RTT} and packet losses for each scenario; \acp{RTT} were calculated by subtracting the synthetic delay from the absolute end-to-end latency\textsuperscript{a}.
-Error bars indicate \SI{95}{\percent} \acp{CI}.}%
-    \small\textsuperscript{a} The systematic offset of \SI{2}{\milli\second} on scenarios with synthetic delay is an artifact of our implementation, and will be fixed in the next release of the framework.
-    \label{fig:network}
-\end{figure}
+In the following, we present and briefly discuss the results for the realistic scenario with network resource contention.
 
-\todo[inline]{Discuss which network results to add.}
+% As before, we start with a brief and simple analysis of plant stability.
+\Cref{fig:video:stability} shows plots representing the fraction of repetitions with toppled penduli as well as the \ac{RMS} value of the angle of those penduli that did not topple.
+\Cref{fig:video:network} shows network metrics for the scenario, with plots showing the amount of dropped \ac{UDP} datagrams and the average \acp{RTT}.
 
-\subsubsection{Realistic Scenario}
+These results are interesting in their counter-intuitiveness.
+Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:stability} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
+\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
+\SI{40}{\hertz} was in turn the second worst configuration --- although no penduli toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.
+
+We can see an explanation for these behaviors in \cref{fig:video:network}.
+Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} datagrams lost.
+The differences in \acp{RTT} results are equally dramatic; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approximately \num{3} times those for the \SIlist{20;40}{\hertz} setups (circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).
+
+These results stem from the contention for shared network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acl{NCS} for deployment on the edge.
+The edge is envisioned as \emph{multi-tenant} and \emph{multi-instance}, and thus any edge \ac{NCS} deployment will have to be performed with shared resources in mind.
+\acp{NCS} will have to dynamically adapt to system limits in order to maintain quality-of-control.
 

### 2022-01-08 cleave.tex
figures

@@
 
-In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}.
+\todo[inline]{
+    For the approach part, mention please also that it is open-source and where it can be found.
+Also, you keep the discussion around the implemented plant and the controller pretty short (and it is only showing up in the numerical part).
+Maybe it is better to discuss the details of the implemented plant and the controller already in Section 2 and mention that the intention is to build a liberary of plants and controllers in the future.
+}
+
+In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}, with a particular focus on edge deployments.
 Our design follows an approach similar to the above-discussed \emph{hardware-in-the-loop} setups, in that we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
-This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of \acp{NCS}, namely the network itself.
+This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of and edge \acp{NCS}, namely the network itself.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{images/CLEAVE_NCS_structure.png}
+    \includegraphics[width=.8\textwidth]{CLEAVE_NCS_structure}
     \caption{
-        Structure of an emulated \ac{NCS} on top of CLEAVE.
+        Structure of an emulated \ac{NCS} on top of CLEAVE.\@
         The State object implements the discrete-time evolution of the Plant.

### 2022-01-08 conclusion.tex
figures

@@
 \section{Conclusion}\label{sec:conclusion}
-\todo[inline]{TODO}
+
+Performance characterization and evaluation of \acl{NCS} is hard, due to the tight interactions and cross-dependencies between network resource utilization and control-loop stability which complicate the scaling and repeatability of benchmarks.
+This issue has been largely glossed over in \ac{NCS} literature; existing research is mostly focused on \emph{theory}, and the small fraction of experimental research studies available tend to implement \emph{ad-hoc} solutions.
+To the best of our knowledge, \textcite{Zoppi2020NCSBench}'s work is the only available framework for \ac{NCS} benchmarking with repeatability and reproducibility in mind.
+
+In this work, we aim to tackle this issue by presenting a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking.
+Our framework, \ac{CLEAVE}, is fully parametrizable and built with industry-standard edge technologies and paradigms in mind.
+We validate the utility of this tool through a series of scenarios, from which we are able to extract relevant metrics relating to the stability of the control system, as well
+
+% Benchmarking human-in-the-loop applications is hard, given
+% their tight interaction with human users who complicate the scaling
+% and repeatability of experiments. In this paper, we have presented
+% a benchmarking suite for this type of applications, called Edge-
+% Droid 1.0, capable of cutting out the need for users in performance
+% evaluations. We achieve this by employing pre-recorded sensory
+% input traces which we play over the network to the real applica-
+% tion backend, employing a parameterized user model to react to
+% feedback. We demonstrate its utility through a series of use case
+% scenarios, from which we are able to extract metrics regarding la-
+% tency both in regards to the application itself and the hardware
+% stack. We believe the EdgeDroid 1.0 suite thus represents an impor-
+% tant rst step towards enabling inexpensive and low-complexity
+% large-scale research on the scaling limits of this type of applications,
+% a requirement for wide adoption of the technology.
+
+% Nonetheless, there is still future work to be done.
+% The user model presented in this paper is only preliminary, and
+% we are currently conducting research in characterizing user behav-
+% ior when interacting with WCA applications in order to present
+% a more complete model in the future. As mentioned in Section 4,
+% in EdgeDroid 1.0, our model is that of a user who does not suer
+% any of the shortcomings of real human users such as annoyance, fa-
+% tigue, frustration, nausea. Rather, EdgeDroid 1.0 models a perfectly
+% stoic user who is like an automaton and responds in a precisely
+% reproducible and deterministic manner to the same system stimu-
+% lus every time. Of course, no real human user is an automaton. In
+% the future, we envision creating many versions of EdgeDroid (i.e.,
+% EdgeDroid 2.0, EdgeDroid 3.0, etc.) that embody more human-like
+% user models that more accurately emulate attributes such as those
+% mentioned above. Experimental validation of these human-like user
+% models via user studies will be an important part of our future work.
+% We are also working on expanding the benchmarking suite to
+% also work rst with other types of Wearable Cognitive Assistance,
+% and later with other categories of human-in-the-loop applications.
+% Other types of WCA we will consider in future iterations of the tool
+% include real-time task-assistance WCA applications (such as the
+% Ping-Pong application described in [7]), which don’t have a linear
+% task model like task-guidance WCA and have tighter latency bounds
+% and context- and information-providing WCA applications, for
+% instance, applications which recognize faces and provide relevant
+% social-media information related to that person. The latter also do
+% not have a linear task model, but present more lax latency bounds.

### 2022-01-08 experiments.tex
figures

@@
 \section{Experimental validation}\label{sec:experiments}
+
+\todo[inline]{
+    In the numerical part, you can gain a lot of space by reformatting the plots into environments with two or three plots next to each other.
+In addition, you need to comment more on Fig 4 (the key result for the single-loop baseline).
+The discussion of Fig 4 should take more space than the discussion of Fig 5.
+}
 
@@
 The \acl{NCS} employed for these experiments corresponds to an inverted pendulum, chosen for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
-The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk 2d physics library: \url{http://www.pymunk.org/en/latest/}}, executed at a constant \SI{120}{\hertz}.
+The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
 For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
@@
 \Cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
-These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value will depend on the parametrization of the controller, and in this case corresponds to \SI{20}{\degree}, or \SI{0.35}{\radian}.} were marked as ``toppled''.
-
+These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
 Next, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
-\Cref{fig:single:network} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios, and packet losses for the networked single-loop scenario\footnote{Packet losses in the local scenario were, of course, \num{0}.}.
+\Cref{fig:single:network} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios, and packet losses for the networked single-loop scenario\footnote{Packet losses in the local scenario were, of course, \SI{0}{\percent}.}.
 
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics[width=.8\columnwidth]{single_loop_toppled}
-    \caption{Fraction of plants that toppled, per experimental setup.}%
+    \begin{subfigure}[t]{.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{single_loop_toppled}
+        \caption{}%
         \label{fig:topple:fraction}
-\end{figure}
+    \end{subfigure}%
+    \begin{subfigure}[t]{.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{single_loop_rms}
+        \caption{}\label{fig:single:rms}
+    \end{subfigure}%
+    \caption[caption]{
+        Stability metrics for the single-loop scenarios.
+        \labelcref{fig:topple:fraction} shows the fraction of plants that toppled, per experimental setup.
+        \labelcref{fig:single:rms} shows the \ac{RMS} value of the angle for all penduli that did not topple in the single loop scenarios.
+        Error bars indicate \SI{95}{\percent} \acp{CI}.
+        }%
+    \label{fig:single:network}
+    \todo[inline]{More focus on this.}
+\end{figure*}
 
@@
     \centering
-    \begin{subfigure}[h]{\columnwidth}
-        \centering
-        \includegraphics[width=.8\columnwidth]{single_loop_rtts}
-        \caption{}\label{fig:single:rtt}
-    \end{subfigure}\\%
-    \begin{subfigure}[h]{\columnwidth}
-        \centering
-        \includegraphics[width=.8\columnwidth]{single_loop_drop_frac}
-        \caption{}\label{fig:single:drop}
-    \end{subfigure}%
-    \caption[caption]{Mean latency due to network and processing (\labelcref{fig:single:rtt}) for both single-loop scenarios, and packet losses (\labelcref{fig:single:drop}) for the single-loop scenario over the WiFi link.
-    Error bars indicate \SI{95}{\percent} \acp{CI}.}%
-    % \vspace{2pt}\raggedright{}\small\textsuperscript{a} The systematic offset of approx. \SI{2}{\milli\second} on scenarios with synthetic delay is due to an artifact of our implementation.
-    \label{fig:single:network}
+    \includegraphics[width=.9\columnwidth]{single_loop_rtts}
+    \caption{
+        Mean latency due to network and processing for both single-loop scenarios.
+        Error bars indicate \SI{95}{\percent} \acp{CI}.
+    }\label{fig:single:rtt}
+    \todo[inline]{less focus on this. mention packet drops are all below 0.2 percent}
 \end{figure}
@@
 
-% As before, we start with a brief and simple analysis of plant stability.
-\Cref{fig:video:stability} shows plots representing the fraction of repetitions with toppled penduli as well as the \ac{RMS} value of the angle of those penduli that did not topple.
-\Cref{fig:video:network} shows network metrics for the scenario, with plots showing the amount of dropped \ac{UDP} datagrams and the average \acp{RTT}.
+\Cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
+These results are interesting in their counter-intuitiveness.
+Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
 
-These results are interesting in their counter-intuitiveness.
-Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:stability} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
 \SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
@@
 
-We can see an explanation for these behaviors in \cref{fig:video:network}.
+We can see an explanation for these behaviors in \cref{fig:video:drop,fig:video:rtt}.
 Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} datagrams lost.
@@
 
-\begin{figure}
+\begin{figure*}[t]
     \centering
-    \begin{subfigure}[h]{\columnwidth}
+    \begin{subfigure}[h]{.25\textwidth}
         \centering
-        \includegraphics[width=.8\columnwidth]{video_topple_frac}
+        \includegraphics[width=.9\textwidth]{video_topple_frac}
         \caption{}\label{fig:video:toppled}
-    \end{subfigure}\\%
-    \begin{subfigure}[h]{\columnwidth}
+    \end{subfigure}%
+    \begin{subfigure}[h]{.25\textwidth}
         \centering
-        \includegraphics[width=.8\columnwidth]{video_angle_rms}
+        \includegraphics[width=.9\textwidth]{video_angle_rms}
         \caption{}\label{fig:video:rms}
     \end{subfigure}%
-    \caption{
-        Stability metrics for the realistic scenario.
-        \labelcref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
-        \labelcref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
-        Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
-    }\label{fig:video:stability}
-\end{figure}
-
-\begin{figure}
+    \begin{subfigure}[h]{.25\textwidth}
         \centering
-    \begin{subfigure}[h]{\columnwidth}
+        \includegraphics[width=.9\textwidth]{video_drop_frac}
+        \caption{}\label{fig:video:drop}
+    \end{subfigure}%
+    \begin{subfigure}[h]{.25\textwidth}
         \centering
-        \includegraphics[width=.8\columnwidth]{video_drop_frac}
-        \caption{}\label{fig:video:drop}
-    \end{subfigure}\\%
-    \begin{subfigure}[h]{\columnwidth}
-        \centering
-        \includegraphics[width=.8\columnwidth]{video_rtt}
+        \includegraphics[width=.9\textwidth]{video_rtt}
         \caption{}\label{fig:video:rtt}
@@
     \caption{
-        Network metrics for the realistic scenario.
+        Results for the realistic scenario.
+        \labelcref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
+        \labelcref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
         \labelcref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
         \labelcref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
-        Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
-    }\label{fig:video:network}
-\end{figure}
+        Error bars indicate \SI{95}{\percent} \ac{CI} in all plots.
+    }\label{fig:video:results}
+\end{figure*}
+
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}[h]{\columnwidth}
+%         \centering
+%         \includegraphics[width=.8\columnwidth]{video_drop_frac}
+%         \caption{}\label{fig:video:drop}
+%     \end{subfigure}\\%
+%     \begin{subfigure}[h]{\columnwidth}
+%         \centering
+%         \includegraphics[width=.8\columnwidth]{video_rtt}
+%         \caption{}\label{fig:video:rtt}
+%     \end{subfigure}%
+%     \caption{
+%         Network metrics for the realistic scenario.
+%         Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
+%     }\label{fig:video:network}
+% \end{figure}

### 2022-01-08 intro.tex
figures

@@
         \item Still missing balance, and too long. Will fix towards the end.
+        \item try to shorten the discussion around NCS.
+It is fine that you introduce it, but you don't need the extensive discussion.
+Likewise, for the experimental NCS works and the simulation works, you can package all that into a single paragraph, where your intent it simply to point out how people typically work in this domain.
+Then discuss the downsides, and lead over to mreo scalable and repeatable frameworks, i.e., NCSbench.
+Keep that discussion short and work out more strongly that it still heavily relies on the physical plant hardware (you just speak of hardware) being operated.
+The rest of the intro is fine.
     \end{itemize}
+
+    CHANGE: removed 5Hz and 100ms cases from plots.
 }

### 2022-01-08 experiments.tex
figures

@@
 
-\subsubsection{Single Loop Baseline Scenarios}
-
-We first run a series of single loop scenarios with varying parametrization of the controlled system.
-These scenarios are intended as initial baselines to evaluate the accuracy of the \ac{CLEAVE} framework and showcase its flexibility.
-
-For these scenarios, we vary:
-\begin{itemize}
-    \item the sampling rate of the Plant state, setting it to \SIlist[list-final-separator={, or }]{5;10;20;40;60}{\hertz};
-    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-final-separator={, or }]{0;25;50;100}{\milli\second} after the processing of each sample.
-\end{itemize}
-
-We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with interesting and relevant results are then repeated an additional \num{20} times for better statistical significance in the results.
-Each repetition lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
-
-Scenarios are executed automatically in batches using a simple Python script which interacts with Docker through the widely adopted \emph{docker-py}\footnote{Docker SDK for Python: \url{https://docker-py.readthedocs.io/en/stable/}} library.
-This is \ac{CLEAVE}'s first advantage over existing frameworks; it is designed with cloud and edge technology and paradigms in mind, making integration with existing systems straightforward.
-
-\subsubsection{Realistic Scenario with Network Resource Contention}
-
-Next, we run a multi-loop scenario to validate the utility of \ac{CLEAVE} in a more realistic setting where network resources are shared with video stream traffic.
-Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
-
-In this scenario, we deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
-On the remaining \num{4} clients we run the \emph{iperf3}\footnote{iperf3: \url{https://iperf.fr/}} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of uplink \ac{UDP} traffic.
-This emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
-We execute this scenario with \ac{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
-Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
-Once again, repetitions of this scenario are executed automatically in batches using a simple Python script and \emph{docker-py}.
-
-\subsection{Results}\label{ssec:results}
-
-The results presented below provide valuable insights on both system limits and on the chosen \ac{NCS} itself, as well as on the capabilities of \ac{CLEAVE}.
-
-\subsubsection{Single Plant Scenario}
-
-We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
-\Cref{fig:topple:fraction} shows a visualization of the fraction of plants that toppled in each scenario.
-These Plants were identified simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
-Next, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
-\Cref{fig:single:network} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios, and packet losses for the networked single-loop scenario\footnote{Packet losses in the local scenario were, of course, \SI{0}{\percent}.}.
-
-From these results we can already infer interesting consequences for \ac{NCS} deployments on the edge.
-For instance, it is clear from \cref{fig:topple:fraction} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
-A corollary of this is that, conversely, at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
-Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.
-
-\begin{figure*}
+\begin{figure*}[h]
     \centering
@@
         \caption{}%
-        \label{fig:topple:fraction}
+        \label{fig:single:topple}
     \end{subfigure}%
@@
         Stability metrics for the single-loop scenarios.
-        \labelcref{fig:topple:fraction} shows the fraction of plants that toppled, per experimental setup.
+        \labelcref{fig:single:topple} shows the fraction of plants that toppled, per experimental setup.
         \labelcref{fig:single:rms} shows the \ac{RMS} value of the angle for all penduli that did not topple in the single loop scenarios.
@@
         }%
-    \label{fig:single:network}
+    \label{fig:single:stability}
     \todo[inline]{More focus on this.}
@@
 \end{figure}
-
-\subsubsection{Realistic Scenario}
-
-In the following, we present and briefly discuss the results for the realistic scenario with network resource contention.
-
-\Cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
-These results are interesting in their counter-intuitiveness.
-Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
-
-\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
-\SI{40}{\hertz} was in turn the second worst configuration --- although no penduli toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.
-
-We can see an explanation for these behaviors in \cref{fig:video:drop,fig:video:rtt}.
-Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} datagrams lost.
-The differences in \acp{RTT} results are equally dramatic; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approximately \num{3} times those for the \SIlist{20;40}{\hertz} setups (circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).
-
-These results stem from the contention for shared network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acl{NCS} for deployment on the edge.
-The edge is envisioned as \emph{multi-tenant} and \emph{multi-instance}, and thus any edge \ac{NCS} deployment will have to be performed with shared resources in mind.
-\acp{NCS} will have to dynamically adapt to system limits in order to maintain quality-of-control.
 
@@
 
-% \begin{figure}
-%     \centering
-%     \begin{subfigure}[h]{\columnwidth}
-%         \centering
-%         \includegraphics[width=.8\columnwidth]{video_drop_frac}
-%         \caption{}\label{fig:video:drop}
-%     \end{subfigure}\\%
-%     \begin{subfigure}[h]{\columnwidth}
-%         \centering
-%         \includegraphics[width=.8\columnwidth]{video_rtt}
-%         \caption{}\label{fig:video:rtt}
-%     \end{subfigure}%
-%     \caption{
-%         Network metrics for the realistic scenario.
-%         Error bars indicate \SI{95}{\percent} \ac{CI} in both plots.
-%     }\label{fig:video:network}
-% \end{figure}
+\subsubsection{Single Loop Baseline Scenarios}
+
+We first run a series of single loop scenarios with varying parametrization of the controlled system.
+These scenarios are intended as initial baselines to evaluate the accuracy of the \ac{CLEAVE} framework and showcase its flexibility.
+
+For these scenarios, we vary:
+\begin{itemize}
+    \item the sampling rate of the Plant state, setting it to \SIlist[list-final-separator={, or }]{10;20;40;60}{\hertz};
+    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-final-separator={, or }]{0;25;50}{\milli\second} after the processing of each sample.
+\end{itemize}
+
+We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with interesting and relevant results are then repeated an additional \num{20} times for better statistical significance in the results.
+Each repetition lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
+
+Scenarios are executed automatically in batches using a simple Python script which interacts with Docker through the widely adopted \emph{docker-py}\footnote{Docker SDK for Python: \url{https://docker-py.readthedocs.io/en/stable/}} library.
+This is \ac{CLEAVE}'s first advantage over existing frameworks; it is designed with cloud and edge technology and paradigms in mind, making integration with existing systems straightforward.
+
+\subsubsection{Realistic Scenario with Network Resource Contention}
+
+Next, we run a multi-loop scenario to validate the utility of \ac{CLEAVE} in a more realistic setting where network resources are shared with video stream traffic.
+Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
+
+In this scenario, we deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
+On the remaining \num{4} clients we run the \emph{iperf3}\footnote{iperf3: \url{https://iperf.fr/}} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of uplink \ac{UDP} traffic.
+This emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
+We execute this scenario with \ac{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
+Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
+Once again, repetitions of this scenario are executed automatically in batches using a simple Python script and \emph{docker-py}.
+
+\subsection{Results}\label{ssec:results}
+
+The results presented below provide valuable insights on both system limits and on the chosen \ac{NCS} itself, as well as on the capabilities of \ac{CLEAVE}.
+
+\subsubsection{Single Plant Scenario}
+
+We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
+\Cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
+\cref{fig:single:topple} shows a visualization of the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} amplitude for the penduli angles.
+
+We identify toppled plants simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
+
+Next, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
+\Cref{fig:single:rtt} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios.
+Packet losses were below \SI{0.2}{\percent} for all parametrizations of the single-loop scenario over the network, and \SI{0}{\percent} for all parametrizations of the local-only setup.
+
+From these results we can already infer interesting consequences for \ac{NCS} deployments on the edge.
+For instance, it is clear from \cref{fig:single:stability} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
+A corollary of this is that, conversely, at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
+Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.
+
+\subsubsection{Realistic Scenario}
+
+In the following, we present and briefly discuss the results for the realistic scenario with network resource contention.
+
+\Cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
+These results are interesting in their counter-intuitiveness.
+Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
+
+\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
+\SI{40}{\hertz} was in turn the second worst configuration --- although no penduli toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.
+
+We can see an explanation for these behaviors in \cref{fig:video:drop,fig:video:rtt}.
+Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} datagrams lost.
+The differences in \acp{RTT} results are equally dramatic; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approximately \num{3} times those for the \SIlist{20;40}{\hertz} setups (circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).
+
+These results stem from the contention for shared network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acl{NCS} for deployment on the edge.
+The edge is envisioned as \emph{multi-tenant} and \emph{multi-instance}, and thus any edge \ac{NCS} deployment will have to be performed with shared resources in mind.
+\acp{NCS} will have to dynamically adapt to system limits in order to maintain quality-of-control.

### 2022-01-08 cleave.tex
fixing

@@
 
-\todo[inline]{
-    For the approach part, mention please also that it is open-source and where it can be found.
-Also, you keep the discussion around the implemented plant and the controller pretty short (and it is only showing up in the numerical part).
-Maybe it is better to discuss the details of the implemented plant and the controller already in Section 2 and mention that the intention is to build a liberary of plants and controllers in the future.
-}
-
 In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}, with a particular focus on edge deployments.
+The framework is distributed as \emph{free and open-source} software, and can be found online in the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
 Our design follows an approach similar to the above-discussed \emph{hardware-in-the-loop} setups, in that we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
@@
 
+\subsection{Implemented Control Loops}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.7\columnwidth]{images/inverted_pendulum.png}
+    \caption{
+        The two-dimensional inverted pendulum system.
+        The cart moves on the X-axis, and the pendulum on top of it swings freely.
+        The objective of the system is to balance the pendulum vertically through the application of horizontal forces on the cart.
+    }\label{fig:invpend}
+\end{figure}
+
+For the purpose of this work, we have used \ac{CLEAVE} to implement an emulated \emph{inverted pendulum} control loop; see \cref{fig:invpend}.
+We have chosen this system as an initial benchmark for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
+However, we eventually plan to build an open library of emulated control loops to be able to provide more thorough and varied benchmarks.
+
+The inverted pendulum Plant is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
+For the Controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
+The plant and controller are then packaged into Docker\footnote{Docker Engine: \url{https://www.docker.com/}} containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
+
 \begin{listing}

### 2022-01-08 experiments.tex
fixing

@@
 \section{Experimental validation}\label{sec:experiments}
-
-\todo[inline]{
-    In the numerical part, you can gain a lot of space by reformatting the plots into environments with two or three plots next to each other.
-In addition, you need to comment more on Fig 4 (the key result for the single-loop baseline).
-The discussion of Fig 4 should take more space than the discussion of Fig 5.
-}
 
@@
 \subsection{Experimental Setup}\label{ssec:expsetup}
-
-The \acl{NCS} employed for these experiments corresponds to an inverted pendulum, chosen for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
-The physical system (``plant''), represented in \cref{fig:invpend}, is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
-For the controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
-The plant and controller are then packaged into Docker\footnote{Docker Engine: \url{https://www.docker.com/}} containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=.7\columnwidth]{images/inverted_pendulum.png}
-    \caption{
-        The two-dimensional inverted pendulum system.
-        The cart moves on the X-axis, and the pendulum on top of it swings freely.
-        The objective of the system is to balance the pendulum vertically through the application of horizontal forces on the cart.
-    }\label{fig:invpend}
-\end{figure}
 
@@
     \label{fig:single:stability}
-    \todo[inline]{More focus on this.}
 \end{figure*}
@@
     }\label{fig:single:rtt}
-    \todo[inline]{less focus on this. mention packet drops are all below 0.2 percent}
 \end{figure}
@@
 
-We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable analysis of relevant metrics.  
+We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable collection and posterior processing of relevant metrics.  
 \Cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
 \cref{fig:single:topple} shows a visualization of the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} amplitude for the penduli angles.
-
 We identify toppled plants simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
 
-Next, we showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
-\Cref{fig:single:rtt} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios.
-Packet losses were below \SI{0.2}{\percent} for all parametrizations of the single-loop scenario over the network, and \SI{0}{\percent} for all parametrizations of the local-only setup.
-
-From these results we can already infer interesting consequences for \ac{NCS} deployments on the edge.
+As expected, higher sampling rates directly correlate to better quality of control (with the exception of the \SIlist{40;60}{\hertz}, \SI{50}{\milli\second} scenarios --- we attribute this to randomness and the relatively short duration of the experiments).
+As sampling rates increased, the system was able to reach stability at higher \ac{RTT} latencies.
+These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.
 For instance, it is clear from \cref{fig:single:stability} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
@@
 Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.
+
+For completeness, we also showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
+\Cref{fig:single:rtt} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios.
+Packet losses were below \SI{0.2}{\percent} for all parametrizations of the single-loop scenario over the network, and \SI{0}{\percent} for all parametrizations of the local-only setup.
 
@@
 Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
-
 \SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.

### 2022-01-08 intro.tex
shortened intro

--- 2022-01-04 intro.tex
@@
 \section{Introduction}\label{sec:intro}
-
-\todo[inline, caption={}]{
-    \begin{itemize}
-        \item Still missing balance, and too long. Will fix towards the end.
-    \end{itemize}
-}
 
@@
 
-The most straightforward approach to experimental research in \acp{NCS} uses setups in which the complete system is built on top of real hardware.
-\textcite{Drew2005NCSWLAN} present \iac{NCS} model which considers both random packet delay and loss on both the sensor and actuator sides of the feedback loop; \textcite{Baumann2018LowPower}, on the other hand, develop an evaluation methodology for wireless \ac{NCS}.
-Both works validate their results on physical testbeds, consisting of inverted penduli plants controlled wirelessly over IEEE 802.11 WiFi from a central computation point.
-Similarly, \textcite{Li2014Wireless} implement a wireless microcontroller-based system for vibration control and validate it on a physical prototype consisting of a cantilever beam controlled over an IEEE 802.15.4 \ac{WPAN} system.
-\textcite{Cuenca2019UAV} design and implement periodic event-triggered sampling and dual-rate control techniques for wireless control, validating their contributions on a four-rotor autonomous drone platform controlled over WiFi.
-
-Conversely, some studies choose to instead use completely \emph{simulated} \ac{NCS} setups~\cite{Andersson2005Simulation,Eyisi2012NCSWT}.
-Such an approach is employed by \textcite{Du2009Smith}; the authors develop a novel \emph{Smith} predictor to compensate for varying latencies in wireless \acp{NCS} and validate it using the \emph{TrueTime}~\cite{Henriksson2002TrueTime} simulator.
-In \textcite{Chen2015synccontrol} the authors validate and evaluate a number of synchronous control strategies for three-motor setups using completely simulated \acp{NCS} environments.
-\citeauthor{Wu2012NPC} introduce the concept of \emph{network predictive control} in\ \cite{Wu2012NPC} and use it to balance a simulated inverted pendulum over a simulated wireless network.
-In \cite{Ma2019DynamicSched}, \citeauthor{Ma2019DynamicSched} propose an optimal dynamic scheduling strategy that optimizes performance of multi-loop control systems and validate it by simulating a four-loop control system over an IEEE 802.15.4 \ac{WPAN}.
-
+One approach to experimental research in \acp{NCS} uses setups in which the complete system is built on top of real hardware.
+This approach is employed in the works of, for instance, \textcite{Drew2005NCSWLAN,Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement and validate their results on physical testbeds of different kinds.
+Conversely, some studies choose to instead use completely \emph{simulated} \ac{NCS} setups.
+\textcite{Du2009Smith,Wu2012NPC,Chen2015synccontrol,Ma2019DynamicSched} are examples of works in which the authors have opted for such completely virtualized approaches.
+These studies often employ combinations of physical and network simulation libraries to try to capture the complex dynamics of \acp{NCS}.
 Finally, a number of experimental studies instead employ \emph{hardware-} and \emph{network-in-the-loop} approaches.
-Hardware-in-the-loop refers to \ac{NCS} setups in which a real network interacts with a simulated or emulated control system.
-This kind of setup is particularly prevalent in the field of \emph{smart grid} control, and has been used by studies such as \textcite{Wang2020VoltageControl}, in which the authors validate a novel three-level coordinated control method for photovoltaic inverters.
-
-On the other hand, network-in-the-loop refers to experimental setups in which the \emph{network} is instead simulated or emulated.
-An example of this approach can be found in \textcite{Natale2004InvPendEthernet}, in which the authors interface a real inverted pendulum system with an emulated IEEE 802.3 Ethernet network to study the effects of the \ac{CSMA/CD} medium access control scheme on plant stability.
+Hardware-in-the-loop refers to \ac{NCS} setups in which a real network interacts with a simulated or emulated control system~\cite{Wang2020VoltageControl}.
+On the other hand, network-in-the-loop refers to experimental setups in which the \emph{network} is instead simulated or emulated~\cite{Natale2004InvPendEthernet}.
 
@@
 Furthermore, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
-We identify thus a gap in knowledge pertaining to the reproducibility and comparison of experimental studies on these systems.
+This has caused a gap in knowledge pertaining to the reproducibility and comparison of experimental studies on these systems.
 
@@
 Although the work of \textcite{Zoppi2020NCSBench} is a relevant and important step towards the reproducibility of \ac{NCS} benchmarks, we believe it still falls short on some aspects.
-In particular, we argue the reliance of \emph{NCSbench} on specific hardware --- accessible as it may be --- hampers the potential of the platform, as it fundamentally limits both the physical systems that can be benchmarked and the scalability of the performed benchmarks.
+In particular, we argue the reliance of \emph{NCSbench} on specific hardware for the Plant hampers the potential of the platform.
+The LEGO Mindstorms platform, although accessible and flexible, still fundamentally limits the physical systems that can be benchmarked to the kind of systems that can be physically implemented on it.
+Furthermore, it is \emph{not} scalable, and thus unsuitable for studies of larger, distributed \acp{NCS}.
 

### 2022-01-08 acks.tex
conclusion and title

\section*{Acknowledgements}\label{sec:acks}
\todo[inline]{TODO: James? Funding? Project course students?}

### 2022-01-08 conclusion.tex
fixing
conclusion and title

@@
 
-Performance characterization and evaluation of \acl{NCS} is hard, due to the tight interactions and cross-dependencies between network resource utilization and control-loop stability which complicate the scaling and repeatability of benchmarks.
-This issue has been largely glossed over in \ac{NCS} literature; existing research is mostly focused on \emph{theory}, and the small fraction of experimental research studies available tend to implement \emph{ad-hoc} solutions.
+The issue of repeatable and scalable benchmarks has been largely glossed over in \ac{NCS} literature; existing research is mostly focused on \emph{theory}, and the small fraction of experimental research studies available tend to implement \emph{ad-hoc} solutions.
 To the best of our knowledge, \textcite{Zoppi2020NCSBench}'s work is the only available framework for \ac{NCS} benchmarking with repeatability and reproducibility in mind.
 
-In this work, we aim to tackle this issue by presenting a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking.
+In this work, we aim to tackle this issue by presenting a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking with a particular focus on edge deployment.
 Our framework, \ac{CLEAVE}, is fully parametrizable and built with industry-standard edge technologies and paradigms in mind.
-We validate the utility of this tool through a series of scenarios, from which we are able to extract relevant metrics relating to the stability of the control system, as well
+We validate the utility of this tool through a series of scenarios, from which we are able to extract relevant metrics relating to the stability of the control system, as well as on the perfomance of the underlying network link.
+We believe \ac{CLEAVE} represents an impor-
+tant step towards enabling inexpensive and low-complexity
+scalable research for real-world deployment of edge-bound \acp{NCS}.
+
+There is still however work to be done.
+To start with, the current implementation of \ac{CLEAVE} only includes a single plant-controller pair.
+We envision creating an open library of plants and controllers to share with the community.
+
+\ac{CLEAVE}'s integration with industry-standard tools and frameworks such as Docker will be refined.
+At the moment, these interactions are still quite superficial.
+Our goal is to achieve a much tighter integration, for instance by providing the complete toolkit as pre-packaged container images.
+
+Finally, the validity of the results obtained by the framework will have to be verified through more thorough and realistic scenarios than what we have been able to show in this work.
+In particular, we intend to eventually perform large-scale experimentation targetting 5G cellular deployments, as this technology is set to become the backbone of edge networks in the near future.
 

### 2022-01-08 abstract.tex
writing
working
abstract

@@
 \begin{abstract}
-    \todo[inline]{TODO}
+    \acfp*{NCS} have long been a subject of theoretical study in academia, however, little work has been devoted to repeatable and scalable experimentation.
+    With the advent of edge computing and the subsequent proliferation of \acp{NCS} this is quickly becoming a key issue that needs to be addressed.
+
+    In this work, we present a novel, completely software-based approach to this challenge.
+    Our core idea rests on a \emph{hardware-in-the-loop} approach in which we emulate a physical plant which communicates over a real network with a controller implemented in software.
+    Our approach is furthermore designed and built for the Edge, using a general-purpose programming language and will full compatibility with industry-standard containerization solutions.
 \end{abstract}

### 2022-01-08 cleave.tex
fixes vishnu

@@
 
-In the following we will explain the approach taken in \ac{CLEAVE} by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation on top of the framework.
+In the following, we will explain the approach taken in \ac{CLEAVE} by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation on top of the framework.
 An overview of the structure of such an emulation on \ac{CLEAVE} can be seen in \cref{fig:cleave:ncs:struct}.
@@
 \texttt{Actuator} objects process the outputs of the Controller and update the values of the \texttt{State} accordingly.
-Conversely \texttt{Sensor} objects sample relevant values from the \texttt{State} object, process them, and finally pass them on to the Controller through the network.
+Conversely, \texttt{Sensor} objects sample relevant values from the \texttt{State} object, process them, and finally pass them on to the Controller through the network.
 

### 2022-01-08 experiments.tex
fixes vishnu

@@
 In this section, we demonstrate the practical utility of \ac{CLEAVE} through a series of experiments on an emulated \ac{NCS} running between a client device and a cloudlet.
-We aim to answer question relating to the ability of our framework to provide relevant, accurate, and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
+We aim to answer questions relating to the ability of our framework to provide relevant, accurate, and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
 
@@
         \labelcref{fig:single:topple} shows the fraction of plants that toppled, per experimental setup.
-        \labelcref{fig:single:rms} shows the \ac{RMS} value of the angle for all penduli that did not topple in the single loop scenarios.
+        \labelcref{fig:single:rms} shows the \ac{RMS} value of the angle for all pendula that did not topple in the single loop scenarios.
         Error bars indicate \SI{95}{\percent} \acp{CI}.
@@
 \Cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
-\cref{fig:single:topple} shows a visualization of the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} amplitude for the penduli angles.
+\cref{fig:single:topple} shows a visualization of the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} amplitude for the pendula angles.
 We identify toppled plants simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
@@
 \SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
-\SI{40}{\hertz} was in turn the second worst configuration --- although no penduli toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.
+\SI{40}{\hertz} was in turn the second worst configuration --- although no pendula toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.
 
 We can see an explanation for these behaviors in \cref{fig:video:drop,fig:video:rtt}.
-Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} datagrams lost.
+Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} of datagrams lost.
 The differences in \acp{RTT} results are equally dramatic; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approximately \num{3} times those for the \SIlist{20;40}{\hertz} setups (circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).
@@
 The edge is envisioned as \emph{multi-tenant} and \emph{multi-instance}, and thus any edge \ac{NCS} deployment will have to be performed with shared resources in mind.
-\acp{NCS} will have to dynamically adapt to system limits in order to maintain quality-of-control.
+\acp{NCS} will have to dynamically adapt to system limits in order to maintain quality of control.

### 2022-01-08 intro.tex
fixes vishnu

@@
 Although \acp{CPS} present huge opportunities for all facets of society, they have yet to reach our daily lives in any relevant scale due to their stringent operational requirements.
-This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing~\cite{Satya2009Case,Satya2017Emergence}, comsumer-grade \acp{CPS} will be made possible.
+This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing~\cite{Satya2009Case,Satya2017Emergence}, consumer-grade \acp{CPS} will be made possible.
 These technologies meet two key requirements of \acp{CPS}: real-time capabilities (through extremely low end-to-end latencies), and context- and locality-awareness, and will most likely become the backbone of \ac{CPS} in the future.

### 2022-01-09 abstract.tex
extend abstract

@@
     Our core idea rests on a \emph{hardware-in-the-loop} approach in which we emulate a physical plant which communicates over a real network with a controller implemented in software.
-    Our approach is furthermore designed and built for the Edge, using a general-purpose programming language and will full compatibility with industry-standard containerization solutions.
+    This approach is furthermore designed and built for the Edge, using a general-purpose programming language and will full compatibility with industry-standard containerization solutions.
+    Our results showcase the utility of this framework as a repeatable and scalable solution to \ac{NCS} performance evaluation.
 \end{abstract}

### 2022-01-09 cleave.tex
andresa fixes

@@
 For the purpose of this work, we have used \ac{CLEAVE} to implement an emulated \emph{inverted pendulum} control loop; see \cref{fig:invpend}.
-We have chosen this system as an initial benchmark for to its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
+We have chosen this system as an initial benchmark for its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
 However, we eventually plan to build an open library of emulated control loops to be able to provide more thorough and varied benchmarks.

### 2022-01-09 conclusion.tex
andresa fixes

@@
 The issue of repeatable and scalable benchmarks has been largely glossed over in \ac{NCS} literature; existing research is mostly focused on \emph{theory}, and the small fraction of experimental research studies available tend to implement \emph{ad-hoc} solutions.
-To the best of our knowledge, \textcite{Zoppi2020NCSBench}'s work is the only available framework for \ac{NCS} benchmarking with repeatability and reproducibility in mind.
+To the best of our knowledge, the work by \textcite{Zoppi2020NCSBench} is the only available framework for \ac{NCS} benchmarking focused on repeatability and reproducibility.
 
@@
 We validate the utility of this tool through a series of scenarios, from which we are able to extract relevant metrics relating to the stability of the control system, as well as on the perfomance of the underlying network link.
-We believe \ac{CLEAVE} represents an impor-
-tant step towards enabling inexpensive and low-complexity
-scalable research for real-world deployment of edge-bound \acp{NCS}.
+\todo[inline]{Talk a bit more about results?}
+We believe \ac{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \acp{NCS}.
 
-There is still however work to be done.
+There is still, however, work to be done.
 To start with, the current implementation of \ac{CLEAVE} only includes a single plant-controller pair.
 We envision creating an open library of plants and controllers to share with the community.
-
 \ac{CLEAVE}'s integration with industry-standard tools and frameworks such as Docker will be refined.
@@
 Our goal is to achieve a much tighter integration, for instance by providing the complete toolkit as pre-packaged container images.
-
 Finally, the validity of the results obtained by the framework will have to be verified through more thorough and realistic scenarios than what we have been able to show in this work.

### 2022-01-09 experiments.tex
andresa fixes

@@
 We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with interesting and relevant results are then repeated an additional \num{20} times for better statistical significance in the results.
+\todo[inline]{Explain criteria for repeated experiments? Or just drop this sentence?}
 Each repetition lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
@@
 We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable collection and posterior processing of relevant metrics.  
-\Cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
+\cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
 \cref{fig:single:topple} shows a visualization of the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} amplitude for the pendula angles.
@@
 For completeness, we also showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
-\Cref{fig:single:rtt} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios.
+\cref{fig:single:rtt} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios.
 Packet losses were below \SI{0.2}{\percent} for all parametrizations of the single-loop scenario over the network, and \SI{0}{\percent} for all parametrizations of the local-only setup.
@@
 
-\Cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
+\cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
 These results are interesting in their counter-intuitiveness.

### 2022-01-09 intro.tex
andresa fixes

@@
 
-The number and applications of \acp{CPS}~\cite{Rajkumar2010CPS} --- that is, systems in which a real, physical mechanism is controlled by a computer --- have exploded in later years.
+The number and applications of \acp{CPS}~\cite{Rajkumar2010CPS} --- that is, systems in which a real, physical mechanism is controlled by a computer --- have exploded in recent years.
 However, this rapid increase in adoption of these systems has been mostly limited to industrial contexts.
@@
 \Acp{NCS} enhance the flexibility of and improve existing control systems by allowing for the distribution of control functions over and across networks.
-This allows for --- for instance --- simultaneous centralized coordination, control, and monitoring of potentially hundreds of devices.
+This allows for, for instance, simultaneous centralized coordination, control, and monitoring of potentially hundreds of devices.
 

### 2022-01-09 cleave.tex
trimming

@@
 
-In the following, we will explain the approach taken in \ac{CLEAVE} by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation on top of the framework.
-An overview of the structure of such an emulation on \ac{CLEAVE} can be seen in \cref{fig:cleave:ncs:struct}.
+Below, we will explain the approach taken in \ac{CLEAVE} by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation.
+An overview of the structure of such a deployment can be seen in \cref{fig:cleave:ncs:struct}.
 
@@
         The State object implements the discrete-time evolution of the Plant.
-        For this purpose, it holds two sets of special variables; Actuated Variables and Sensed Variables.
-        Actuated Variables are updated from the values obtained from the Actuator objects at the beginning of each time step.
-        These are used to perform a time step update in the State, and update the values of the Sensed Variables.
-        The values of the Sensed Variables are then, at the end of each time step, sampled and passed on to the Sensor objects, which in turn process them (for instance, to add noise) and send the processed values to the Controller.
-        The Controller uses these values to calculate the new actuation commands and sends these to the Actuators.
-        Finally, the Actuators process the actuation commands and hold the new values for the Actuated Variables until they are read at the next time step.
+        Actuated Variables are updated from the values obtained from the Actuator objects at the beginning of each time step and are used to update the State before each time-step.
+        Sensed Variables are sampled at the end of each time-step and passed on to the Sensor objects, which process them (e.g. to add noise) and send the processed values to the Controller.
+        Controller calculates actuation values and sends these to the Actuators.
+        Finally, the Actuators process the actuation values and hold the results until they are read at the next time-step.
     }\label{fig:cleave:ncs:struct}
@@
 When invoked, it is passed a mapping of property names to sampled values.
-The framework also expects implementations of this method to return yet another mapping, this time of actuated property names to target values.
+Implementations of this method should return yet another mapping of actuated property names to target values.
 Users are otherwise free to implement any kind of logic or process within extending classes.
@@
         The two-dimensional inverted pendulum system.
-        The cart moves on the X-axis, and the pendulum on top of it swings freely.
-        The objective of the system is to balance the pendulum vertically through the application of horizontal forces on the cart.
+        The objective of the system is to balance the vertically free-swinging pendulum by moving the cart.
     }\label{fig:invpend}
@@
 We have chosen this system as an initial benchmark for its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
-However, we eventually plan to build an open library of emulated control loops to be able to provide more thorough and varied benchmarks.
+However, we eventually plan to build an open library of emulated control loops.
 
@@
 For the Controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
-The plant and controller are then packaged into Docker\footnote{Docker Engine: \url{https://www.docker.com/}} containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
+The plant and controller are then packaged into Docker containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
 
@@
         Example configuration file for a \ac{CLEAVE} Plant, defining the parameters for a single emulation.
-        ``Plant emulation parameters'' define key emulation parameters, such as the emulation discrete-time update rate (the \emph{tick rate}).
-        The Sensors and Actuators are then provided in lists, each Sensor and Actuator attached to a specific property identified by a string.
-        Additionally, Sensors have a sampling rate in Hz which needs to be provided as an integer.
+        % ``Plant emulation parameters'' define key emulation parameters, such as the emulation discrete-time update rate (the \emph{tick rate}).
+        % The Sensors and Actuators are then provided in lists, each Sensor and Actuator attached to a specific property identified by a string.
+        % Additionally, Sensors have a sampling rate in Hz which needs to be provided as an integer.
     }

### 2022-01-09 experiments.tex
trimming

@@
 
-In this section, we demonstrate the practical utility of \ac{CLEAVE} through a series of experiments on an emulated \ac{NCS} running between a client device and a cloudlet.
-We aim to answer questions relating to the ability of our framework to provide relevant, accurate, and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
+In this section, we demonstrate the utility of \ac{CLEAVE} through a series of experiments on an emulated \ac{NCS} running on a number of client devices and a cloudlet.
+We aim to answer questions relating to the ability of our framework to provide accurate and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
 
@@
     \includegraphics[width=.95\columnwidth]{images/CLEAVE_experiment_setup}
-    \caption{Experimental setup.
-Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Swarm Overlay Network spanning \num{10} Raspberry Pi 4B clients connected to a single Cloudlet over an IEEE 802.11n \ac{AP}.}\label{fig:cleave:expsetup}
+    \caption{
+        The setup used for our experimentation. 
+        % Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Swarm Overlay Network spanning \num{10} Raspberry Pi 4B clients connected to a single Cloudlet over an IEEE 802.11n \ac{AP}.
+    }\label{fig:cleave:expsetup}
 \end{figure}
@@
 
-We first run a series of single loop scenarios with varying parametrization of the controlled system.
-These scenarios are intended as initial baselines to evaluate the accuracy of the \ac{CLEAVE} framework and showcase its flexibility.
+We first run a series of single loop scenarios with varying parametrization of the \ac{NCS}, intended as initial baselines to evaluate the accuracy of the \ac{CLEAVE} framework and showcase its flexibility.
 
@@
     \item the sampling rate of the Plant state, setting it to \SIlist[list-final-separator={, or }]{10;20;40;60}{\hertz};
-    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-final-separator={, or }]{0;25;50}{\milli\second} after the processing of each sample.
+    \item the responsiveness of the Controller, adding fixed delays of \SIlist[list-final-separator={, or }]{0;25;50}{\milli\second} after the processing of each sample.
 \end{itemize}
@@
 \cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
-\cref{fig:single:topple} shows a visualization of the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} amplitude for the pendula angles.
-We identify toppled plants simply by analyzing the emulation data --- instances whose pendulum angles reached values above a certain value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
+\cref{fig:single:topple} shows the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} for the pendula angles.
+Toppled plants were identified simply by analysis of the emulation data; instances whose pendulum angles reached values above a threshold\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
 
@@
 
-These results stem from the contention for shared network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acl{NCS} for deployment on the edge.
-The edge is envisioned as \emph{multi-tenant} and \emph{multi-instance}, and thus any edge \ac{NCS} deployment will have to be performed with shared resources in mind.
+These results stem from the contention for network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acl{NCS} for deployment on the edge.
+The edge will be \emph{multi-tenant} and \emph{multi-instance}, and thus any edge \ac{NCS} deployment will have to be designed with shared resources in mind.
 \acp{NCS} will have to dynamically adapt to system limits in order to maintain quality of control.

### 2022-01-09 intro.tex
trimming

@@
 Although \acp{CPS} present huge opportunities for all facets of society, they have yet to reach our daily lives in any relevant scale due to their stringent operational requirements.
-This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing~\cite{Satya2009Case,Satya2017Emergence}, consumer-grade \acp{CPS} will be made possible.
+This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing~\cite{Satya2017Emergence}, consumer-grade \acp{CPS} will be made possible.
 These technologies meet two key requirements of \acp{CPS}: real-time capabilities (through extremely low end-to-end latencies), and context- and locality-awareness, and will most likely become the backbone of \ac{CPS} in the future.
@@
 A subcategory of \acp{CPS} which stands to particularly benefit from the adoption of the above technologies is \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system.
-Depending on the physical system being controlled, \acp{NCS} can have extremely stringent timing and reliability requirements for the communication between plant and controller; requirements that conventional cloud computing and cellular networks are unable to meet~\cite{Liu2017Review,Wan2020Efficient}.
+Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for the communication between components that conventional cloud paradigms and cellular networks are unable to meet~\cite{Liu2017Review,Wan2020Efficient}.
 
-% These systems have been of great interest to the scientific community in the last decade.
-Due to their potential advantages for industrial and commercial settings~\cite{Lu2015WSAN}, a large body of work exists dedicated to the modelling and performance characterization of \aclp{NCS}~\cite{Hespanha2007Survey,Zhang2013Survey,Zhang2016Survey}.
+Due to their potential advantages for industrial and commercial settings, a large body of work exists dedicated to the modelling and performance characterization of \aclp{NCS}~\cite{Zhang2013Survey,Zhang2016Survey}.
 \Acp{NCS} enhance the flexibility of and improve existing control systems by allowing for the distribution of control functions over and across networks.
@@
 One approach to experimental research in \acp{NCS} uses setups in which the complete system is built on top of real hardware.
-This approach is employed in the works of, for instance, \textcite{Drew2005NCSWLAN,Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement and validate their results on physical testbeds of different kinds.
+This approach is employed in the works of, for instance, \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement and validate their results on physical testbeds of different kinds.
 Conversely, some studies choose to instead use completely \emph{simulated} \ac{NCS} setups.
-\textcite{Du2009Smith,Wu2012NPC,Chen2015synccontrol,Ma2019DynamicSched} are examples of works in which the authors have opted for such completely virtualized approaches.
+\textcite{Wu2012NPC,Chen2015synccontrol,Ma2019DynamicSched} are examples of works in which the authors have opted for such completely virtualized approaches.
 These studies often employ combinations of physical and network simulation libraries to try to capture the complex dynamics of \acp{NCS}.
@@
 \textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge with their \emph{NCSbench} platform.
-\emph{NCSbench} is an open-source \ac{NCS} benchmarking platform designed with reproducibility in mind, built using the \citetitle{LEGOMindstormsEV3}~\cite{LEGOMindstormsEV3} platform.
-It allows for easy and highly reproducible experimentation and benchmarking of \acp{NCS}, due to its low-cost, accessibility, and ease of reconfiguration.
+\emph{NCSbench} is an open-source \ac{NCS} benchmarking platform designed with reproducibility in mind, built on top of the LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform\footnote{\url{https://education.lego.com/en-us/products/lego-mindstorms-education-ev3-core-set/5003400\#lego-mindstorms-education-ev3}}.
+It allows for easy and highly reproducible benchmarking of \acp{NCS}, thanks to its low-cost, accessibility, and ease of reconfiguration.
 
@@
 ``Near'', ``far'', ``core'', and ``telco'' edge, among others, are all terms which describe variations of the original concept and which are becoming ubiquitous in new research.
-While the main idea of Edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the usefulness, strenghts, and weaknesses of such different edge concepts is becoming more and more important.
+While the core idea of Edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts, and weaknesses of such different edge concepts is of paramount importance.
 
 Our framework, which we name \ac{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems. 
-It is built on top of the \emph{Python 3.8}~\cite{Python3.8} language, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries and packages~\cite{pypi}.
-Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}~\cite{merkel2014docker}, making it suitable for automated deployment and benchmarking on industry-standard cloud and (more importantly) edge setups using container orchestration technologies.
+It is built on top of the \emph{Python 3.8} language, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries and packages.
+Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment and benchmarking on industry-standard cloud and (more importantly) edge setups using container orchestration technologies.
 
@@
 \Cref{sec:approach} presents the design principles of the framework.
-Next, in \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.
-Finally, we conclude this paper in \cref{sec:conclusion} with a brief discussion on the strengths and weaknesses of the platform, as well as on future directions to take.
+In \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.
+Finally,in \cref{sec:conclusion} we conclude this paper and briefly discuss future work.

### 2022-01-10 abstract.tex
fixing

--- 2020-11-30 abstract.tex
@@
 \begin{abstract}
-    \todo[inline]{TODO}
+    \acfp*{NCS} have long been a subject of theoretical research, however, little work has been devoted to repeatable and scalable experimentation.
+    The advent of edge computing and the subsequent proliferation of \acsp*{NCS} has quickly led this to become a key issue that needs to be addressed.
+
+    In this work, we present a novel, completely software-based approach to this challenge.
+    Our core idea rests on a framework for the emulation of physical plants which communicate with controllers implemented in software over a real network.
+    This approach is furthermore designed and built for the Edge, using a general-purpose programming language and will full compatibility with industry-standard containerization solutions.
+    We validate this framework using an initial implementation of an inverted pendulum \acs*{NCS}.
+    Our results showcase the utility of the tool as a repeatable, extensible, and scalable solution to \acs*{NCS} performance evaluation.
 \end{abstract}

### 2022-01-10 cleave.tex
fixing

@@
 The framework is distributed as \emph{free and open-source} software, and can be found online in the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
-Our design follows an approach similar to the above-discussed \emph{hardware-in-the-loop} setups, in that we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
+Our design follows a virtualized emulation approach; we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
 This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of and edge \acp{NCS}, namely the network itself.
+In other words, \ac{CLEAVE} is a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
+Our vision is to furthermore provide an open library of \ac{NCS} implementations in the future, this way making the tool even more accessible to the community.
 
@@
 We have chosen this system as an initial benchmark for its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
-However, we eventually plan to build an open library of emulated control loops.
+However, as mentioned above, we eventually plan to build an open library of emulated control loops.
 

### 2022-01-10 conclusion.tex
fixing

@@
 In this work, we aim to tackle this issue by presenting a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking with a particular focus on edge deployment.
-Our framework, \ac{CLEAVE}, is fully parametrizable and built with industry-standard edge technologies and paradigms in mind.
-We validate the utility of this tool through a series of scenarios, from which we are able to extract relevant metrics relating to the stability of the control system, as well as on the perfomance of the underlying network link.
-\todo[inline]{Talk a bit more about results?}
+Our framework, \ac{CLEAVE}, is fully parametrizable, extendable, and built with industry-standard edge technologies and paradigms in mind.
+We validate the utility of this tool through a series of scenarios; our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the perfomance of the underlying network link.
 We believe \ac{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \acp{NCS}.

### 2022-01-10 experiments.tex
fixing

@@
 
-We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with interesting and relevant results are then repeated an additional \num{20} times for better statistical significance in the results.
-\todo[inline]{Explain criteria for repeated experiments? Or just drop this sentence?}
+We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with sampling rates \( \geq \) \SI{20}{\hertz} and artificial delays \( \geq \) \SI{25}{\milli\second} where repeated an additional \num{20} times for better statistical significance.
 Each repetition lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.

### 2022-01-10 intro.tex
fixing

@@
 
-The number and applications of \acp{CPS}~\cite{Rajkumar2010CPS} --- that is, systems in which a real, physical mechanism is controlled by a computer --- have exploded in recent years.
-However, this rapid increase in adoption of these systems has been mostly limited to industrial contexts.
+The number and applications of \acp{CPS}~\cite{Rajkumar2010CPS} --- i.e.\ systems in which a real, physical mechanism is controlled by a computer --- have exploded in recent years.
+However, this rapid increase in adoption has mostly been limited to industrial contexts.
 Although \acp{CPS} present huge opportunities for all facets of society, they have yet to reach our daily lives in any relevant scale due to their stringent operational requirements.
-This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge (and Fog) Computing~\cite{Satya2017Emergence}, consumer-grade \acp{CPS} will be made possible.
+This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge Computing~\cite{Satya2017Emergence}, consumer-grade \acp{CPS} will be made possible.
 These technologies meet two key requirements of \acp{CPS}: real-time capabilities (through extremely low end-to-end latencies), and context- and locality-awareness, and will most likely become the backbone of \ac{CPS} in the future.
@@
 
-Due to their potential advantages for industrial and commercial settings, a large body of work exists dedicated to the modelling and performance characterization of \aclp{NCS}~\cite{Zhang2013Survey,Zhang2016Survey}.
-\Acp{NCS} enhance the flexibility of and improve existing control systems by allowing for the distribution of control functions over and across networks.
-This allows for, for instance, simultaneous centralized coordination, control, and monitoring of potentially hundreds of devices.
+Due to their potential advantages for industrial and commercial settings, a large body of work exists dedicated to the modelling and performance characterization of \acp{NCS}~\cite{Zhang2013Survey,Zhang2016Survey}.
+They improve and flexibilize existing control systems by allowing for the distribution of control functions over and across networks.
+This allows for, e.g.\ centralized coordination, control, and monitoring of potentially hundreds of devices.
 
@@
 One approach to experimental research in \acp{NCS} uses setups in which the complete system is built on top of real hardware.
-This approach is employed in the works of, for instance, \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement and validate their results on physical testbeds of different kinds.
-Conversely, some studies choose to instead use completely \emph{simulated} \ac{NCS} setups.
+This approach is employed in the works of, for instance, \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement and validate their results on physical testbeds.
+Conversely, other studies choose to instead use completely \emph{simulated} \ac{NCS} setups.
 \textcite{Wu2012NPC,Chen2015synccontrol,Ma2019DynamicSched} are examples of works in which the authors have opted for such completely virtualized approaches.
-These studies often employ combinations of physical and network simulation libraries to try to capture the complex dynamics of \acp{NCS}.
-Finally, a number of experimental studies instead employ \emph{hardware-} and \emph{network-in-the-loop} approaches.
-Hardware-in-the-loop refers to \ac{NCS} setups in which a real network interacts with a simulated or emulated control system~\cite{Wang2020VoltageControl}.
-On the other hand, network-in-the-loop refers to experimental setups in which the \emph{network} is instead simulated or emulated~\cite{Natale2004InvPendEthernet}.
+These studies often employ combinations of physical and network simulation tools to try to capture the complex dynamics of \acp{NCS}.
+Finally, a number of experimental studies instead employ \emph{virtualized} approaches, in which either
+\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; or }}]
+    \item a real network interacts with a simulated or emulated control system~\cite{Wang2020VoltageControl}
+    \item an emulated or simulated network interacts with a real control system~\cite{Natale2004InvPendEthernet}.
+\end{enumerate*}
 
-As evidenced above, experimental research in \acp{NCS} includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \acp{KPI}.
+As evidenced above, experimental research in \acp{NCS} includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \aclp*{KPI}.
 This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
@@
 \textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge with their \emph{NCSbench} platform.
-\emph{NCSbench} is an open-source \ac{NCS} benchmarking platform designed with reproducibility in mind, built on top of the LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform\footnote{\url{https://education.lego.com/en-us/products/lego-mindstorms-education-ev3-core-set/5003400\#lego-mindstorms-education-ev3}}.
-It allows for easy and highly reproducible benchmarking of \acp{NCS}, thanks to its low-cost, accessibility, and ease of reconfiguration.
+\emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built on top of the LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform, with stated goals of reproducibility and ease of use.
+It achives this thanks to its low-cost, accessibility, and ease of reconfiguration.
 
-Although the work of \textcite{Zoppi2020NCSBench} is a relevant and important step towards the reproducibility of \ac{NCS} benchmarks, we believe it still falls short on some aspects.
-In particular, we argue the reliance of \emph{NCSbench} on specific hardware for the Plant hampers the potential of the platform.
+Although \emph{NCSbench}~\cite{Zoppi2020NCSBench} is a relevant and important step towards the reproducibility of \ac{NCS} benchmarks, we believe it still falls short on some aspects.
+In particular, we argue the reliance of the tool on specific hardware for the Plant hampers the potential of the platform.
 The LEGO Mindstorms platform, although accessible and flexible, still fundamentally limits the physical systems that can be benchmarked to the kind of systems that can be physically implemented on it.

### 2022-01-10 cleave.tex
trimming

@@
 The framework is distributed as \emph{free and open-source} software, and can be found online in the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
-Our design follows a virtualized emulation approach; we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
-This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of and edge \acp{NCS}, namely the network itself.
+Our design follows a virtualized emulation approach; we provide a framework for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
+This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of the network, arguably the most complex and limiting component of and edge \acp{NCS}.
 In other words, \ac{CLEAVE} is a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
-Our vision is to furthermore provide an open library of \ac{NCS} implementations in the future, this way making the tool even more accessible to the community.
+Our vision is furthermore to provide an open library of \ac{NCS} implementations in the future, this way making the tool even more accessible to the community.
 
@@
     \caption{
-        Structure of an emulated \ac{NCS} on top of CLEAVE.\@
+        Structure of an emulated \ac{NCS} in \ac{CLEAVE}.\@
         The State object implements the discrete-time evolution of the Plant.
@@
 The first step in developing an \ac{NCS} emulation on top of \ac{CLEAVE} is the definition of a Plant which represents the physical system to be controlled.
-
 In abstract terms, Plants in \ac{CLEAVE} can be understood to be composed of
@@
 While \cref{item:plant:network} is handled transparently by CLEAVE, by hiding it through layers of abstraction, the rest of the items need to be provided in more explicit ways by the user.
-
 For \cref{item:plant:state}, the user needs to extend an abstract base class \texttt{State} provided by the framework.
@@
 \end{enumerate*}
-
-\Cref{item:plant:actuators,item:plant:sensors} are in the same way provided by the user through the extension of the \texttt{Actuator} and \texttt{Sensor} abstract base classes, respectively.
+\cref{item:plant:actuators,item:plant:sensors} are in the same way provided by the user through the extension of the \texttt{Actuator} and \texttt{Sensor} abstract base classes, respectively.
 \texttt{Actuator} objects process the outputs of the Controller and update the values of the \texttt{State} accordingly.
@@
 
-After the Plant has been defined, users need to implement a Controller for it.
+After the Plant has been defined, a Controller needs to be implemented for it.
 This is once again done by extending an abstract base class, \texttt{Controller}, provided by CLEAVE.\@
@@
 An example of such a configuration file can be seen in \cref{lst:config:plant}.
-These Python configuration files are then passed as arguments to the \texttt{cleave} bash command provided by the framework, which sets up and executes the emulation.
+These Python configuration files are then passed as arguments to the \texttt{cleave} command line provided by the framework, which sets up and executes the emulation.
 The use of Python scripts for configuration allows for a high degree of flexibility and complexity in the definition of emulation setups.
@@
     \caption{
-        The two-dimensional inverted pendulum system.
-        The objective of the system is to balance the vertically free-swinging pendulum by moving the cart.
+        The 2d inverted pendulum system, the objective of which is to balance the vertically free-swinging pendulum on the cart.
     }\label{fig:invpend}
@@
 \begin{listing}
-    \begin{minted}[linenos=true,
+    \begin{minted}[linenos=false,
         breaklines, 
         breakafter=d,
-        fontsize=\small]{Python}
+        fontsize=\footnotesize]{Python}
 from cleave.api.plant import SimpleConstantActuator, SimpleSensor

### 2022-01-10 experiments.tex
trimming

@@
 \section{Experimental validation}\label{sec:experiments}
-
-\input{tables/hardware.tex}
 
@@
 
-We depict our experimental edge setup in \cref{fig:cleave:expsetup}.
+We detail our experimental edge setup in \cref{fig:cleave:expsetup,tab:hardware}.
 It consists of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose Cloudlet.
@@
 Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which plants and controllers are executed co-located on the Cloudlet.
+
+\input{tables/hardware.tex}
 
@@
 As expected, higher sampling rates directly correlate to better quality of control (with the exception of the \SIlist{40;60}{\hertz}, \SI{50}{\milli\second} scenarios --- we attribute this to randomness and the relatively short duration of the experiments).
-As sampling rates increased, the system was able to reach stability at higher \ac{RTT} latencies.
+As sampling rates increased, the system was able to reach stability at higher \acp{RTT}.
 These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.
@@
 
-These results stem from the contention for network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acl{NCS} for deployment on the edge.
-The edge will be \emph{multi-tenant} and \emph{multi-instance}, and thus any edge \ac{NCS} deployment will have to be designed with shared resources in mind.
-\acp{NCS} will have to dynamically adapt to system limits in order to maintain quality of control.
+These results stem from the contention for network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acp{NCS} for deployment on the edge.
+The edge will be \emph{multi-tenant} and \emph{multi-instance}. 
+\acp{NCS} deployments will have to be designed with shared resources in mind, and given the complexity of these systems, experimental tools like \ac{CLEAVE} will be key for their succesful adoption and massification.

### 2022-01-10 intro.tex
trimming

@@
 \emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built on top of the LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform, with stated goals of reproducibility and ease of use.
-It achives this thanks to its low-cost, accessibility, and ease of reconfiguration.
+It achieves this thanks to its low-cost, accessibility, and ease of reconfiguration.
 
@@
 
-In this paper, we present the first fully-software-based \ac{NCS} benchmarking framework with scalability and repeatability as its main focus, built with the Edge in mind.
-In the years since the advent of the Edge computing paradigm, more and more variations of it have appeared in literature as the paradigm starts to be adopted by industry.
+In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \ac{NCS}.
+As the Edge computing paradigm starts to be adopted by industry, more and more variations of it have started to appear in literature.
 ``Near'', ``far'', ``core'', and ``telco'' edge, among others, are all terms which describe variations of the original concept and which are becoming ubiquitous in new research.
-While the core idea of Edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts, and weaknesses of such different edge concepts is of paramount importance.
+While the core idea of Edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts and weaknesses of such different edge concepts is of paramount importance.
 
@@
 It is built on top of the \emph{Python 3.8} language, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries and packages.
-Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment and benchmarking on industry-standard cloud and (more importantly) edge setups using container orchestration technologies.
+Additionally, it is compatible with \emph{containerization} technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment and benchmarking on industry-standard cloud and (more importantly) edge setups using container orchestration technologies.
 
 The rest of this paper is structured as follows.
-\Cref{sec:approach} presents the design principles of the framework.
+\cref{sec:approach} presents the design principles of the framework.
 In \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.
-Finally,in \cref{sec:conclusion} we conclude this paper and briefly discuss future work.
+Finally, in \cref{sec:conclusion} we conclude this paper and briefly discuss future work.

### 2022-01-10 acks.tex
version 11 jan

@@
 \section*{Acknowledgements}\label{sec:acks}
-\todo[inline]{TODO: James? Funding? Project course students?}
+
+This research has been partially funded by
+\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
+    \item the VINNOVA Competence Center for \ac{TECoSA} at KTH Royal Institute of Technology
+    \item the \ac{SSF}, through grant number ITM17--0246, awarded to James Gross, Ph.D.
+\end{enumerate*}
+Any opinions, findings, conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the view(s) of their employers or funding sources.
+
+We also wish to thank S. Mostafavi, V. N. Moothedath, N. Roy, M. Alsakati, O. Ferm, and S. Vojcic, without whose effort and support this work not have been possible.

### 2022-01-10 intro.tex
version 11 jan

@@
 A subcategory of \acp{CPS} which stands to particularly benefit from the adoption of the above technologies is \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system.
-Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for the communication between components that conventional cloud paradigms and cellular networks are unable to meet~\cite{Liu2017Review,Wan2020Efficient}.
+Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for the communication between components that conventional cloud paradigms and cellular networks are unable to meet~\cite{Wan2020Efficient}.
 

### 2022-01-11 cleave.tex
new approach section

@@
-\section{Approach}\label{sec:approach}
+\section{The \ac{CLEAVE} framework}\label{sec:approach}
 
@@
 
-Below, we will explain the approach taken in \ac{CLEAVE} by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation.
-An overview of the structure of such a deployment can be seen in \cref{fig:cleave:ncs:struct}.
+% Below, we will explain the approach taken in \ac{CLEAVE} by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation.
+% An overview of the structure of such a deployment can be seen in \cref{fig:cleave:ncs:struct}.
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=.8\textwidth]{CLEAVE_NCS_structure}
-    \caption{
-        Structure of an emulated \ac{NCS} in \ac{CLEAVE}.\@
-        The State object implements the discrete-time evolution of the Plant.
-        Actuated Variables are updated from the values obtained from the Actuator objects at the beginning of each time step and are used to update the State before each time-step.
-        Sensed Variables are sampled at the end of each time-step and passed on to the Sensor objects, which process them (e.g. to add noise) and send the processed values to the Controller.
-        Controller calculates actuation values and sends these to the Actuators.
-        Finally, the Actuators process the actuation values and hold the results until they are read at the next time-step.
-    }\label{fig:cleave:ncs:struct}
-\end{figure*}
+\subsection{Design and Implementation}
 
-\subsection{Defining a Plant}
+\cref{fig:cleave:ncs:struct} shows an overview of the conceptual structure and operation of \ac{CLEAVE}.
+The \emph{State} implements the discrete-time evolution of the Plant.
+At the beginning of each time-step, actuated variables in the \emph{State} are updated with values obtained from the \emph{Actuator} objects.
+Conversely, at the end of each time-step, sensed variables are sampled and used to update values associated with \emph{Sensor} objects.
+These values are processed at the \emph{Sensors} (e.g.\ to add noise) before being sent over the network to the \emph{Controller}.
+The \emph{Controller} calculates actuation values and updates the \emph{Actuators} over the network.
+Finally, the \emph{Actuators} process the actuation values and hold the results until they are read at the next time-step.
 
-The first step in developing an \ac{NCS} emulation on top of \ac{CLEAVE} is the definition of a Plant which represents the physical system to be controlled.
-In abstract terms, Plants in \ac{CLEAVE} can be understood to be composed of
+Implementation-wise, \ac{CLEAVE} is built on top of the well-established and stable \emph{Twisted}\footnote{\url{https://twistedmatrix.com}} asynchronous programming and networking framework.
+The aforementioned core components (\emph{State}, \emph{Controller}, \emph{Sensors}, and \emph{Actuators}) are provided as abstract base classes with core functionality that users need to extend for their specific use case and \ac{NCS}.
+On the other hand, the framework handles network communication and metric collection autonomously; users only need to select the desired transport protocol (we do however provide an \ac{API} for extending the available transports).
+
+Below we present a brief overview of the design and implementation of the core components.
+
+\subsubsection{State}
+
+The \emph{State} class implements the discrete-time emulation of the plant side of the \ac{NCS}.
+It is provided as an abstract base class with an abstract method
+\mintinline{python}{advance()} which users must implement.
+This method is called by the framework periodically according to the configured emulation update rate, and should perform a single time-step of the discrete-time emulation.
+Two additional, optionally extensible methods, \mintinline{python}{initialize()} and \mintinline{python}{shutdown()}, are provided in the \emph{State} base class.
+These are provided for users to implement single-fire set-up and tear-down procedures.
+
+The \emph{State} class also provides functionality for the automatic tracking of sensed values and updating of actuated values.
+This is handled through special classes for variables, \emph{SensorVariable} and \emph{ActuatorVariable}, which when instantiated and assigned to an attribute of the \emph{State} allow the framework to track and modify them.
+
+\subsubsection{Controller}
+
+The \emph{Controller} class defines a single abstract method \mintinline{python}{process()}, which takes a mapping of names to values corresponding to the sensed variables of the \emph{State}, and must return a similar mapping corresponding to the actuated variables.
+This method is called by \ac{CLEAVE} in an event-based manner whenever samples are received from the plant side; however, we intend to extend this implementation to time-interval-based approaches eventually as well.
+
+\subsubsection{Sensors}
+
+\emph{Sensors} implement the processing of monitored stated variables before sending them over the network to the \emph{Controller}.
+This base class must be initialized with a property name (corresponding to the name of the monitored \emph{State} attribute) and a sampling rate in \si{\hertz}, and provides an abstract method \mintinline{python}{process_sample()} which must be extended with an implementation of the processing performed on each sample.
+
+\subsubsection{Actuators}
+
+Finally, \emph{Actuators} perform the processing of actuation values obtained over the network from the \emph{Controller}.
+The base class is initialized with a property name corresponding to the name of the actuated \emph{State} attribute, and provides two abstract methods which must be implemented by extending classes:
 \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
-    \item\label{item:plant:state} an emulation of a physical system which is updated in a discrete-time fashion
-    \item\label{item:plant:actuators} procedures to modify or act upon the physical emulation at each time-step
-    \item\label{item:plant:sensors} procedures to sample relevant values from the physical emulation at each time-step
-    \item\label{item:plant:network} procedures to interact with the network and send and receive data from the Controller.
+    \item \mintinline{python}{set_value()}, which is called by the framework whenever a new value for the corresponding actuated attribute is received
+    \item \mintinline{python}{get_actuation()}, which is called on every time-step and must return a value to apply to the actuated attribute.
 \end{enumerate*}
 
-While \cref{item:plant:network} is handled transparently by CLEAVE, by hiding it through layers of abstraction, the rest of the items need to be provided in more explicit ways by the user.
-For \cref{item:plant:state}, the user needs to extend an abstract base class \texttt{State} provided by the framework.
-This class provides the base functionality and defines a simple API for the implementation of discrete-time emulations by
-\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
-    \item automatically tracking variables that can be modified and/or sampled by the framework
-    \item defining a single abstract method \texttt{State.advance()} which is invoked by the framework on each time step of the emulation.
-\end{enumerate*}
-\cref{item:plant:actuators,item:plant:sensors} are in the same way provided by the user through the extension of the \texttt{Actuator} and \texttt{Sensor} abstract base classes, respectively.
-\texttt{Actuator} objects process the outputs of the Controller and update the values of the \texttt{State} accordingly.
-Conversely, \texttt{Sensor} objects sample relevant values from the \texttt{State} object, process them, and finally pass them on to the Controller through the network.
-
-\subsection{Defining a Controller}
-
-After the Plant has been defined, a Controller needs to be implemented for it.
-This is once again done by extending an abstract base class, \texttt{Controller}, provided by CLEAVE.\@
-This class defines a single abstract method, \texttt{Controller.process()} which is called by the framework whenever samples are provided by the Plant.
-When invoked, it is passed a mapping of property names to sampled values.
-Implementations of this method should return yet another mapping of actuated property names to target values.
-Users are otherwise free to implement any kind of logic or process within extending classes.
-
-\subsection{Configuring and running the emulation}
+\subsection{Configuring and Executing \iac{NCS}}
 
 Once the core components for the emulation have been defined and implemented, the emulation needs to be configured to register them with the framework.
-This is done through fully-featured Python scripts which simply define a number of required (and some optional) top-level variables.
-An example of such a configuration file can be seen in \cref{lst:config:plant}.
-These Python configuration files are then passed as arguments to the \texttt{cleave} command line provided by the framework, which sets up and executes the emulation.
-The use of Python scripts for configuration allows for a high degree of flexibility and complexity in the definition of emulation setups.
+This is done through Python scripts which simply define a number of required (and some optional) top-level variables, corresponding to emulation parameters such as the \emph{State}, \emph{Controller}, \emph{Sensor}, and \emph{Actuator} subclasses to use; the emulation update rate; the controller side host and address; data output directories; and so on.
+These Python configuration files are then provided as arguments to the \texttt{cleave} command line provided by the framework.
+\ac{CLEAVE} imports and parses them, and then sets up and executes the emulation.
 
-\subsection{Implemented Control Loops}
+\subsection{Current Implementations}
 
@@
     \caption{
-        The 2d inverted pendulum system, the objective of which is to balance the vertically free-swinging pendulum on the cart.
+        The 2d inverted pendulum system, the goal of which is to balance the vertically free-swinging pendulum by applying horizontal forces on the cart.
     }\label{fig:invpend}
@@
 
-The inverted pendulum Plant is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
-For the Controller, a proportional-differential strategy is employed, implemented using the framework Controller API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
+The inverted pendulum plant \emph{State} is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
+\emph{Sensors} for the angle, position, angular velocity, and velocity of the \emph{State}, and an \emph{Actuator} for the horizontal force applied to the cart are implemented as ``perfect'', i.e.\ values are returned as-is, without any added noise.
+For the controller side, a proportional-differential strategy is employed, implemented using the framework \emph{Controller} API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
 The plant and controller are then packaged into Docker containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
-
-\begin{listing}
-    \begin{minted}[linenos=false,
-        breaklines, 
-        breakafter=d,
-        fontsize=\footnotesize]{Python}
-from cleave.api.plant import SimpleConstantActuator, SimpleSensor
-from cleave.impl.inverted_pendulum import InvPendulumState
-
-# Plant emulation parameters
-tick_rate = 100
-emu_duration = '30m'
-state = InvPendulumState(fail_angle_rad=0.34)
-output_dir = './plant_metrics'
-
-# Sensors to attach to the Plant
-sensors = [
-    SimpleSensor('position', 100),
-    SimpleSensor('speed', 100),
-    SimpleSensor('angle', 100),
-    SimpleSensor('ang_vel', 100),
-]
-
-actuators = [
-    SimpleConstantActuator(
-        initial_value=0,
-        prop_name='force')
-]
-    \end{minted}
-    \caption{
-        Example configuration file for a \ac{CLEAVE} Plant, defining the parameters for a single emulation.
-        % ``Plant emulation parameters'' define key emulation parameters, such as the emulation discrete-time update rate (the \emph{tick rate}).
-        % The Sensors and Actuators are then provided in lists, each Sensor and Actuator attached to a specific property identified by a string.
-        % Additionally, Sensors have a sampling rate in Hz which needs to be provided as an integer.
-    }
-    \label{lst:config:plant}
-\end{listing}

### 2022-01-12 abstract.tex
neel fixes

@@
     Our core idea rests on a framework for the emulation of physical plants which communicate with controllers implemented in software over a real network.
-    This approach is furthermore designed and built for the Edge, using a general-purpose programming language and will full compatibility with industry-standard containerization solutions.
+    This approach is furthermore designed and built for the Edge, using a general-purpose programming language and with full compatibility with industry-standard containerization solutions.
     We validate this framework using an initial implementation of an inverted pendulum \acs*{NCS}.

### 2022-01-12 cleave.tex
neel fixes

@@
 
-In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}, with a particular focus on edge deployments.
+In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \aclp{NCS}, with a particular focus on edge deployments.
 The framework is distributed as \emph{free and open-source} software, and can be found online in the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
@@
 In other words, \ac{CLEAVE} is a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
-Our vision is furthermore to provide an open library of \ac{NCS} implementations in the future, this way making the tool even more accessible to the community.
+Our vision is, furthermore, to provide an open library of \ac{NCS} implementations in the future, this way making the tool even more accessible to the community.
 
@@
 \cref{fig:cleave:ncs:struct} shows an overview of the conceptual structure and operation of \ac{CLEAVE}.
-The \emph{State} implements the discrete-time evolution of the Plant.
+The \emph{State} implements the discrete-time behavior of the Plant.
 At the beginning of each time-step, actuated variables in the \emph{State} are updated with values obtained from the \emph{Actuator} objects.

### 2022-01-12 experiments.tex
neel fixes

@@
 
-We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with sampling rates \( \geq \) \SI{20}{\hertz} and artificial delays \( \geq \) \SI{25}{\milli\second} where repeated an additional \num{20} times for better statistical significance.
+We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with sampling rates \( \geq \) \SI{20}{\hertz} and artificial delays \( \geq \) \SI{25}{\milli\second} were repeated an additional \num{20} times for better statistical significance.
 Each repetition lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
@@
 These results are interesting in their counter-intuitiveness.
-Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly shows this not to be the case for \ac{NCS} in resource-constrained scenarios.
+Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly show this not to be the case for \ac{NCS} in resource-constrained scenarios.
 \SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
@@
 The edge will be \emph{multi-tenant} and \emph{multi-instance}. 
-\acp{NCS} deployments will have to be designed with shared resources in mind, and given the complexity of these systems, experimental tools like \ac{CLEAVE} will be key for their succesful adoption and massification.
+\ac{NCS} deployments will have to be designed with shared resources in mind, and given the complexity of these systems, experimental tools like \ac{CLEAVE} will be key for their succesful adoption and massification.

### 2022-01-12 intro.tex
new approach section
neel fixes

@@
 \section{Introduction}\label{sec:intro}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=.8\textwidth]{CLEAVE_NCS_structure}
+    \caption{
+        Structure of an emulated \ac{NCS} in \acs*{CLEAVE}.
+    }\label{fig:cleave:ncs:struct}
+\end{figure*}
 
@@
 
-Most of the large literature concerning \acp{NCS}, however, follows a theoretical approach, and only a small fraction of it deals with experimental studies.
+Most of the literature concerning \acp{NCS}, however, follows a theoretical approach, and only a small fraction of it deals with experimental studies.
 The inherent inter-domain nature of \acp{NCS}, which intertwines knowledge from the fields of communications, computing, and control theory, coupled with the complexity of experimental platforms make these kinds of studies hard to perform.

### 2022-01-13 abstract.tex
fixes that james wanted

@@
 \begin{abstract}
-    \acfp*{NCS} have long been a subject of theoretical research, however, little work has been devoted to repeatable and scalable experimentation.
+    \acfp*{NCS} have long been a subject of both theoretical and experimental research. 
+    However, the inter-domain nature of such systems has led to a number of task-specific approaches to experimental research, with little work devoted to generalizable, repeatable, and scalable experimentation.
     The advent of edge computing and the subsequent proliferation of \acsp*{NCS} has quickly led this to become a key issue that needs to be addressed.
@@
     Our core idea rests on a framework for the emulation of physical plants which communicate with controllers implemented in software over a real network.
-    This approach is furthermore designed and built for the Edge, using a general-purpose programming language and with full compatibility with industry-standard containerization solutions.
+    This approach is furthermore designed and built for the edge, using a general-purpose programming language and with full compatibility with industry-standard containerization solutions.
     We validate this framework using an initial implementation of an inverted pendulum \acs*{NCS}.

### 2022-01-13 cleave.tex
fixes that james wanted

@@
 In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \aclp{NCS}, with a particular focus on edge deployments.
-The framework is distributed as \emph{free and open-source} software, and can be found online in the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
-Our design follows a virtualized emulation approach; we provide a framework for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
-This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of the network, arguably the most complex and limiting component of and edge \acp{NCS}.
+Our design follows a virtualized emulation approach; we provide a framework for the real-time \emph{emulation} of physical control systems and the easy implementation of softwarized controllers which interact with \emph{real} networks.
 In other words, \ac{CLEAVE} is a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
-Our vision is, furthermore, to provide an open library of \ac{NCS} implementations in the future, this way making the tool even more accessible to the community.
+This allows for great flexibility, as the core components of the \ac{NCS} can easily be switched out while maintaining the realism of the network, arguably the most complex and limiting component of edge \acp{NCS}.
 
-% Below, we will explain the approach taken in \ac{CLEAVE} by walking the reader through the steps necessary to deploy an arbitrary \ac{NCS} emulation.
-% An overview of the structure of such a deployment can be seen in \cref{fig:cleave:ncs:struct}.
+The framework is distributed as \emph{free and open-source} softwarethrough the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
+Furthermore, we plan to eventually provide an open library of \ac{NCS} implementations, thus making the tool even more accessible to the community.
 
@@
 
-The \emph{State} class also provides functionality for the automatic tracking of sensed values and updating of actuated values.
-This is handled through special classes for variables, \emph{SensorVariable} and \emph{ActuatorVariable}, which when instantiated and assigned to an attribute of the \emph{State} allow the framework to track and modify them.
+The \emph{State} class also provides functionality for the automatic tracking and updating of sensed and actuated values.
+This is handled through special classes, \emph{SensorVariable} and \emph{ActuatorVariable}, which when instantiated and assigned to an attribute of the \emph{State} allow the framework to track and modify them.
 
@@
 \emph{Sensors} implement the processing of monitored stated variables before sending them over the network to the \emph{Controller}.
-This base class must be initialized with a property name (corresponding to the name of the monitored \emph{State} attribute) and a sampling rate in \si{\hertz}, and provides an abstract method \mintinline{python}{process_sample()} which must be extended with an implementation of the processing performed on each sample.
+This base class must be initialized with a property name (corresponding to the monitored \emph{State} attribute) and a sampling rate in \si{\hertz}, and provides an abstract method \mintinline{python}{process_sample()} which must be extended with an implementation of the processing performed on each sample.
 
@@
 
-Once the core components for the emulation have been defined and implemented, the emulation needs to be configured to register them with the framework.
+Once the core components for the emulation have been defined and implemented, they need to be registered with the framework.
 This is done through Python scripts which simply define a number of required (and some optional) top-level variables, corresponding to emulation parameters such as the \emph{State}, \emph{Controller}, \emph{Sensor}, and \emph{Actuator} subclasses to use; the emulation update rate; the controller side host and address; data output directories; and so on.
@@
 
-The inverted pendulum plant \emph{State} is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz}.
+The inverted pendulum plant \emph{State} is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz} (this value is configurable; in this case it corresponds to the maximum achievable stable rate on the clients).
 \emph{Sensors} for the angle, position, angular velocity, and velocity of the \emph{State}, and an \emph{Actuator} for the horizontal force applied to the cart are implemented as ``perfect'', i.e.\ values are returned as-is, without any added noise.
-For the controller side, a proportional-differential strategy is employed, implemented using the framework \emph{Controller} API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
-The plant and controller are then packaged into Docker containers, for ease of orchestration, deployment, and re-parametrization, as well as to mimic real-world deployment.
+For the controller side, a proportional-differential strategy is implemented using the framework \emph{Controller} API and the \emph{NumPy} numeric computation library\footnote{NumPy: \url{https://numpy.org/}}.
+Plant and controller are then packaged into containers, for ease of orchestration and reparametrization, and to mimic real-world deployment.

### 2022-01-13 conclusion.tex
fixes that james wanted

@@
 
-The issue of repeatable and scalable benchmarks has been largely glossed over in \ac{NCS} literature; existing research is mostly focused on \emph{theory}, and the small fraction of experimental research studies available tend to implement \emph{ad-hoc} solutions.
+The issue of repeatable and scalable benchmarks has been largely glossed over in \ac{NCS} literature, as existing experimental research studies tend to implement \emph{ad-hoc} solutions.
 To the best of our knowledge, the work by \textcite{Zoppi2020NCSBench} is the only available framework for \ac{NCS} benchmarking focused on repeatability and reproducibility.
@@
 In this work, we aim to tackle this issue by presenting a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking with a particular focus on edge deployment.
-Our framework, \ac{CLEAVE}, is fully parametrizable, extendable, and built with industry-standard edge technologies and paradigms in mind.
+Our framework, \ac{CLEAVE}, allows for the easy implementation of emulated \acp{NCS} on real networks.
+It is fully parametrizable, extendable, and built with industry-standard edge technologies and paradigms in mind.
 We validate the utility of this tool through a series of scenarios; our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the perfomance of the underlying network link.
@@
 There is still, however, work to be done.
-To start with, the current implementation of \ac{CLEAVE} only includes a single plant-controller pair.
-We envision creating an open library of plants and controllers to share with the community.
-\ac{CLEAVE}'s integration with industry-standard tools and frameworks such as Docker will be refined.
-At the moment, these interactions are still quite superficial.
-Our goal is to achieve a much tighter integration, for instance by providing the complete toolkit as pre-packaged container images.
-Finally, the validity of the results obtained by the framework will have to be verified through more thorough and realistic scenarios than what we have been able to show in this work.
-In particular, we intend to eventually perform large-scale experimentation targetting 5G cellular deployments, as this technology is set to become the backbone of edge networks in the near future.
-
-% Benchmarking human-in-the-loop applications is hard, given
-% their tight interaction with human users who complicate the scaling
-% and repeatability of experiments. In this paper, we have presented
-% a benchmarking suite for this type of applications, called Edge-
-% Droid 1.0, capable of cutting out the need for users in performance
-% evaluations. We achieve this by employing pre-recorded sensory
-% input traces which we play over the network to the real applica-
-% tion backend, employing a parameterized user model to react to
-% feedback. We demonstrate its utility through a series of use case
-% scenarios, from which we are able to extract metrics regarding la-
-% tency both in regards to the application itself and the hardware
-% stack. We believe the EdgeDroid 1.0 suite thus represents an impor-
-% tant rst step towards enabling inexpensive and low-complexity
-% large-scale research on the scaling limits of this type of applications,
-% a requirement for wide adoption of the technology.
-
-% Nonetheless, there is still future work to be done.
-% The user model presented in this paper is only preliminary, and
-% we are currently conducting research in characterizing user behav-
-% ior when interacting with WCA applications in order to present
-% a more complete model in the future. As mentioned in Section 4,
-% in EdgeDroid 1.0, our model is that of a user who does not suer
-% any of the shortcomings of real human users such as annoyance, fa-
-% tigue, frustration, nausea. Rather, EdgeDroid 1.0 models a perfectly
-% stoic user who is like an automaton and responds in a precisely
-% reproducible and deterministic manner to the same system stimu-
-% lus every time. Of course, no real human user is an automaton. In
-% the future, we envision creating many versions of EdgeDroid (i.e.,
-% EdgeDroid 2.0, EdgeDroid 3.0, etc.) that embody more human-like
-% user models that more accurately emulate attributes such as those
-% mentioned above. Experimental validation of these human-like user
-% models via user studies will be an important part of our future work.
-% We are also working on expanding the benchmarking suite to
-% also work rst with other types of Wearable Cognitive Assistance,
-% and later with other categories of human-in-the-loop applications.
-% Other types of WCA we will consider in future iterations of the tool
-% include real-time task-assistance WCA applications (such as the
-% Ping-Pong application described in [7]), which don’t have a linear
-% task model like task-guidance WCA and have tighter latency bounds
-% and context- and information-providing WCA applications, for
-% instance, applications which recognize faces and provide relevant
-% social-media information related to that person. The latter also do
-% not have a linear task model, but present more lax latency bounds.
+The current implementation only includes a single plant-controller pair; we envision creating an open library of \acp{NCS} to share with the community.
+At the moment, the interactions of \ac{CLEAVE} and tools such as Docker are still quite superficial.
+Our goal is to achieve a much tighter integration, e.g.\ by providing the toolkit as pre-packaged container images.
+Finally, the validity of the results obtained by the framework will have to be verified through more thorough, realistic scenarios than what we have been able to show in this work.
+In particular, we intend to perform large-scale experimentation targetting 5G cellular deployments, as this technology is set to become the backbone of edge networks in the near future.

### 2022-01-13 experiments.tex
fixes that james wanted

@@
         \includegraphics[width=.9\textwidth]{single_loop_toppled}
+        % \todo[inline]{Check consistency??}
         \caption{}%
@@
         \includegraphics[width=.9\textwidth]{single_loop_rms}
+        % \todo[inline]{Mark fully toppled experiments with an X.}
         \caption{}\label{fig:single:rms}
@@
 For these scenarios, we vary:
-\begin{itemize}
-    \item the sampling rate of the Plant state, setting it to \SIlist[list-final-separator={, or }]{10;20;40;60}{\hertz};
+\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
+    \item the sampling rate of the Plant state, setting it to \SIlist[list-final-separator={, or }]{10;20;40;60}{\hertz}
     \item the responsiveness of the Controller, adding fixed delays of \SIlist[list-final-separator={, or }]{0;25;50}{\milli\second} after the processing of each sample.
-\end{itemize}
+\end{enumerate*}
 
@@
 For instance, it is clear from \cref{fig:single:stability} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
-A corollary of this is that, conversely, at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
-Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.
+A corollary of this is, conversely, that at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
+Adaptive sampling might thus be a viable method for optimizing \ac{NCS} resource utilization at the edge.
 
@@
 
-In the following, we present and briefly discuss the results for the realistic scenario with network resource contention.
-
 \cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
@@
 Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly show this not to be the case for \ac{NCS} in resource-constrained scenarios.
-\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and average pendulum angle \ac{RMS} about \num{3} times that of the \SI{40}{\hertz} scenario.
-\SI{40}{\hertz} was in turn the second worst configuration --- although no pendula toppled in this scenario, average angle \ac{RMS} were double that of the \SI{20}{\hertz} setup.
+\SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and mean pendulum angle \ac{RMS} circa \num{3} times that of the \SI{40}{\hertz} scenario.
+\SI{40}{\hertz} was in turn the second worst configuration --- although it presented no toppled pendula, average angle \ac{RMS} doubled that of the \SI{20}{\hertz} setup.
 
 We can see an explanation for these behaviors in \cref{fig:video:drop,fig:video:rtt}.
-Whereas both the \SIlist{20;40}{\hertz} setups show datagram losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} of datagrams lost.
-The differences in \acp{RTT} results are equally dramatic; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approximately \num{3} times those for the \SIlist{20;40}{\hertz} setups (circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).
+Whereas both the \SIlist{20;40}{\hertz} setups show losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} of datagrams lost.
+The differences in \acp{RTT} results are equally telling; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approx.\ \num{3} times those for the \SIlist{20;40}{\hertz} setups.
 
-These results stem from the contention for network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acp{NCS} for deployment on the edge.
+%(circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).
+
+These results stem from the contention for network resources, and hint at important trade-offs system designers will have to take into consideration when designing and developing \acp{NCS} for deployment on the edge.
 The edge will be \emph{multi-tenant} and \emph{multi-instance}. 

### 2022-01-13 intro.tex
acronym typo
fixes that james wanted

@@
     \caption{
-        Structure of an emulated \ac{NCS} in \acs*{CLEAVE}.
+        Structure of an emulated \acl*{NCS} in \acs*{CLEAVE}.
     }\label{fig:cleave:ncs:struct}
@@
 Although \acp{CPS} present huge opportunities for all facets of society, they have yet to reach our daily lives in any relevant scale due to their stringent operational requirements.
-This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and Edge Computing~\cite{Satya2017Emergence}, consumer-grade \acp{CPS} will be made possible.
+This is about to change, however, as with the advent of novel wireless communication technologies as well as networking paradigms, such as cellular 5G and edge computing~\cite{Satya2017Emergence}, consumer-grade \acp{CPS} will be made possible.
 These technologies meet two key requirements of \acp{CPS}: real-time capabilities (through extremely low end-to-end latencies), and context- and locality-awareness, and will most likely become the backbone of \ac{CPS} in the future.
@@
 
-Most of the literature concerning \acp{NCS}, however, follows a theoretical approach, and only a small fraction of it deals with experimental studies.
-The inherent inter-domain nature of \acp{NCS}, which intertwines knowledge from the fields of communications, computing, and control theory, coupled with the complexity of experimental platforms make these kinds of studies hard to perform.
+% Most of the literature concerning \acp{NCS}, follows a theoretical approach, and only a small fraction of it deals with experimental studies.
+An important fraction of the literature concerning \acp{NCS} deals with experimental studies.
+\acp{NCS} have an inherently inter-domain nature that intertwines knowledge from the fields of communications, computing, and control theory in ways that cannot be studied in isolation.
+Experimental research is therefore key to fully understanding the behavior of these systems in real-world deployments.
+This same inter-domain nature, however, has led to a number of different approaches to this kind of studies, as researchers attempt to tackle related research questions from different domains of knowledge.
 
-One approach to experimental research in \acp{NCS} uses setups in which the complete system is built on top of real hardware.
+One such approach uses setups in which the complete system is built on top of real hardware.
 This approach is employed in the works of, for instance, \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement and validate their results on physical testbeds.
@@
 In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \ac{NCS}.
-As the Edge computing paradigm starts to be adopted by industry, more and more variations of it have started to appear in literature.
+As the edge computing paradigm starts to be adopted by industry, more and more variations of it have started to appear in literature.
 ``Near'', ``far'', ``core'', and ``telco'' edge, among others, are all terms which describe variations of the original concept and which are becoming ubiquitous in new research.
-While the core idea of Edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts and weaknesses of such different edge concepts is of paramount importance.
+While the core idea of edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts and weaknesses of such different edge concepts is of paramount importance.
 
-Our framework, which we name \ac{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems. 
-It is built on top of the \emph{Python 3.8} language, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries and packages.
-Additionally, it is compatible with \emph{containerization} technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment and benchmarking on industry-standard cloud and (more importantly) edge setups using container orchestration technologies.
+Our framework, \ac{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems.
+Our approach is fully virtualized, inspired by our previous work on benchmarking human-in-the-loop applications on the edge~\cite{Olguin2019EdgeDroid}.
+We provide a benchmarking framework and software development kit for the development of emulated physical systems and softwarized controllers.
+These virtual \acp{NCS} can then be deployed on real networks for reparametrizable, repeatable, and reproducible benchmarking of the complete system.
+
+\ac{CLEAVE} is built using \emph{Python 3.8}, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries.
+It is furthermore compatible with container technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment and benchmarking on industry-standard edge setups using container orchestration solutions.
 

### 2022-01-13 experiments.tex
results

@@
         \labelcref{fig:single:topple} shows the fraction of plants that toppled, per experimental setup.
-        \labelcref{fig:single:rms} shows the \ac{RMS} value of the angle for all pendula that did not topple in the single loop scenarios.
-        Error bars indicate \SI{95}{\percent} \acp{CI}.
+        \labelcref{fig:single:rms} shows the mean \ac{RMS} value of the angle with respect to the vertical axis in logarithmic scale; the red line indicates \( y = 180 \).
+        Error bars indicate \SI{95}{\percent} \acp{CI} in both plots.
         }%
@@
 \cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
-\cref{fig:single:topple} shows the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} for the pendula angles.
+\cref{fig:single:topple} shows the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} for the absolute pendula angles.
 Toppled plants were identified simply by analysis of the emulation data; instances whose pendulum angles reached values above a threshold\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.

### 2022-01-14 acks.tex
new approach section
fixes that james wanted
fixes
fix samie's name

--- 2022-01-08 acks.tex
@@
 \section*{Acknowledgements}\label{sec:acks}
-\todo[inline]{TODO: James? Funding? Project course students?}
+
+This research has been partially funded by
+\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
+    \item the VINNOVA Competence Center for \ac{TECoSA} at KTH Royal Institute of Technology
+    \item the \ac{SSF}, through grant number ITM17--0246.
+\end{enumerate*}
+
+We also wish to thank S.~S.~Mostafavi, V.~N.~Moothedath, M.~Alsakati, O.~Ferm, S.~Vojcic, and A.~Bhattacharya for their support and valuable contributions to this work.

### 2022-01-14 experiments.tex
new plots

@@
 
-As expected, higher sampling rates directly correlate to better quality of control (with the exception of the \SIlist{40;60}{\hertz}, \SI{50}{\milli\second} scenarios --- we attribute this to randomness and the relatively short duration of the experiments).
+As expected, higher sampling rates directly correlate to better quality of control (the \SI{40}{\hertz}, \SI{50}{\milli\second} scenario has a slightly inconsistent behavior that we attribute to randomness and the relatively short duration of the experiments).
 As sampling rates increased, the system was able to reach stability at higher \acp{RTT}.

### 2022-01-15 experiments.tex
new topple plot with errorbars

--- 2022-01-13 experiments.tex
@@
 
-As expected, higher sampling rates directly correlate to better quality of control (with the exception of the \SIlist{40;60}{\hertz}, \SI{50}{\milli\second} scenarios --- we attribute this to randomness and the relatively short duration of the experiments).
-As sampling rates increased, the system was able to reach stability at higher \acp{RTT}.
+As expected, higher sampling rates tend to correlate with better quality of control.
+% (the \SI{40}{\hertz}, \SI{50}{\milli\second} scenario has a slightly inconsistent behavior that we attribute to randomness and the relatively short duration of the experiments).
+At higher sampling rates the system was able to reach stability at higher \acp{RTT}.
 These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.

### 2022-01-15 cleave.tex
final fixes

@@
 Implementation-wise, \ac{CLEAVE} is built on top of the well-established and stable \emph{Twisted}\footnote{\url{https://twistedmatrix.com}} asynchronous programming and networking framework.
-The aforementioned core components (\emph{State}, \emph{Controller}, \emph{Sensors}, and \emph{Actuators}) are provided as abstract base classes with core functionality that users need to extend for their specific use case and \ac{NCS}.
+The aforementioned core components (\emph{State}, \emph{Controller}, \emph{Sensors}, and \emph{Actuators}) are provided as abstract base classes with core functionality that users must extend for their specific use cases and \acp{NCS}.
 On the other hand, the framework handles network communication and metric collection autonomously; users only need to select the desired transport protocol (we do however provide an \ac{API} for extending the available transports).
@@
 Once the core components for the emulation have been defined and implemented, they need to be registered with the framework.
-This is done through Python scripts which simply define a number of required (and some optional) top-level variables, corresponding to emulation parameters such as the \emph{State}, \emph{Controller}, \emph{Sensor}, and \emph{Actuator} subclasses to use; the emulation update rate; the controller side host and address; data output directories; and so on.
+This is done through Python scripts which simply define a number of required (and some optional) top-level variables, corresponding to emulation parameters such as the \emph{State}, \emph{Controller}, \emph{Sensor}, and \emph{Actuator} subclasses to use; the emulation update rate; the controller-side host and address; data output directories; \emph{et cetera}.
 These Python configuration files are then provided as arguments to the \texttt{cleave} command line provided by the framework.

### 2022-01-15 conclusion.tex
final fixes

@@
 There is still, however, work to be done.
-The current implementation only includes a single plant-controller pair; we envision creating an open library of \acp{NCS} to share with the community.
+We are extending the number of plant and controller implementations on the framework, with the goal of creating an open library of \acp{NCS} to share with the community.
 At the moment, the interactions of \ac{CLEAVE} and tools such as Docker are still quite superficial.

### 2022-01-15 experiments.tex
final fixes

--- 2022-01-13 experiments.tex
@@
 
-As expected, higher sampling rates directly correlate to better quality of control (with the exception of the \SIlist{40;60}{\hertz}, \SI{50}{\milli\second} scenarios --- we attribute this to randomness and the relatively short duration of the experiments).
-As sampling rates increased, the system was able to reach stability at higher \acp{RTT}.
+As expected, higher sampling rates tend to correlate with better quality of control; at higher sampling rates the system was able to reach stability at higher \acp{RTT}.
 These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.

### 2022-01-15 intro.tex
final fixes

@@
 In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \ac{NCS}.
-As the edge computing paradigm starts to be adopted by industry, more and more variations of it have started to appear in literature.
+As the edge computing paradigm starts to be adopted by industry, more and more variations of it have begun to appear in literature.
 ``Near'', ``far'', ``core'', and ``telco'' edge, among others, are all terms which describe variations of the original concept and which are becoming ubiquitous in new research.
@@
 The rest of this paper is structured as follows.
-\cref{sec:approach} presents the design principles of the framework.
+\cref{sec:approach} presents the design principles and implementation of the framework.
 In \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.

### 2022-02-10 cleave.tex
complex loops

@@
 The aforementioned core components (\emph{State}, \emph{Controller}, \emph{Sensors}, and \emph{Actuators}) are provided as abstract base classes with core functionality that users must extend for their specific use cases and \acp{NCS}.
-On the other hand, the framework handles network communication and metric collection autonomously; users only need to select the desired transport protocol (we do however provide an \ac{API} for extending the available transports).
+The framework makes no assumptions about the inner workings of the user implementations of these classes other than the interfaces they expose.
+Users are therefore free to implement arbitrarily complex functionality in these components.
+On the other hand, \ac{CLEAVE} handles network communication and metric collection autonomously; users only need to select the desired transport protocol (we do however provide an \ac{API} for extending the available transports).
 
@@
 
+Although the core design of \ac{CLEAVE} follows a single-loop approach, multi-loop scenarios can be easily implemented through the combination of multiple \ac{CLEAVE} instances.
+Furthermore, the framework makes no assumptions about the individual components other than the signatures of methods on the base classes that must be extended and implemented.
+Combined with the plain-Python emulation configuration scripts, this makes development of more complex scenarios merely a matter of implementing the desired functionality within the framework itself.
+
 \subsection{Current Implementations}

### 2022-02-10 cleave.tex
tweaking

@@
 Implementation-wise, \ac{CLEAVE} is built on top of the well-established and stable \emph{Twisted}\footnote{\url{https://twistedmatrix.com}} asynchronous programming and networking framework.
-The aforementioned core components (\emph{State}, \emph{Controller}, \emph{Sensors}, and \emph{Actuators}) are provided as abstract base classes with core functionality that users must extend for their specific use cases and \acp{NCS}.
+The aforementioned core components (\emph{State}, \emph{Controller}, \emph{Sensors}, and \emph{Actuators}) are provided as \acp{ABC} with core functionality that users must extend for their specific use cases and \acp{NCS}.
 The framework makes no assumptions about the inner workings of the user implementations of these classes other than the interfaces they expose.
@@
 
-\subsubsection{State}
+\begin{description}[style=nextline]
+    \item[State Base Class]
 
-The \emph{State} class implements the discrete-time emulation of the plant side of the \ac{NCS}.
-It is provided as an abstract base class with an abstract method
-\mintinline{python}{advance()} which users must implement.
-This method is called by the framework periodically according to the configured emulation update rate, and should perform a single time-step of the discrete-time emulation.
-Two additional, optionally extensible methods, \mintinline{python}{initialize()} and \mintinline{python}{shutdown()}, are provided in the \emph{State} base class.
-These are provided for users to implement single-fire set-up and tear-down procedures.
+    \ac{ABC} with an abstract method
+    \mintinline{python}{advance()} which users must extend with the implementation of discrete-time emulation of the plant side of the \ac{NCS}.
+    The \mintinline{python}{advance()} method is called by the framework periodically according to the configured emulation update rate, and should perform a single time-step of the discrete-time emulation.
+    Two additional, optionally extensible methods, \mintinline{python}{initialize()} and \mintinline{python}{shutdown()}, are provided for users to implement single-fire set-up and tear-down procedures.
 
-The \emph{State} class also provides functionality for the automatic tracking and updating of sensed and actuated values.
-This is handled through special classes, \emph{SensorVariable} and \emph{ActuatorVariable}, which when instantiated and assigned to an attribute of the \emph{State} allow the framework to track and modify them.
+    % This class also provides functionality for the automatic tracking and updating of sensed and actuated values.
+    % This is handled through special auxiliary classes, \emph{SensorVariable} and \emph{ActuatorVariable}, which when instantiated and assigned to an attribute of the \emph{State} allow the framework to track and modify them.
 
-\subsubsection{Controller}
+    \item[Controller Base Class]
 
-The \emph{Controller} class defines a single abstract method \mintinline{python}{process()}, which takes a mapping of names to values corresponding to the sensed variables of the \emph{State}, and must return a similar mapping corresponding to the actuated variables.
+    Defines a single abstract method \mintinline{python}{process()}, which takes a mapping of names to values corresponding to the sensed variables of the \emph{State}, and must return a similar mapping corresponding to the actuated variables.
     This method is called by \ac{CLEAVE} in an event-based manner whenever samples are received from the plant side; however, we intend to extend this implementation to time-interval-based approaches eventually as well.
 
-\subsubsection{Sensors}
+    \item[Sensor Base Class]
 
@@
 
-\subsubsection{Actuators}
+    \item[Actuator Base Class]
 
@@
     \end{enumerate*}
+
+\end{description}
 
@@
     \caption{
-        The 2d inverted pendulum system, the goal of which is to balance the vertically free-swinging pendulum by applying horizontal forces on the cart.
+        The 2d inverted pendulum system.
     }\label{fig:invpend}
@@
 
-For the purpose of this work, we have used \ac{CLEAVE} to implement an emulated \emph{inverted pendulum} control loop; see \cref{fig:invpend}.
+For the purpose of this work, we have used \ac{CLEAVE} to implement an emulated \emph{inverted pendulum} control loop (see \cref{fig:invpend}), the goal of which is to balance the vertically free-swinging pendulum by applying horizontal forces on the cart.
 We have chosen this system as an initial benchmark for its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.

### 2022-02-10 experiments.tex
tweaking

@@
 
-We detail our experimental edge setup in \cref{fig:cleave:expsetup,tab:hardware}.
-It consists of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose Cloudlet.
+We detail our experimental edge setup in \cref{fig:cleave:expsetup}.
+It consists of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose \verb|x86_64| desktop computer acting as a Cloudlet/Edge server.
 On top of this physical architecture we configure a Docker Swarm\footnote{Swarm mode overview: \url{https://docs.docker.com/engine/swarm/}} managed centrally from the Cloudlet.
@@
 
-\input{tables/hardware.tex}
+% \input{tables/hardware.tex}
 
@@
 \end{figure}
-
-\begin{figure*}[h]
-    \centering
-    \begin{subfigure}[t]{.5\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{single_loop_toppled}
-        % \todo[inline]{Check consistency??}
-        \caption{}%
-        \label{fig:single:topple}
-    \end{subfigure}%
-    \begin{subfigure}[t]{.5\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{single_loop_rms}
-        % \todo[inline]{Mark fully toppled experiments with an X.}
-        \caption{}\label{fig:single:rms}
-    \end{subfigure}%
-    \caption[caption]{
-        Stability metrics for the single-loop scenarios.
-        \labelcref{fig:single:topple} shows the fraction of plants that toppled, per experimental setup.
-        \labelcref{fig:single:rms} shows the mean \ac{RMS} value of the angle with respect to the vertical axis in logarithmic scale; the red line indicates \( y = 180 \).
-        Error bars indicate \SI{95}{\percent} \acp{CI} in both plots.
-        }%
-    \label{fig:single:stability}
-\end{figure*}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=.9\columnwidth]{single_loop_rtts}
-    \caption{
-        Mean latency due to network and processing for both single-loop scenarios.
-        Error bars indicate \SI{95}{\percent} \acp{CI}.
-    }\label{fig:single:rtt}
-\end{figure}
-
-\begin{figure*}[t]
-    \centering
-    \begin{subfigure}[h]{.25\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{video_topple_frac}
-        \caption{}\label{fig:video:toppled}
-    \end{subfigure}%
-    \begin{subfigure}[h]{.25\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{video_angle_rms}
-        \caption{}\label{fig:video:rms}
-    \end{subfigure}%
-    \begin{subfigure}[h]{.25\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{video_drop_frac}
-        \caption{}\label{fig:video:drop}
-    \end{subfigure}%
-    \begin{subfigure}[h]{.25\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{video_rtt}
-        \caption{}\label{fig:video:rtt}
-    \end{subfigure}%
-    \caption{
-        Results for the realistic scenario.
-        \labelcref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
-        \labelcref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
-        \labelcref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
-        \labelcref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
-        Error bars indicate \SI{95}{\percent} \ac{CI} in all plots.
-    }\label{fig:video:results}
-\end{figure*}
 
@@
 
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[t]{.45\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{single_loop_toppled}
+        \caption{Fraction of toppled plants.}\label{fig:single:topple}
+    \end{subfigure}\\
+    \begin{subfigure}[t]{.45\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{single_loop_rms}
+        \caption{Mean angle \acs*{RMS}; red line indicates \( y = 180 \).}\label{fig:single:rms}
+    \end{subfigure}\\
+    \begin{subfigure}[t]{.45\textwidth}
+        \centering
+        \includegraphics[width=.9\columnwidth]{single_loop_rtts}
+        \caption{
+            Mean \acsp*{RTT}.
+        }\label{fig:single:rtt}
+    \end{subfigure}%
+    \caption[caption]{
+        Metrics for the single-loop scenarios.
+        % \labelcref{fig:single:topple} shows the fraction of plants that toppled, per experimental setup.
+        % \labelcref{fig:single:rms} shows the mean \ac{RMS} value of the angle with respect to the vertical axis in logarithmic scale; the red line indicates \( y = 180 \).
+        Error bars indicate \SI{95}{\percent} \acp{CI}.
+        }%
+    \label{fig:single}
+\end{figure}
+
 We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable collection and posterior processing of relevant metrics.  
-\cref{fig:single:stability} shows the results relating to the stability of the plants in the single-loop scenarios:
+\cref{fig:single} shows a summary of the results relating to the single-loop scenarios:
 \cref{fig:single:topple} shows the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} for the absolute pendula angles.
-Toppled plants were identified simply by analysis of the emulation data; instances whose pendulum angles reached values above a threshold\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.} were marked as ``toppled''.
+Toppled plants were identified simply by analysis of the emulation data, indentifying instances whose pendulum angles reached values above a threshold value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.}.
 
@@
 These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.
-For instance, it is clear from \cref{fig:single:stability} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
+For instance, it is clear from \cref{fig:single:topple,fig:single:rms} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
 A corollary of this is, conversely, that at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
@@
 
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[h]{.22\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{video_topple_frac}
+        \caption{Toppled plants.}\label{fig:video:toppled}
+    \end{subfigure}%
+    \hfill%
+    \begin{subfigure}[h]{.22\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{video_angle_rms}
+        \caption{Angle \ac{RMS}.}\label{fig:video:rms}
+    \end{subfigure}\\
+    \begin{subfigure}[h]{.22\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{video_drop_frac}
+        \caption{Packet losses.}\label{fig:video:drop}
+    \end{subfigure}%
+    \hfill%
+    \begin{subfigure}[h]{.22\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{video_rtt}
+        \caption{\acsp{RTT}.}\label{fig:video:rtt}
+    \end{subfigure}%
+    \caption{
+        Results for the realistic scenario.
+        % \labelcref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
+        % \labelcref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
+        % \labelcref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
+        % \labelcref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
+        Error bars indicate \SI{95}{\percent} \acp{CI} in all plots.
+    }\label{fig:video:results}
+\end{figure}
+
 \cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
+\cref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
+\cref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
+\cref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
+\cref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
+
 These results are interesting in their counter-intuitiveness.

### 2022-02-10 abstract.tex
more tweaks

--- 2022-01-12 abstract.tex
@@
 \begin{abstract}
-    \acfp*{NCS} have long been a subject of theoretical research, however, little work has been devoted to repeatable and scalable experimentation.
-    The advent of edge computing and the subsequent proliferation of \acsp*{NCS} has quickly led this to become a key issue that needs to be addressed.
+    In this work, we present \acs*{CLEAVE}, a novel, completely software-based framework for repeatable and scalable experimentation in \acfp*{NCS}.
+    While these systems have long been a subject of both theoretical and experimental research, their inter-domain nature  has led to a wide number of task-specific approaches to research, and little work has been devoted to generalizable, repeatable, and scalable experimentation.
+    We aim to tackle this gap in knowledge through an approach based on the emulation of physical plants which communicate over a real network with controllers implemented in software.
+    \acs*{CLEAVE} is furthermore designed and built for the edge, using Python3 and with full compatibility with industry-standard containerization solutions.
+    Although designed for single-loop emulations, the flexibility afforded by the aformentioned characteristics allow our framework to be adaptated to and employed in a multitude of more complex scenarios.
 
-    In this work, we present a novel, completely software-based approach to this challenge.
-    Our core idea rests on a framework for the emulation of physical plants which communicate with controllers implemented in software over a real network.
-    This approach is furthermore designed and built for the Edge, using a general-purpose programming language and with full compatibility with industry-standard containerization solutions.
-    We validate this framework using an initial implementation of an inverted pendulum \acs*{NCS}.
+    We validate \acs*{CLEAVE} using an initial implementation of an inverted pendulum \acs*{NCS}.
     Our results showcase the utility of the tool as a repeatable, extensible, and scalable solution to \acs*{NCS} performance evaluation.

### 2022-02-10 intro.tex
more tweaks

@@
 Our framework, \ac{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems.
-Our approach is fully virtualized, inspired by our previous work on benchmarking human-in-the-loop applications on the edge~\cite{Olguin2019EdgeDroid}.
-We provide a benchmarking framework and software development kit for the development of emulated physical systems and softwarized controllers.
+\ac{CLEAVE} is fully virtualized, inspired by our previous work on benchmarking human-in-the-loop applications on the edge~\cite{Olguin2019EdgeDroid}.
+The tool consists of a benchmarking framework and software development kit for the development of emulated physical systems and softwarized controllers.
 These virtual \acp{NCS} can then be deployed on real networks for reparametrizable, repeatable, and reproducible benchmarking of the complete system.
@@
 \ac{CLEAVE} is built using \emph{Python 3.8}, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries.
-It is furthermore compatible with container technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment and benchmarking on industry-standard edge setups using container orchestration solutions.
+It is furthermore compatible with container technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment, scaling, and benchmarking on industry-standard edge setups using container orchestration solutions.
 

### 2022-02-11 abstract.tex
abstract

@@
 \begin{abstract}
-    In this work, we present \acs*{CLEAVE}, a novel, completely software-based framework for repeatable and scalable experimentation in \acfp*{NCS}.
+    In this work, we present \acs*{CLEAVE}, a novel, completely software based framework for repeatable and scalable experimentation in \acfp*{NCS}.
     While these systems have long been a subject of both theoretical and experimental research, their inter-domain nature  has led to a wide number of task-specific approaches to research, and little work has been devoted to generalizable, repeatable, and scalable experimentation.

### 2022-02-11 cleave.tex

@@
 
-The framework is distributed as \emph{free and open-source} softwarethrough the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
+The framework is distributed as \emph{free and open-source} software through the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
 Furthermore, we plan to eventually provide an open library of \ac{NCS} implementations, thus making the tool even more accessible to the community.

### 2022-02-11 conclusion.tex
Update on Overleaf.

@@
 The issue of repeatable and scalable benchmarks has been largely glossed over in \ac{NCS} literature, as existing experimental research studies tend to implement \emph{ad-hoc} solutions.
-To the best of our knowledge, the work by \textcite{Zoppi2020NCSBench} is the only available framework for \ac{NCS} benchmarking focused on repeatability and reproducibility.
 
@@
 It is fully parametrizable, extendable, and built with industry-standard edge technologies and paradigms in mind.
-We validate the utility of this tool through a series of scenarios; our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the perfomance of the underlying network link.
+We validate the utility of this tool through a series of scenarios; our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
 We believe \ac{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \acp{NCS}.

### 2022-02-11 intro.tex
Update on Overleaf.

@@
 Due to their potential advantages for industrial and commercial settings, a large body of work exists dedicated to the modelling and performance characterization of \acp{NCS}~\cite{Zhang2013Survey,Zhang2016Survey}.
-They improve and flexibilize existing control systems by allowing for the distribution of control functions over and across networks.
-This allows for, e.g.\ centralized coordination, control, and monitoring of potentially hundreds of devices.
+They improve and flexibilize existing control systems by allowing for the distribution of control functions across networks, allowing for centralized coordination, control, and monitoring of a multitude of devices.
 
 % Most of the literature concerning \acp{NCS}, follows a theoretical approach, and only a small fraction of it deals with experimental studies.
-An important fraction of the literature concerning \acp{NCS} deals with experimental studies.
-\acp{NCS} have an inherently inter-domain nature that intertwines knowledge from the fields of communications, computing, and control theory in ways that cannot be studied in isolation.
-Experimental research is therefore key to fully understanding the behavior of these systems in real-world deployments.
-This same inter-domain nature, however, has led to a number of different approaches to this kind of studies, as researchers attempt to tackle related research questions from different domains of knowledge.
-
+A number of works concerning \acp{NCS} deal with experimental studies.
+\acp{NCS} have an inherently inter-domain nature intertwining knowledge from the fields of communications, computing, and control theory in ways that cannot be studied in isolation, leading to various different approaches to such studies.
 One such approach uses setups in which the complete system is built on top of real hardware.
-This approach is employed in the works of, for instance, \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement and validate their results on physical testbeds.
+This approach is employed in the works of, for instance, \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement their approach on physical testbeds.
 Conversely, other studies choose to instead use completely \emph{simulated} \ac{NCS} setups.
@@
 These studies often employ combinations of physical and network simulation tools to try to capture the complex dynamics of \acp{NCS}.
-Finally, a number of experimental studies instead employ \emph{virtualized} approaches, in which either
+Finally, some experimental studies instead employ \emph{virtualized} approaches, in which either
 \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; or }}]
     \item a real network interacts with a simulated or emulated control system~\cite{Wang2020VoltageControl}
-    \item an emulated or simulated network interacts with a real control system~\cite{Natale2004InvPendEthernet}.
+    \item a simulated network interacts with a real control system~\cite{Natale2004InvPendEthernet}.
 \end{enumerate*}
 
-As evidenced above, experimental research in \acp{NCS} includes a large variety of heterogeneous hardware and software platforms, as well as methodologies and \aclp*{KPI}.
+As evidenced above, experimental research in \acp{NCS} includes varied heterogeneous hardware and software platforms, methodologies and \aclp*{KPI}.
+
 This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
@@
 
-\textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge with their \emph{NCSbench} platform.
-\emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built on top of the LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform, with stated goals of reproducibility and ease of use.
-It achieves this thanks to its low-cost, accessibility, and ease of reconfiguration.
+\textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge in their work.
+Their proposed platform \emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built utilizing LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform, with stated goals of reproducibility and ease of use.
+It achieves them thanks to its low-cost, accessibility, and ease of reconfiguration.
 
-Although \emph{NCSbench}~\cite{Zoppi2020NCSBench} is a relevant and important step towards the reproducibility of \ac{NCS} benchmarks, we believe it still falls short on some aspects.
-In particular, we argue the reliance of the tool on specific hardware for the Plant hampers the potential of the platform.
-The LEGO Mindstorms platform, although accessible and flexible, still fundamentally limits the physical systems that can be benchmarked to the kind of systems that can be physically implemented on it.
+Although it is an important step towards the reproducibility of \ac{NCS} benchmarks, we believe it still falls short on some aspects.
+In particular, we argue the reliance on specific hardware for the Plant which although flexible, limits the benchmarking to only those systems which can be physically implemented on it.
 Furthermore, it is \emph{not} scalable, and thus unsuitable for studies of larger, distributed \acp{NCS}.
@@
 
-\ac{CLEAVE} is built using \emph{Python 3.8}, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries.
+\ac{CLEAVE} is built using \emph{Python 3.8}, making it highly extensible and able to harness the multitude of already existing user-provided libraries.
 It is furthermore compatible with container technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment, scaling, and benchmarking on industry-standard edge setups using container orchestration solutions.

### 2022-02-11 cleave.tex

@@
 
-\begin{description}[style=nextline]
-    \item[State Base Class]
-
+%\begin{description}[style=nextline]
+%\begin{itemize}
+\textbf{State Base Class}\\
 \ac{ABC} with an abstract method
-    \mintinline{python}{advance()} which users must extend with the implementation of discrete-time emulation of the plant side of the \ac{NCS}.
+\mintinline{python}{advance()} which users must extend with the implementation of discrete-time emulation of the plant.
 The \mintinline{python}{advance()} method is called by the framework periodically according to the configured emulation update rate, and should perform a single time-step of the discrete-time emulation.
-    Two additional, optionally extensible methods, \mintinline{python}{initialize()} and \mintinline{python}{shutdown()}, are provided for users to implement single-fire set-up and tear-down procedures.
+Two optionally extensible methods, \mintinline{python}{initialize()} and \mintinline{python}{shutdown()} are provided for users to implement single-fire set-up and tear-down procedures.
 
@@
 
-    \item[Controller Base Class]
+\textbf{Controller Base Class}\\
+Defines a single abstract method \mintinline{python}{process()}, which takes a mapping of names to values corresponding to the sensed variables of the \emph{State}, and must return a similar mapping corresponding to the actuated variables.
+This method is called by \ac{CLEAVE} in an event-based manner whenever samples are received from the plant side; however, we intend to extend this implementation to time-interval-based approaches.
 
-    Defines a single abstract method \mintinline{python}{process()}, which takes a mapping of names to values corresponding to the sensed variables of the \emph{State}, and must return a similar mapping corresponding to the actuated variables.
-    This method is called by \ac{CLEAVE} in an event-based manner whenever samples are received from the plant side; however, we intend to extend this implementation to time-interval-based approaches eventually as well.
-
-    \item[Sensor Base Class]
-
+\textbf{Sensor Base Class}\\
 \emph{Sensors} implement the processing of monitored stated variables before sending them over the network to the \emph{Controller}.
@@
 
-    \item[Actuator Base Class]
-
+\textbf{Actuator Base Class}\\
 Finally, \emph{Actuators} perform the processing of actuation values obtained over the network from the \emph{Controller}.
@@
 \end{enumerate*}
-
-\end{description}
+%\end{itemize}
+%\end{description}
 
@@
     \caption{
-        The 2d inverted pendulum system.
+        The 2D inverted pendulum system.
     }\label{fig:invpend}

### 2022-02-11 experiments.tex

@@
 Scenarios are executed automatically in batches using a simple Python script which interacts with Docker through the widely adopted \emph{docker-py}\footnote{Docker SDK for Python: \url{https://docker-py.readthedocs.io/en/stable/}} library.
-This is \ac{CLEAVE}'s first advantage over existing frameworks; it is designed with cloud and edge technology and paradigms in mind, making integration with existing systems straightforward.
+This is \ac{CLEAVE}'s first advantage over existing frameworks; it is designed with cloud and edge technology and paradigms in mind, making integration with existing systems convenient.
 

### 2022-02-11 abstract.tex
Update on Overleaf.

@@
 \begin{abstract}
-    In this work, we present \acs*{CLEAVE}, a novel, completely software based framework for repeatable and scalable experimentation in \acfp*{NCS}.
-    While these systems have long been a subject of both theoretical and experimental research, their inter-domain nature  has led to a wide number of task-specific approaches to research, and little work has been devoted to generalizable, repeatable, and scalable experimentation.
-    We aim to tackle this gap in knowledge through an approach based on the emulation of physical plants which communicate over a real network with controllers implemented in software.
-    \acs*{CLEAVE} is furthermore designed and built for the edge, using Python3 and with full compatibility with industry-standard containerization solutions.
-    Although designed for single-loop emulations, the flexibility afforded by the aformentioned characteristics allow our framework to be adaptated to and employed in a multitude of more complex scenarios.
+    As the number of cyber-physical systems rises, it becomes increasingly crucial to study \acfp*{NCS} combining control communication co-design.
+This nature of NCSs has led to task-specific approaches to research, creating a dearth of generalizable, repeatable, and scalable experimentation.
+Further, with the advent of edge computing solutions, it is of paramount importance to explore its relevance in such applications.
+In this work, we present \acs*{CLEAVE}, a novel, completely software-based framework for repeatable and scalable experimentation in edge native NCSs.
+    Our approach is based on the emulation of physical plants communicating over a real network with software-based controllers.
+    \acs*{CLEAVE} is designed and built for the edge, using Python3 and with full compatibility with industry-standard containerization solutions.
+    Although designed for single-loop emulations, the flexibility afforded by the aformentioned characteristics allow our framework to be adapted to a multitude of complex scenarios.
 
     We validate \acs*{CLEAVE} using an initial implementation of an inverted pendulum \acs*{NCS}.
-    Our results showcase the utility of the tool as a repeatable, extensible, and scalable solution to \acs*{NCS} performance evaluation.
+    Our results showcase the utility of the tool as a repeatable, extensible, and scalable solution to \acs*{NCS} performance evaluation and benchmarking.
 \end{abstract}

### 2022-02-11 cleave.tex
Update on Overleaf.

@@
 
-In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \aclp{NCS}, with a particular focus on edge deployments.
-Our design follows a virtualized emulation approach; we provide a framework for the real-time \emph{emulation} of physical control systems and the easy implementation of softwarized controllers which interact with \emph{real} networks.
-In other words, \ac{CLEAVE} is a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
-This allows for great flexibility, as the core components of the \ac{NCS} can easily be switched out while maintaining the realism of the network, arguably the most complex and limiting component of edge \acp{NCS}.
+In this section, we detail our \ac{CLEAVE} framework for the performance evaluation of \aclp{NCS}, with an emphasis on edge deployments.
+Our design follows a virtualized emulation approach; providing a framework for the real-time \emph{emulation} of physical control systems and the easy implementation of softwarized controllers interacting with \emph{real} networks.
+\ac{CLEAVE} is thus, a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
+This allows for great flexibility, as the core components of the \ac{NCS} can be switched out while maintaining the realism of the network, the most complex and limiting component of edge \acp{NCS}.
 
 The framework is distributed as \emph{free and open-source} software through the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
-Furthermore, we plan to eventually provide an open library of \ac{NCS} implementations, thus making the tool even more accessible to the community.
+We also plan to provide a library of \ac{NCS} implementations, making the tool even more accessible.
 
@@
 %\begin{itemize}
-\textbf{State Base Class}\\
+\textbf{State Base Class}:
 \ac{ABC} with an abstract method
-\mintinline{python}{advance()} which users must extend with the implementation of discrete-time emulation of the plant.
+\mintinline{python}{advance()} which must be extended with the implementation of discrete-time emulation of the plant.
 The \mintinline{python}{advance()} method is called by the framework periodically according to the configured emulation update rate, and should perform a single time-step of the discrete-time emulation.
-Two optionally extensible methods, \mintinline{python}{initialize()} and \mintinline{python}{shutdown()} are provided for users to implement single-fire set-up and tear-down procedures.
+Two optionally extensible methods, \mintinline{python}{initialize()} and \mintinline{python}{shutdown()} are provided to implement single-fire set-up and tear-down procedures.
 
@@
 
-\textbf{Controller Base Class}\\
-Defines a single abstract method \mintinline{python}{process()}, which takes a mapping of names to values corresponding to the sensed variables of the \emph{State}, and must return a similar mapping corresponding to the actuated variables.
-This method is called by \ac{CLEAVE} in an event-based manner whenever samples are received from the plant side; however, we intend to extend this implementation to time-interval-based approaches.
+\textbf{Controller Base Class}:
+Defines a single abstract method \mintinline{python}{process()}, taking a mapping of names to values corresponding to the sensed variables of the \emph{State}, and returning a similar mapping corresponding to the actuated variables.
+This method is called by \ac{CLEAVE} in an event-based manner whenever samples are received from the plant side; however, we intend to extend this to time-interval-based approaches.
 
-\textbf{Sensor Base Class}\\
-\emph{Sensors} implement the processing of monitored stated variables before sending them over the network to the \emph{Controller}.
+\textbf{Sensor Base Class}:
+Implements processing of monitored stated variables before sending them over the network to the \emph{Controller}.
 This base class must be initialized with a property name (corresponding to the monitored \emph{State} attribute) and a sampling rate in \si{\hertz}, and provides an abstract method \mintinline{python}{process_sample()} which must be extended with an implementation of the processing performed on each sample.
 
-\textbf{Actuator Base Class}\\
-Finally, \emph{Actuators} perform the processing of actuation values obtained over the network from the \emph{Controller}.
-The base class is initialized with a property name corresponding to the name of the actuated \emph{State} attribute, and provides two abstract methods which must be implemented by extending classes:
+\textbf{Actuator Base Class}:
+Performs the processing of actuation values obtained over the network from the \emph{Controller}.
+It is initialized with a property name corresponding to the name of the actuated \emph{State} attribute, and provides two abstract methods which must be implemented by extending classes:
 \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
-\item \mintinline{python}{set_value()}, which is called by the framework whenever a new value for the corresponding actuated attribute is received
-\item \mintinline{python}{get_actuation()}, which is called on every time-step and must return a value to apply to the actuated attribute.
+\item \mintinline{python}{set_value()}, called by the framework whenever a new value for the corresponding actuated attribute is received
+\item \mintinline{python}{get_actuation()}, called on every time-step and must return a value to apply to the actuated attribute.
 \end{enumerate*}
@@
 Once the core components for the emulation have been defined and implemented, they need to be registered with the framework.
-This is done through Python scripts which simply define a number of required (and some optional) top-level variables, corresponding to emulation parameters such as the \emph{State}, \emph{Controller}, \emph{Sensor}, and \emph{Actuator} subclasses to use; the emulation update rate; the controller-side host and address; data output directories; \emph{et cetera}.
+This is done through Python scripts which define a number of required (and some optional) top-level variables, corresponding to emulation parameters such as the \emph{State}, \emph{Controller}, \emph{Sensor}, and \emph{Actuator} subclasses to use; the emulation update rate; the controller-side host and address; data output directories; \emph{et cetera}.
 These Python configuration files are then provided as arguments to the \texttt{cleave} command line provided by the framework.
@@
 Although the core design of \ac{CLEAVE} follows a single-loop approach, multi-loop scenarios can be easily implemented through the combination of multiple \ac{CLEAVE} instances.
-Furthermore, the framework makes no assumptions about the individual components other than the signatures of methods on the base classes that must be extended and implemented.
-Combined with the plain-Python emulation configuration scripts, this makes development of more complex scenarios merely a matter of implementing the desired functionality within the framework itself.
+The framework doesn't assume the individual components other than the signatures of methods on the base classes that must be extended and implemented.
+Combined with the Python emulation configuration scripts, this makes development of more complex scenarios merely a matter of implementing the desired functionality within the framework itself.
 

### 2022-02-11 intro.tex
Update on Overleaf.

@@
 
-A subcategory of \acp{CPS} which stands to particularly benefit from the adoption of the above technologies is \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system.
-Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for the communication between components that conventional cloud paradigms and cellular networks are unable to meet~\cite{Wan2020Efficient}.
+\acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system will benefit from the adoption of these technologies.
+Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks can't meet\cite{Wan2020Efficient}.
 
-Due to their potential advantages for industrial and commercial settings, a large body of work exists dedicated to the modelling and performance characterization of \acp{NCS}~\cite{Zhang2013Survey,Zhang2016Survey}.
-They improve and flexibilize existing control systems by allowing for the distribution of control functions across networks, allowing for centralized coordination, control, and monitoring of a multitude of devices.
+There are multiple works dedicated to the modelling and performance characterization of \acp{NCS}~\cite{Zhang2013Survey,Zhang2016Survey}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
 
@@
 One such approach uses setups in which the complete system is built on top of real hardware.
-This approach is employed in the works of, for instance, \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement their approach on physical testbeds.
+This approach is employed in the works of  \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement their approach on physical testbeds.
 Conversely, other studies choose to instead use completely \emph{simulated} \ac{NCS} setups.
 \textcite{Wu2012NPC,Chen2015synccontrol,Ma2019DynamicSched} are examples of works in which the authors have opted for such completely virtualized approaches.
-These studies often employ combinations of physical and network simulation tools to try to capture the complex dynamics of \acp{NCS}.
+These studies often employ combinations of physical and network simulation tools trying to capture the complex dynamics of \acp{NCS}.
 Finally, some experimental studies instead employ \emph{virtualized} approaches, in which either
@@
 As evidenced above, experimental research in \acp{NCS} includes varied heterogeneous hardware and software platforms, methodologies and \aclp*{KPI}.
+This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
 
-This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
 Furthermore, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
@@
 \textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge in their work.
-Their proposed platform \emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built utilizing LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform, with stated goals of reproducibility and ease of use.
-It achieves them thanks to its low-cost, accessibility, and ease of reconfiguration.
-
-Although it is an important step towards the reproducibility of \ac{NCS} benchmarks, we believe it still falls short on some aspects.
-In particular, we argue the reliance on specific hardware for the Plant which although flexible, limits the benchmarking to only those systems which can be physically implemented on it.
-Furthermore, it is \emph{not} scalable, and thus unsuitable for studies of larger, distributed \acp{NCS}.
+Their proposed platform \emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built utilizing LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform.
+It is reproducible because of its low-cost and accessibility.
+Although important for the reproducibility of \ac{NCS} benchmarks, we believe it is still inadequate given the reliance on specific hardware for the Plant which although flexible, limits the benchmarking to only those systems which can be physically implemented.
+It is \emph{not} scalable, and thus unsuitable for distributed \acp{NCS}.
 
 In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \ac{NCS}.
-As the edge computing paradigm starts to be adopted by industry, more and more variations of it have begun to appear in literature.
-``Near'', ``far'', ``core'', and ``telco'' edge, among others, are all terms which describe variations of the original concept and which are becoming ubiquitous in new research.
+As edge computing begins being adopted by industry, more and more variations have begun to appear in literature.
+``Near'', ``far'', ``core'', and ``telco'' edge describe variations of the original concept and are becoming ubiquitous in new research.
 While the core idea of edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts and weaknesses of such different edge concepts is of paramount importance.
@@
 Our framework, \ac{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems.
-\ac{CLEAVE} is fully virtualized, inspired by our previous work on benchmarking human-in-the-loop applications on the edge~\cite{Olguin2019EdgeDroid}.
+It is fully virtualized, inspired by our previous work on benchmarking human-in-the-loop applications on the edge~\cite{Olguin2019EdgeDroid}.
 The tool consists of a benchmarking framework and software development kit for the development of emulated physical systems and softwarized controllers.
@@
 
-\ac{CLEAVE} is built using \emph{Python 3.8}, making it highly extensible and able to harness the multitude of already existing user-provided libraries.
+\ac{CLEAVE} is built using \emph{Python 3.8}, making it highly extensible and able to harness the multitude of already existing user libraries.
 It is furthermore compatible with container technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment, scaling, and benchmarking on industry-standard edge setups using container orchestration solutions.
@@
 In \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.
-Finally, in \cref{sec:conclusion} we conclude this paper and briefly discuss future work.
+Finally, in \cref{sec:conclusion} we conclude this work.

### 2022-02-11 conclusion.tex
rewriting

@@
 It is fully parametrizable, extendable, and built with industry-standard edge technologies and paradigms in mind.
-We validate the utility of this tool through a series of scenarios; our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
+We validate the utility of this tool through an example use case showcasing the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
 We believe \ac{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \acp{NCS}.

### 2022-02-11 experiments.tex
rewriting

@@
-\section{Experimental validation}\label{sec:experiments}
+\section{Use case validation}\label{sec:experiments}
 
-In this section, we demonstrate the utility of \ac{CLEAVE} through a series of experiments on an emulated \ac{NCS} running on a number of client devices and a cloudlet.
-We aim to answer questions relating to the ability of our framework to provide accurate and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
+In this section, we demonstrate the utility of \ac{CLEAVE} by walking readers through an example use-case for the tool.
+With this, we aim to answer questions relating to the ability of our framework to provide accurate and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
 
-This section is structured in two parts.
-\Cref{ssec:expsetup} details the experimental setup and describes the experiments performed.
-\Cref{ssec:results} then presents and discusses the numerical results.
+We present the following use case scenario.
+A simple \ac{NCS} consisting of an \emph{inverted pendulum} plant controlled by a \emph{proportional-differential} controller is to be deployed on an Edge server.
+The connection between plant and server is over a WiFi (IEEE 802.11n) link.
+Moreover, there are a number of video-streaming applications running concurrently on this Edge setup.
+Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
+However, before deployment can become a reality, key questions such as: 
+``what are the baseline requirements of the \ac{NCS}?''; ``what are the achievable best-case end-to-end latencies in the system?''; and
+``how does the concurrent video-streaming traffic affect \ac{NCS} stability?'' need to be answered.
 
-\subsection{Experimental Setup}\label{ssec:expsetup}
-
-We detail our experimental edge setup in \cref{fig:cleave:expsetup}.
-It consists of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose \verb|x86_64| desktop computer acting as a Cloudlet/Edge server.
+We attempt to study these questions using the experimental setup in \cref{fig:cleave:expsetup}.
+\ac{CLEAVE} is deployed on a testbed consisting of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose \verb|x86_64| desktop computer acting as a Cloudlet/Edge server.
 On top of this physical architecture we configure a Docker Swarm\footnote{Swarm mode overview: \url{https://docs.docker.com/engine/swarm/}} managed centrally from the Cloudlet.
-For each experimental scenario we deploy a number of control loops inside this Swarm, executing plant containers exclusively in the Raspberry Pi clients (up to a single plant per client) and controller containers in the Cloudlet.
+We deploy varying numbers of control loops inside this Swarm, executing plant containers exclusively in the Raspberry Pi clients and controller containers in the Cloudlet.
 Plants and controllers communicate using the \ac{UDP} over an overlay network which sits on top of and abstracts away the real network configuration.
-Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which plants and controllers are executed co-located on the Cloudlet.
 
@@
 
-\subsubsection{Single Loop Baseline Scenarios}
-
-We first run a series of single loop scenarios with varying parametrization of the \ac{NCS}, intended as initial baselines to evaluate the accuracy of the \ac{CLEAVE} framework and showcase its flexibility.
-
-For these scenarios, we vary:
+The first step in aswering the questions posed above is to evaluate the baseline performance of the control system, both locally and over the network.
+We first configure \ac{CLEAVE} to run a single loop with both plant and controller co-located on the Edge server; next we configure an identical loop to be executed with the plant located on a Raspberry Pi client and the controller on the Edge server.
+For each setup, we execute a series of scenarios varying:
 \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
-    \item the sampling rate of the Plant state, setting it to \SIlist[list-final-separator={, or }]{10;20;40;60}{\hertz}
+    \item the sampling rate of the Plant state, setting it to \SIlist[list-final-separator={, or }]{5;10;20;40;60}{\hertz}
     \item the responsiveness of the Controller, adding fixed delays of \SIlist[list-final-separator={, or }]{0;25;50}{\milli\second} after the processing of each sample.
@@
 
-We repeat each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups; experiments with sampling rates \( \geq \) \SI{20}{\hertz} and artificial delays \( \geq \) \SI{25}{\milli\second} were repeated an additional \num{20} times for better statistical significance.
-Each repetition lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
+We run repeated executions of each combination of these parameters at least \num{10} times, for both the networked and ``local-only'' setups.
+Each execution lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
+After the initial \num{10} runs, we identify that setups with sampling rates of \SIlist{5;10}{\hertz} were consistently too unstable to consider, and discard them from further analyses.
+For all other setups, we further note that setups with no additional delay are always stable and thus decided to also discard them from further analyses.
+Remaining setups, i.e.\ sampling rates \( \geq \) \SI{20}{\hertz} and artificial delays \( \geq \) \SI{25}{\milli\second}, are then repeated an additional \num{20} times for better statistical significance.
 
+This repeated experimentation and data collection while ``zooming in'' on particular setups is facilitated by \ac{CLEAVE}s design.
 Scenarios are executed automatically in batches using a simple Python script which interacts with Docker through the widely adopted \emph{docker-py}\footnote{Docker SDK for Python: \url{https://docker-py.readthedocs.io/en/stable/}} library.
@@
 
-\subsubsection{Realistic Scenario with Network Resource Contention}
-
-Next, we run a multi-loop scenario to validate the utility of \ac{CLEAVE} in a more realistic setting where network resources are shared with video stream traffic.
-Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
-
-In this scenario, we deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
-On the remaining \num{4} clients we run the \emph{iperf3}\footnote{iperf3: \url{https://iperf.fr/}} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of uplink \ac{UDP} traffic.
-This emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
-We execute this scenario with \ac{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
-Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
-Once again, repetitions of this scenario are executed automatically in batches using a simple Python script and \emph{docker-py}.
-
-\subsection{Results}\label{ssec:results}
-
-The results presented below provide valuable insights on both system limits and on the chosen \ac{NCS} itself, as well as on the capabilities of \ac{CLEAVE}.
-
-\subsubsection{Single Plant Scenario}
-
 \begin{figure}[t]
     \centering
-    \begin{subfigure}[t]{.45\textwidth}
+    \begin{subfigure}[t]{\columnwidth}
         \centering
@@
     \end{subfigure}\\
-    \begin{subfigure}[t]{.45\textwidth}
+    \begin{subfigure}[t]{\columnwidth}
         \centering
@@
     \end{subfigure}\\
-    \begin{subfigure}[t]{.45\textwidth}
+    \begin{subfigure}[t]{\columnwidth}
         \centering
-        \includegraphics[width=.9\columnwidth]{single_loop_rtts}
+        \includegraphics[width=\textwidth]{single_loop_rtts}
         \caption{
@@
 
-We begin with a simple analysis of the stability of the Plant, as \ac{CLEAVE} allows for straightforward and repeatable collection and posterior processing of relevant metrics.  
 \cref{fig:single} shows a summary of the results relating to the single-loop scenarios:
 \cref{fig:single:topple} shows the fraction of plants that toppled in each scenario, and \cref{fig:single:rms} shows the average \ac{RMS} for the absolute pendula angles.
-Toppled plants were identified simply by analysis of the emulation data, indentifying instances whose pendulum angles reached values above a threshold value\footnote{This value depends on the parametrization of the controller; in this case it corresponds to \SI{0.35}{\radian}.}.
-
-As expected, higher sampling rates tend to correlate with better quality of control; at higher sampling rates the system was able to reach stability at higher \acp{RTT}.
-These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.
-For instance, it is clear from \cref{fig:single:topple,fig:single:rms} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
-A corollary of this is, conversely, that at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
-Adaptive sampling might thus be a viable method for optimizing \ac{NCS} resource utilization at the edge.
-
-For completeness, we also showcase \ac{CLEAVE}'s ability to obtain network performance data from the experiments.
 \cref{fig:single:rtt} shows average latency due to network and processing (excluding synthetic delays) for both single-loop scenarios.
@@
 
-\subsubsection{Realistic Scenario}
+As expected, higher sampling rates tend to correlate with better quality of control; at higher sampling rates the system was able to reach stability at higher \acp{RTT}.
+These initial results already hint at interesting consequences for such an edge-bound \ac{NCS} deployment.
+For instance, it is clear from \cref{fig:single:topple,fig:single:rms} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
+A corollary of this is, conversely, that at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
+Adaptive sampling might thus be a viable method for optimizing resource utilization.
+
+Now that we have established baselines for single loops in the use case scenario, we can proceed to studying the interaction between the \acp{NCS} and the video-streaming applications.
+We deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
+On the remaining \num{4} clients we run the \emph{iperf3}\footnote{iperf3: \url{https://iperf.fr/}} traffic load generator, each generating \SI[per-mode=symbol]{6.5}{\mega\bit\per\second} of uplink \ac{UDP} traffic.
+This emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
+Based on the baseline results obtained beforehand, we execute setups with \ac{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
+Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
+Once again, repetitions of this scenario are executed automatically in batches using a simple Python script and \emph{docker-py}.
 
@@
 
-%(circa \SI{41}{\milli\second} versus \SIrange[range-phrase=--]{13}{15}{\milli\second}).
-
 These results stem from the contention for network resources, and hint at important trade-offs system designers will have to take into consideration when designing and developing \acp{NCS} for deployment on the edge.

### 2022-02-11 intro.tex
Update on Overleaf.

@@
 
-There are multiple works dedicated to the modelling and performance characterization of \acp{NCS}~\cite{Zhang2013Survey,Zhang2016Survey}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
+\cite{Zhang2016Survey} is dedicated to the modelling and performance characterization of \acp{NCS}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
 
@@
 One such approach uses setups in which the complete system is built on top of real hardware.
-This approach is employed in the works of  \textcite{Li2014Wireless,Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement their approach on physical testbeds.
+This approach is employed in the works of  \textcite{Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement their approach on physical testbeds.
 Conversely, other studies choose to instead use completely \emph{simulated} \ac{NCS} setups.
-\textcite{Wu2012NPC,Chen2015synccontrol,Ma2019DynamicSched} are examples of works in which the authors have opted for such completely virtualized approaches.
+The authors in \cite{Ma2019DynamicSched} have opted for such a completely virtualized approach.
 These studies often employ combinations of physical and network simulation tools trying to capture the complex dynamics of \acp{NCS}.

### 2022-02-11 experiments.tex
trimming

@@
 
-In this section, we demonstrate the utility of \ac{CLEAVE} by walking readers through an example use-case for the tool.
-With this, we aim to answer questions relating to the ability of our framework to provide accurate and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
+In this section, we demonstrate the utility of \ac{CLEAVE} by walking readers through an example use case.
+With this, we aim to showcase the ability of our framework to provide accurate and repeatable measurements of the performance of \acp{NCS} deployed on edge computing infrastructure.
 
@@
 The connection between plant and server is over a WiFi (IEEE 802.11n) link.
-Moreover, there are a number of video-streaming applications running concurrently on this Edge setup.
+Moreover, there are a number of video-streaming applications running concurrently on the Edge setup.
 Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
@@
 We attempt to study these questions using the experimental setup in \cref{fig:cleave:expsetup}.
-\ac{CLEAVE} is deployed on a testbed consisting of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose \verb|x86_64| desktop computer acting as a Cloudlet/Edge server.
-On top of this physical architecture we configure a Docker Swarm\footnote{Swarm mode overview: \url{https://docs.docker.com/engine/swarm/}} managed centrally from the Cloudlet.
-We deploy varying numbers of control loops inside this Swarm, executing plant containers exclusively in the Raspberry Pi clients and controller containers in the Cloudlet.
-Plants and controllers communicate using the \ac{UDP} over an overlay network which sits on top of and abstracts away the real network configuration.
+\ac{CLEAVE} is deployed on a testbed consisting of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}.
+Connected to the Ethernet backbone of this \ac{AP} is a general-purpose \verb|x86_64| desktop computer acting as a Cloudlet/Edge server.
 
@@
 
-The first step in aswering the questions posed above is to evaluate the baseline performance of the control system, both locally and over the network.
-We first configure \ac{CLEAVE} to run a single loop with both plant and controller co-located on the Edge server; next we configure an identical loop to be executed with the plant located on a Raspberry Pi client and the controller on the Edge server.
-For each setup, we execute a series of scenarios varying:
+The first step in our case study is to evaluate the baseline performance of the control system, both locally and over the network.
+We configure \ac{CLEAVE} to run a single loop with both plant and controller co-located on the Edge server, and we also configure a separate setup with an identical loop executed over the wireless link.
+For both of these setups, we execute a series of scenarios varying:
 \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
@@
 Each execution lasts for \SI{5}{\minute}, during which we collect detailed data on both the state of the controlled system as well as on the data sent over the network.
-After the initial \num{10} runs, we identify that setups with sampling rates of \SIlist{5;10}{\hertz} were consistently too unstable to consider, and discard them from further analyses.
-For all other setups, we further note that setups with no additional delay are always stable and thus decided to also discard them from further analyses.
-Remaining setups, i.e.\ sampling rates \( \geq \) \SI{20}{\hertz} and artificial delays \( \geq \) \SI{25}{\milli\second}, are then repeated an additional \num{20} times for better statistical significance.
+After the initial \num{10} runs, we identify that setups with sampling rates of \SIlist{5;10}{\hertz} were consistently too unstable to consider, and disregard them in further analyses.
+We further note that setups with \SI{0}{\milli\second} additional delay are always stable and thus also disregard them.
+Remaining setups, i.e.\ sampling rates \( \geq \) \SI{20}{\hertz} and artificial delays \( \geq \) \SI{25}{\milli\second}, are repeated an additional \num{20} times for better statistical significance.
 
-This repeated experimentation and data collection while ``zooming in'' on particular setups is facilitated by \ac{CLEAVE}s design.
+This repeated experimentation and data collection while ``zooming in'' on particular setups is facilitated by \ac{CLEAVE}'s design.
 Scenarios are executed automatically in batches using a simple Python script which interacts with Docker through the widely adopted \emph{docker-py}\footnote{Docker SDK for Python: \url{https://docker-py.readthedocs.io/en/stable/}} library.
@@
 
-Now that we have established baselines for single loops in the use case scenario, we can proceed to studying the interaction between the \acp{NCS} and the video-streaming applications.
+Once baselines for single loops in the use case scenario have been established, we can proceed to studying the interaction between the \acp{NCS} and the video-streaming applications.
 We deploy \num{6} control loops on the experimental setup depicted in \cref{fig:cleave:expsetup}.
@@
 This emulates the load generated on the network by \SI{1080}{p} Full-HD video streaming, originating from the clients and terminating in the cloudlet.
-Based on the baseline results obtained beforehand, we execute setups with \ac{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
+Based on the baseline results, we execute setups with \ac{NCS} plant sampling rates of \SIlist{20;40;60}{\hertz}.
 Each sampling rate configuration is run for \SI{5}{\minute}, and then repeated \num{30} times to obtain statistical significance.
@@
 
-\cref{fig:video:results} shows a summary of the results obtained from the realistic scenario.
-\cref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
-\cref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
-\cref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
-\cref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
+\cref{fig:video:results} shows a summary of the results obtained.
+\cref{fig:video:toppled} shows the fraction of repetitions of each setup in which \emph{at least} one plant failed to maintain stability and toppled.
+\cref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each setup, only considering data from plants that did not topple.
+\cref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per setup.
+\cref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per setup.
 
@@
 Whereas both the \SIlist{20;40}{\hertz} setups show losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} of datagrams lost.
-The differences in \acp{RTT} results are equally telling; \acp{RTT} for the \SI{60}{\hertz} scenario were on average approx.\ \num{3} times those for the \SIlist{20;40}{\hertz} setups.
+The differences in \acp{RTT} results are equally telling; \acp{RTT} for the \SI{60}{\hertz} setup were on average approx.\ \num{3} times those for the \SIlist{20;40}{\hertz} setups.
 

### 2022-02-11 abstract.tex
abstract

@@
     We validate \acs*{CLEAVE} using an initial implementation of an inverted pendulum \acs*{NCS}.
-    Our results showcase the utility of the tool as a repeatable, extensible, and scalable solution to \acs*{NCS} performance evaluation and benchmarking.
+    Our results showcase the utility of the tool as a repeatable, extensible, and scalable solution to \acs*{NCS} performance evaluation and benchmarking on the Edge.
 \end{abstract}

### 2022-02-11 intro.tex
use case intro
Update on Overleaf.

@@
 
-\cite{Zhang2016Survey} is dedicated to the modelling and performance characterization of \acp{NCS}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
+Due to their potential advantages for industrial and commercial settings, there exist works\cite{Zhang2016Survey} dedicated to the modelling and performance characterization of \acp{NCS}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
 
@@
 \cref{sec:approach} presents the design principles and implementation of the framework.
-In \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.
+In \cref{sec:experiments}, we present a use case scenario which validates the utility, flexibility, and repeatability of \ac{CLEAVE}.
 Finally, in \cref{sec:conclusion} we conclude this work.

### 2022-02-11 intro.tex
Update on Overleaf.

@@
 \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system will benefit from the adoption of these technologies.
-Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks can't meet\cite{Wan2020Efficient}.
+Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks are unable to meet\cite{Wan2020Efficient}.
 

### 2022-02-11 intro.tex
fixes

--- 2022-02-11 intro.tex
@@
 \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system will benefit from the adoption of these technologies.
-Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks can't meet\cite{Wan2020Efficient}.
+Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks are unable to meet~\cite{Wan2020Efficient}.
 
-Due to their potential advantages for industrial and commercial settings, there exist works\cite{Zhang2016Survey} dedicated to the modelling and performance characterization of \acp{NCS}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
+Due to their potential advantages for industrial and commercial settings, there exist works~\cite{Zhang2016Survey} dedicated to the modelling and performance characterization of \acp{NCS}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
 

### 2022-02-11 intro.tex
Update on Overleaf.

--- 2022-02-11 intro.tex
@@
 \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system will benefit from the adoption of these technologies.
-Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks can't meet\cite{Wan2020Efficient}.
+Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks cannot meet\cite{Wan2020Efficient}.
 

### 2022-02-11 intro.tex

--- 2022-02-11 intro.tex
@@
 \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system will benefit from the adoption of these technologies.
-Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks are unable to meet~\cite{Wan2020Efficient}.
+Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks cannot meet\cite{Wan2020Efficient}.
 

### 2022-02-11 intro.tex
space

@@
 \acp{NCS}~\cite{Gupta2010NCSOverview}, a type of \ac{CPS} wherein multiple networked actuators and sensors form a part of the same automatic control system will benefit from the adoption of these technologies.
-Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks cannot meet\cite{Wan2020Efficient}.
+Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks cannot meet~\cite{Wan2020Efficient}.
 

### 2022-02-11 intro.tex
kpi

@@
 
-As evidenced above, experimental research in \acp{NCS} includes varied heterogeneous hardware and software platforms, methodologies and \aclp*{KPI}.
+As evidenced above, experimental research in \acp{NCS} includes varied heterogeneous hardware and software platforms, methodologies and key performance indicators.
 This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.

### 2022-02-11 experiments.tex
experiments

@@
     \caption[caption]{
-        Metrics for the single-loop scenarios.
-        % \labelcref{fig:single:topple} shows the fraction of plants that toppled, per experimental setup.
-        % \labelcref{fig:single:rms} shows the mean \ac{RMS} value of the angle with respect to the vertical axis in logarithmic scale; the red line indicates \( y = 180 \).
+        Baseline results.
         Error bars indicate \SI{95}{\percent} \acp{CI}.
@@
     \caption{
-        Results for the realistic scenario.
-        % \labelcref{fig:video:toppled} shows the fraction of repetitions of each scenario in which \emph{at least} one plant failed to maintain stability and toppled.
-        % \labelcref{fig:video:rms} shows the \ac{RMS} for the pendulum angles for each scenario, only considering data from plants that did not topple.
-        % \labelcref{fig:video:drop} shows the fraction of \ac{UDP} datagrams dropped, averaged over all plants and repetitions per scenario.
-        % \labelcref{fig:video:rtt} shows the measured end-to-end plant-side \ac{RTT}, averaged over all plants and repetitions per scenario.
+        Multi-loop, resource-constrained setup results.
         Error bars indicate \SI{95}{\percent} \acp{CI} in all plots.
@@
 
-These results are interesting in their counter-intuitiveness.
-Conventional wisdom would lead us to think that higher sampling rates are always better for the stability of control systems; however, \cref{fig:video:toppled,fig:video:rms} clearly show this not to be the case for \ac{NCS} in resource-constrained scenarios.
+These results are interesting in their counter-intuitiveness compared to the single-loop baseline results.
+While the baselines might lead us to think that higher sampling rates are always better for the stability of control systems, \cref{fig:video:toppled,fig:video:rms} show this not to be the case for \ac{NCS} in resource-constrained scenarios.
 \SI{60}{\hertz} was the least stable configuration, with at least one pendulum toppling in around \SI{16}{\percent} of the repetitions, and mean pendulum angle \ac{RMS} circa \num{3} times that of the \SI{40}{\hertz} scenario.
@@
 
-We can see an explanation for these behaviors in \cref{fig:video:drop,fig:video:rtt}.
+\cref{fig:video:drop,fig:video:rtt} explains these behaviors.
 Whereas both the \SIlist{20;40}{\hertz} setups show losses well below \SI{5}{\percent}, the \SI{60}{\hertz} scenario shows an average of around \SI{13}{\percent} of datagrams lost.
 The differences in \acp{RTT} results are equally telling; \acp{RTT} for the \SI{60}{\hertz} setup were on average approx.\ \num{3} times those for the \SIlist{20;40}{\hertz} setups.
-
-These results stem from the contention for network resources, and hint at important trade-offs system designers will have to take into consideration when designing and developing \acp{NCS} for deployment on the edge.
-The edge will be \emph{multi-tenant} and \emph{multi-instance}. 
+These results stem from the contention for network resources, and hint at important trade-offs system designers will have to take into consideration when designing and developing \acp{NCS} for deployment on the Edge.
+The Edge will be \emph{multi-tenant} and \emph{multi-instance}. 
 \ac{NCS} deployments will have to be designed with shared resources in mind, and given the complexity of these systems, experimental tools like \ac{CLEAVE} will be key for their succesful adoption and massification.

### 2022-02-11 intro.tex
line

@@
 Depending on the physical system being controlled, \acp{NCS} can have stringent timing and reliability requirements for communication that conventional cloud paradigms and cellular networks cannot meet~\cite{Wan2020Efficient}.
+This necessitates sophisticated tools for the performance evaluation of future system architectures, as well as novel NCS design paradigm.
 
-Due to their potential advantages for industrial and commercial settings, there exist works~\cite{Zhang2016Survey} dedicated to the modelling and performance characterization of \acp{NCS}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
+%Due to their potential advantages for industrial and commercial settings, there exist works~\cite{Zhang2016Survey} dedicated to the modelling and performance characterization of \acp{NCS}, improving NCSs by distributing control functions across networks, facilitating centralized coordination, control, and monitoring.
 
-% Most of the literature concerning \acp{NCS}, follows a theoretical approach, and only a small fraction of it deals with experimental studies.
-A number of works concerning \acp{NCS} deal with experimental studies.
-\acp{NCS} have an inherently inter-domain nature intertwining knowledge from the fields of communications, computing, and control theory in ways that cannot be studied in isolation, leading to various different approaches to such studies.
+One the one hand, related literature in \acp{NCS} leverages to a large extent theoretical models, at the price of being able to capture networked systems effects only on a coarse level. 
+%follows a theoretical approach, and only a small fraction of it deals with experimental studies.
+On the other, there exist several approaches when considering experimental methodologies. 
+%A number of works concerning \acp{NCS} deal with experimental studies.
+%\acp{NCS} have an inherently inter-domain nature intertwining knowledge from the fields of communications, computing, and control theory in ways that cannot be studied in isolation, leading to various different approaches to such studies.
 One such approach uses setups in which the complete system is built on top of real hardware.
@@
 This, in turn, leads to hardware, software, and methodology fragmentation, as different studies tend to prefer approaches more favored in their respective communities.
-
 Furthermore, existing studies tend to focus on individual aspects and components of a system, thus producing results which do not provide a complete image of the \ac{NCS}.
@@
 Their proposed platform \emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built utilizing LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform.
-It is reproducible because of its low-cost and accessibility.
+It is  because of its low-cost and accessibility.
 Although important for the reproducibility of \ac{NCS} benchmarks, we believe it is still inadequate given the reliance on specific hardware for the Plant which although flexible, limits the benchmarking to only those systems which can be physically implemented.

### 2022-02-11 cleave.tex
use case
fixes

@@
 Our design follows a virtualized emulation approach; providing a framework for the real-time \emph{emulation} of physical control systems and the easy implementation of softwarized controllers interacting with \emph{real} networks.
-\ac{CLEAVE} is thus, a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
+\ac{CLEAVE} is thus a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
 This allows for great flexibility, as the core components of the \ac{NCS} can be switched out while maintaining the realism of the network, the most complex and limiting component of edge \acp{NCS}.
@@
 Although the core design of \ac{CLEAVE} follows a single-loop approach, multi-loop scenarios can be easily implemented through the combination of multiple \ac{CLEAVE} instances.
-The framework doesn't assume the individual components other than the signatures of methods on the base classes that must be extended and implemented.
+The framework does not make any assumptions about the individual components other than the signatures of methods on the \acp{ABC} that must be extended and implemented.
 Combined with the Python emulation configuration scripts, this makes development of more complex scenarios merely a matter of implementing the desired functionality within the framework itself.
@@
     \centering
-    \includegraphics[width=.7\columnwidth]{images/inverted_pendulum.png}
+    \includegraphics[width=.6\columnwidth]{images/inverted_pendulum.png}
     \caption{
@@
 
-The inverted pendulum plant \emph{State} is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz} (this value is configurable; in this case it corresponds to the maximum achievable stable rate on the clients).
+The inverted pendulum plant \emph{State} is implemented as a real-time discrete-time physical emulation using CLEAVE's API and a 2D physics library\footnote{Pymunk: \url{http://www.pymunk.org/en/latest/}}, updated at a constant \SI{120}{\hertz} (this value is configurable; in this case it corresponds to the maximum achievable stable rate on our available hardware).
 \emph{Sensors} for the angle, position, angular velocity, and velocity of the \emph{State}, and an \emph{Actuator} for the horizontal force applied to the cart are implemented as ``perfect'', i.e.\ values are returned as-is, without any added noise.

### 2022-02-11 intro.tex

@@
 Their proposed platform \emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built utilizing LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform. 
-It is  because of its low-cost and accessibility.
+NCSbench facilitates to some extent reproducible and It is  because of its low-cost and accessibility. 
 Although important for the reproducibility of \ac{NCS} benchmarks, we believe it is still inadequate given the reliance on specific hardware for the Plant which although flexible, limits the benchmarking to only those systems which can be physically implemented.
 It is \emph{not} scalable, and thus unsuitable for distributed \acp{NCS}.
-
+(((( Zoppi \textit{et al.} in \cite{9045199} proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS.
+Their methodology utilizes joint knowledge of control, computation, and communication.
+In their work various architectural elements and the corresponding delays associated with the NCS are modelled.
+Multiple experimental parameters and certain observable Key Performance Indicators (KPIs) are defined and utilized in the implementation.
+This work however utilizes a physical LEGO \cite{lego} based plant for the implementation, preventing instantaneous changes in the plant characteristics and their nature, in terms of flexibility.
+We overcome this issue by proposing a completely virtual plant allowing for unparalleled flexibility in changing the plant model and characteristic features of the experiments.)))
 In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \ac{NCS}.

### 2022-02-11 intro.tex
Update on Overleaf.

--- 2022-02-11 intro.tex
@@
 \textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge in their work.
-Their proposed platform \emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built utilizing LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform.
-It is  because of its low-cost and accessibility.
-Although important for the reproducibility of \ac{NCS} benchmarks, we believe it is still inadequate given the reliance on specific hardware for the Plant which although flexible, limits the benchmarking to only those systems which can be physically implemented.
-It is \emph{not} scalable, and thus unsuitable for distributed \acp{NCS}.
+In their work, they proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS. 
+Their methodology utilizes joint knowledge of control, computation, and communication. 
+In their work various architectural elements and the corresponding delays associated with the NCS are modelled. 
+Multiple experimental parameters and certain observable key performance indicators are defined and utilized in the implementation. 
+This work however utilizes a physical LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\  based plant for the implementation, preventing instantaneous changes in the plant characteristics and their nature, in terms of flexibility.
+Furthermore, relying on physical objects like an inverted pendulum limits scalability of the experimentation in practise.
 
+We overcome this issue by proposing a completely virtual plant allowing for unparalleled flexibility in changing the plant model and characteristic features of the experiments.
 In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \ac{NCS}.
@@
 ``Near'', ``far'', ``core'', and ``telco'' edge describe variations of the original concept and are becoming ubiquitous in new research.
-While the core idea of edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts and weaknesses of such different edge concepts is of paramount importance.
+While the core idea of edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strengths and weaknesses of such different edge concepts is of paramount importance.
 

### 2022-02-11 intro.tex
fixes

@@
 \textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge in their work.
-In their work, they proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS. 
+In their work, they proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS.\ 
 Their methodology utilizes joint knowledge of control, computation, and communication. 
@@
 Multiple experimental parameters and certain observable key performance indicators are defined and utilized in the implementation. 
-This work however utilizes a physical LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\  based plant for the implementation, preventing instantaneous changes in the plant characteristics and their nature, in terms of flexibility.
-Furthermore, relying on physical objects like an inverted pendulum limits scalability of the experimentation in practise.
+This work however utilizes a physical LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\  based plant for the implementation, preventing instantaneous changes in plant characteristics and component parametrizations.
+Furthermore, relying on physical objects like an inverted pendulum limits scalability of the experimentation in practice.
 

### 2022-02-11 abstract.tex
fix abstract

@@
     As the number of cyber-physical systems rises, it becomes increasingly crucial to study \acfp*{NCS} combining control communication co-design.
-This nature of NCSs has led to task-specific approaches to research, creating a dearth of generalizable, repeatable, and scalable experimentation.
+This nature of \acsp*{NCS} has led to task-specific approaches to research, creating a dearth of generalizable, repeatable, and scalable experimentation.
 Further, with the advent of edge computing solutions, it is of paramount importance to explore its relevance in such applications.
-In this work, we present \acs*{CLEAVE}, a novel, completely software-based framework for repeatable and scalable experimentation in edge native NCSs.
+In this work, we present \acs*{CLEAVE}, a novel, completely software-based framework for repeatable and scalable experimentation in edge native \acsp*{NCS}.
     Our approach is based on the emulation of physical plants communicating over a real network with software-based controllers.
     \acs*{CLEAVE} is designed and built for the edge, using Python3 and with full compatibility with industry-standard containerization solutions.
-    Although designed for single-loop emulations, the flexibility afforded by the aformentioned characteristics allow our framework to be adapted to a multitude of complex scenarios.
+    Although designed for single-loop emulations, the flexibility afforded by the aforementioned characteristics allow our framework to be adapted to a multitude of complex scenarios.
 

### 2022-03-13 experiments.tex
adding outline of changes

@@
 Moreover, there are a number of video-streaming applications running concurrently on the Edge setup.
+
+{
+    \color{ACMRed}
+    Why is this an interesting use-case?
+    \begin{itemize}
+        \item Inverted pendulum is broadly used as a benchmark in NCS (already mentioned); worth mentioning related work which does the same thing!
+        \item Easy to re-parameterize for different system dynamics.
+        \item Partly a reality, similar to deployments in industry right now (although I'm pretty sure industry uses ZigBee or WirelessHART).
+        \item WiFi is a bit of a problem - best case would've been to have 4G or 5G.
+Argument could however be that 4G at least has higher RTTS, so if it doesn't work for WiFi it will definitely not work for 4G\dots
+        \item Video streaming/analytics are core use-case for edge computing (already in text).
+    \end{itemize}
+}
+
+\todo[inline]{
+    More strongly argue why this is an interesting and general use-case scenario.
+    Paper needs to argue and describe why this use case chosen is one of general interest, addressing key aspects of demands, and should refer to a comparative view with related work.
+}
+
 Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.

### 2022-03-13 intro.tex
no bibtex

@@
 One such approach uses setups in which the complete system is built on top of real hardware.
-This approach is employed in the works of  \textcite{Baumann2018LowPower,Cuenca2019UAV}; in all of these, the authors implement their approach on physical testbeds.
+This approach is employed in the works of Baumann \emph{et al.}~\cite{Baumann2018LowPower} and Cuenca \emph{et al.}~\cite{Cuenca2019UAV}; in both of these, the authors implement their approach on physical testbeds.
 Conversely, other studies choose to instead use completely \emph{simulated} \ac{NCS} setups.
@@
 
-\textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge in their work.
+Zoppi \emph{et al.}~\cite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge in their work.
 In their work, they proposed a platform called NCSBench, to be used for reproducible benchmarking in NCS.\ 

### 2022-03-13 experiments.tex
copyrights

@@
 Moreover, there are a number of video-streaming applications running concurrently on the Edge setup.
+
+We believe this to be an interesting and representative use-case scenario for a number of reasons.
+Firstly, the inverted pendulum is broadly used as a benchmark in control-system literature.
+\todo[inline]{Examples}
+A key reason for this is that such a relatively simplistic system allows for straightforward reparametrization to obtain varied system dynamics, which in turn makes a broader range of experiments possible.
+Secondly, similar, albeit non-Edge-enabled, setups have been a reality in industry for almost two decades, as evidenced by
+\todo[inline]{citation here pls}.
+Thirdly, although real deployments would most assuredly employ mobile technologies due to the added benefits of mobility and range, the \acp{RTT} offered by these technologies are higher, or at best equal, those offered by WiFi.
+4G, for instance, has average \acp{RTT} of around a few hundred milliseconds, which is orders of magnitude greater than the values measured in our scenario, and current 5G deployments offer \ac{RTT} values barely matching the sub-\SI{10}{\milli\second} values we obtain for WiFi.
+Thus we argue that using WiFi allows for more experimental freedom, as the system can be tweaked to obtain worse \acp{RTT} more akin to those of mobile technologies while still allowing us the option to study best-case scenarios with very low-latencies.
 

### 2022-03-13 experiments.tex
thank reviewers

@@
 Thus we argue that using WiFi allows for more experimental freedom, as the system can be tweaked to obtain worse \acp{RTT} more akin to those of mobile technologies while still allowing us the option to study best-case scenarios with very low-latencies.
-
-{
-    \color{ACMRed}
-    Why is this an interesting use-case?
-    \begin{itemize}
-        \item Inverted pendulum is broadly used as a benchmark in NCS (already mentioned); worth mentioning related work which does the same thing!
-        \item Easy to re-parameterize for different system dynamics.
-        \item Partly a reality, similar to deployments in industry right now (although I'm pretty sure industry uses ZigBee or WirelessHART).
-        \item WiFi is a bit of a problem - best case would've been to have 4G or 5G.
-Argument could however be that 4G at least has higher RTTS, so if it doesn't work for WiFi it will definitely not work for 4G\dots
-        \item Video streaming/analytics are core use-case for edge computing (already in text).
-    \end{itemize}
-}
-
-\todo[inline]{
-    More strongly argue why this is an interesting and general use-case scenario.
-    Paper needs to argue and describe why this use case chosen is one of general interest, addressing key aspects of demands, and should refer to a comparative view with related work.
-}
-
-Video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
+Finally, video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.
 However, before deployment can become a reality, key questions such as: 

### 2022-03-13 experiments.tex
use case validation

@@
 We believe this to be an interesting and representative use-case scenario for a number of reasons.
-Firstly, the inverted pendulum is broadly used as a benchmark in control-system literature.
-\todo[inline]{Examples}
+Firstly, the inverted pendulum is broadly used as a benchmark in control-system literature, see for instance\ \cite{Baumann2018LowPower} and\ \cite{Natale2004InvPendEthernet}.
 A key reason for this is that such a relatively simplistic system allows for straightforward reparametrization to obtain varied system dynamics, which in turn makes a broader range of experiments possible.
-Secondly, similar, albeit non-Edge-enabled, setups have been a reality in industry for almost two decades, as evidenced by
-\todo[inline]{citation here pls}.
+Secondly, similar, albeit non-Edge-enabled, setups have been a reality in industry for almost two decades, as evidenced by\ \cite{Gupta2010NCSOverview}.
 Thirdly, although real deployments would most assuredly employ mobile technologies due to the added benefits of mobility and range, the \acp{RTT} offered by these technologies are higher, or at best equal, those offered by WiFi.

### 2022-03-14 conclusion.tex
adding outline of changes
conclusion

@@
 
-In this work, we aim to tackle this issue by presenting a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking with a particular focus on edge deployment.
-Our framework, \ac{CLEAVE}, allows for the easy implementation of emulated \acp{NCS} on real networks.
-It is fully parametrizable, extendable, and built with industry-standard edge technologies and paradigms in mind.
-We validate the utility of this tool through an example use case showcasing the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
+In this work, we aimed to tackle this issue through a fully software-based framework for repeatable, reproducible, and easily scalable \ac{NCS} benchmarking with a particular focus on edge deployment.
+We argue our approach, \ac{CLEAVE}, embodies a better solution than previous work for a number of reasons:
+\begin{enumerate}
+    \item Compared to fully physical approaches, such as those used in\ \cite{Baumann2018LowPower} and\ \cite{Cuenca2019UAV}, our approach allows for greater flexibility and scalability.
+    The aforementioned approaches rely on specialized and sometimes entirely custom-built physical platforms, and although flexible and cheap approaches such as Zoppi \emph{et al.'s} --- which uses a LEGO-based physical plant --- exist, these still do not reach the level of flexibility afforded by a fully software-based framework.
+    Experimenters still would need copies of the hardware, making anything other than small-scale setups unfeasible.
+    In contrast, our approach requires only general-purpose computing platforms, and can be employed by basically anyone with access to a computer.
+    Scalable deployments can in turn easily and cheaply be set up using single-board computers and/or virtual cloud instances.
+    \item When compared to simulated approaches such as\ \cite{Ma2019DynamicSched}, \ac{CLEAVE} provides a higher level of realism, in particular with regards to the network segment of the system.
+    \item Finally, although it shares much in common with previous emulated approaches such as the one employed in\ \cite{Wang2020VoltageControl}, \ac{CLEAVE} has an advantage by specifically targeting a general-purpose approach using industry-standard, cloud- and edge-native tools and software.
+    The tool can easily be deployed and scaled using widely-used frameworks such as Docker Swarm and Kubernetes.
+\end{enumerate}
+
+We validate the utility of this tool through an example use case approximating a proposed edge deployment of inverted pendula control loops co-located with video analytics services.
+We argue such a use case represents a realistic scenario and appropriate benchmark for the tool, since
+\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
+    \item the inverted pendulum plant is ubiquitous in \ac{NCS} research
+    \item similar setups exist in real-world industrial use
+    \item video analytics has long been proposed as a ``killer app'' for edge computing.
+\end{enumerate*}
+Our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
 We believe \ac{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \acp{NCS}.

### 2022-03-14 cleave.tex
adding outline of changes
changes

@@
 
-\subsection{Design and Implementation}
+\subsection{Design and Implementation}\label{sec:design}
+
+The design of \ac{CLEAVE} attempts to target a broad a set of different control systems as possible, and tries to minimize the number of assumptions made about the systems implemented on top of the framework.
+These are:
+\begin{enumerate}
+    \item Controllers can be stateful or stateless, but they exclusively control a single plant.
+    \item Plants and controllers interact in a ``request-response'' manner.
+    We currently don't support setups in which the controller generates an actuation command without there first being a request from the plant.
+    \item Controller and plant communicate over a TCP/IP network.
+    \item The physical state behavior of the plant can be implemented in a discrete-time manner.
+    \item Values associated to sensors and actuators can be represented by integers, floating point numbers, booleans, or arrays of bytes.
+    \item A plant has zero or more sensors attached to it, with potentially different sampling rates.
+    \item A plant has zero or more actuators attached to.
+\end{enumerate}
 
@@
 We have chosen this system as an initial benchmark for its relative simplicity as well as prevalence in the field of automatic control as one of the fundamental examples of linear control.
-However, as mentioned above, we eventually plan to build an open library of emulated control loops.
 
@@
 Plant and controller are then packaged into containers, for ease of orchestration and reparametrization, and to mimic real-world deployment.
+
+A previously mentioned, we eventually plan to build an open library of emulated control loops.\footnote{%
+    At the moment of writing, we are in the process of implementing \emph{model-predictive} as well as \emph{deep neural network} controllers for the inverted pendula plants.%
+}
+The inverted pendulum plant is merely a proof-of-concept, and \ac{CLEAVE} could be used to implement much more complex physical systems with even more stringent latency requirements.
+For instance, the flexibility afforded by the choice of programming language and the abstractions detailed in \cref{sec:design}, allow for relatively straightforward emulation of autonomous drone dynamics.
+Propeller dynamics could be implemented by \emph{Actuator} objects, gyroscopes by \emph{Sensors}, and the physical behavior of the plant could be modeled using a general purpose 3D physics engine (\emph{e.g.} the Panda3D\footnote{\url{https://docs.panda3d.org/1.10/python/programming/physics/builtin/index}{https://docs.panda3d.org/1.10/python/programming/physics/builtin/index}} engine).

### 2022-03-14 experiments.tex
changes

@@
 4G, for instance, has average \acp{RTT} of around a few hundred milliseconds, which is orders of magnitude greater than the values measured in our scenario, and current 5G deployments offer \ac{RTT} values barely matching the sub-\SI{10}{\milli\second} values we obtain for WiFi.
-Thus we argue that using WiFi allows for more experimental freedom, as the system can be tweaked to obtain worse \acp{RTT} more akin to those of mobile technologies while still allowing us the option to study best-case scenarios with very low-latencies.
+Thus, we argue that using WiFi allows for more experimental freedom, as the system can be tweaked to obtain worse \acp{RTT} more akin to those of mobile technologies while still having the option to study best-case scenarios with very low-latencies.
 Finally, video analytics is one of the main proposed use cases for edge computing~\cite{Ananthanarayanan2017Analytics,Yi2017Analytics,Wang2018Analytics}, and thus we foresee edge \ac{NCS} deployments being deployed in parallel with such applications in the future.

### 2022-03-14 cleave.tex
final?

@@
     \item The physical state behavior of the plant can be implemented in a discrete-time manner.
-    \item Values associated to sensors and actuators can be represented by integers, floating point numbers, booleans, or arrays of bytes.
+    \item Values associated to sensors and actuators can be represented as integers, floating point numbers, booleans, or sequences of bytes.
     \item A plant has zero or more sensors attached to it, with potentially different sampling rates.
-    \item A plant has zero or more actuators attached to.
+    \item A plant has zero or more actuators attached to it.
 \end{enumerate}
